---
title: "WKMAMBY outputs"
author: "WKMAMBY"
date: "2026-09-06"
output:
  html_document:
    lightbox: true
    toc: true
    toc_float: false
    toc_depth: 4
    code_folding: hide
    number_section: false
    df_print: paged
    css: styles.css
    self_contained: true
editor_options:
  chunk_output_type: inline
  markdown:
    wrap: 80
---

```{r setup, echo=FALSE}
operationnel <- TRUE
knitr::opts_chunk$set(
  message = !operationnel,
  warning = !operationnel,
  echo = !operationnel,
  progress = !operationnel,
  verbose = !operationnel,
  fig.show = "asis",
  fig.width = 12,
  fig.height = 10
)
```

```{r library_setup, message = FALSE, warning = FALSE}
# Load libraries
library(icesTAF)
library(here)
library(ggalluvial)
library("icesVocab")
library(sf)
library(ggplot2)
library(dplyr)
library(rnaturalearth)
library(rnaturalearthdata)
library(lwgeom)
library(viridis)
library(stringr)
library(ggforce)
library(readr)
library(cowplot)
library(here)
library(data.table)
library(dplyr)
library(sf)
library(ggrepel)
library(plotly)
library(rmapshaper)
library(readxl)
library(tidyverse)
library(leaflet)
library(mapview)
library(ggspatial)
library(paletteer)
library(ggnewscale)
library(raster)
library(stringr)
library(paletteer)
library(RColorBrewer)
here::i_am("DataComparisons.Rproj")
```

## Data preparation

Load ICES vocab

```{r ICES vocab}
# Extract the country codes, gear codes and monitoring program tables
codes <- getCodeList("BYC_MonProgr")
write.taf(codes,file="data/monprog.csv", row.names = F, quote = T)
codes <- getCodeList("BYC_MonMethod")
write.taf(codes,file="data/monmeth.csv", row.names = F, quote = T)
codes <- getCodeList("GearGroup")
write.taf(codes,file="data/gearcodes.csv", row.names = F, quote = T)
codes <- getCodeList("ISO_3166")
write.taf(codes,file="data/ctrycodes.csv", row.names = F, quote = T)

# Read and simplify the country codes table
ctrycodes <- read_csv("data/ctrycodes.csv") %>%
  transmute(country = Key, ctryname = tolower(Description)) %>%
  mutate(ctryname = paste0(toupper(substr(ctryname, 1, 1)), substr(ctryname, 2, 999))) %>%
  mutate(ctryname = ifelse(ctryname == "United kingdom of great britain and northern ireland (the)", 
                           "United Kingdom", ctryname)) %>%
  mutate(ctryname = ifelse(ctryname == "Netherlands (kingdom of the)", 
                           "Netherlands", ctryname)) %>%
  distinct()

# Read and simplify the monitoring methods names table
monmeth <- read_csv("data/monmeth.csv") %>%
  transmute(monitoringMethod = Key, monmethname = Description) %>%
  distinct()
# Read and simplify gear codes table
gearcodes <- read_csv("data/gearcodes.csv") %>%
  transmute(metierL3 = Key, L3name = Description) %>%
  distinct()
# Read and simplify th monitoring program names table
monprog <- read_csv("data/monprog.csv") %>%
  transmute(monitoringProgramType = Key, monprog = Description)%>%
  distinct()
```

Load data produced by ICES and chairs

```{r Load data and merge WKMAMBY and WGBYC data, message = F, results = "hide"}
# Get AUs info with prop by icesarea
# If needed, we can add areacodes that we can fetch here : https://gis.ices.dk/shapefiles/ICES_StatRec_mapto_ICES_Areas.zip
# Not needed here because the information is in the AUs file
#statrec_map_areas <- here("data/ICES_StatRec_mapto_ICES_Areas/StatRec_map_Areas_Full_20170124.shp")
#statrec_map_areas <- st_read(statrec_map_areas)
AUsfile <- here("data/20260605_OSPAR_IA2029_AssessmentUnits_WKMAMBY_wgs84.gpkg")
geoinfo <- suppressWarnings(suppressMessages(st_read(AUsfile,
                                                     layer = "ICES statistical rectangles")))

# ------------------------------------------------------------
# Common dolphin disaggregated AUs
# Official MSFD Atlantic subregions from EEA
# ------------------------------------------------------------

msfd_url <- paste0(
  "https://marine.discomap.eea.europa.eu/",
  "arcgis/rest/services/Marine/Marine_regions_subregions/",
  "MapServer/0/query?",
  "where=1%3D1",
  "&outFields=*",
  "&returnGeometry=true",
  "&f=geojson"
)

# Read MSFD Atlantic subregions
msfd <- suppressMessages(
  sf::st_read(
    msfd_url,
    quiet = TRUE
  )
)

# Names expected from the WKMAMBY AU definition
cd_dis_names <- sort(
  unique(
    na.omit(
      as.character(
        geoinfo$common_dolphin_disaggregated
      )
    )
  )
)

# Keep exactly the three Common dolphin disaggregated AUs
cd_dis_sf <- msfd %>%
  dplyr::filter(
    name_text %in% cd_dis_names
  ) %>%
  dplyr::transmute(
    ID = "Common Dolphin",
    Name = name_text,
    geometry = geometry
  )

# Use exactly the same CRS metadata as geoinfo
cd_dis_sf <- sf::st_transform(
  cd_dis_sf,
  sf::st_crs(geoinfo)
)

# Check that all expected AUs were found
missing_cd_dis <- setdiff(
  cd_dis_names,
  as.character(cd_dis_sf$Name)
)

if (length(missing_cd_dis) > 0) {
  stop(
    "Missing Common dolphin disaggregated AU polygon(s): ",
    paste(missing_cd_dis, collapse = " | ")
  )
}

# Simplify ONCE for map production
# Keep the original cd_dis_sf untouched
cd_dis_sf_map <- rmapshaper::ms_simplify(
  cd_dis_sf,
  keep = 0.01,
  keep_shapes = TRUE
)



### Load D1 for WGBYC and D1 for WKMAMBY
#WGBYC
load('data/WGBYCD1_for_WKMAMBY20260623_111342.RData')
D1byc <- WGBYCD1_for_WKMAMBY
unique(D1byc$year)

# WKMAMBY
load("data/WKMAMBYD1_withoutES9a8c_harbour_porpoise_corrected20260623_095707.RData")

D1mby <- WKMAMBYD1_withoutES9a8c %>% 
  mutate(year = as.integer(year)) %>%
  left_join(as.data.frame(geoinfo) %>%
              dplyr::select(ICESNAME, Area_27) %>%
              dplyr::rename(statRec = ICESNAME, areaCode = Area_27), by = "statRec") %>% 
  mutate(areaCode = paste0("27.", areaCode))
  
unique(D1mby$year)

### Load D2 for WGBYC and D2 for WKMAMBY

# WGBYC
load("data/WGBYCD2_for_WKMAMBY20260623_111520.RData")

D2byc <- WGBYCD2_for_WKMAMBY %>%
  mutate(monitoringProgramType = as.numeric(monitoringProgramType))
unique(D2byc$year)

# WKMAMBY
load("data/WKMAMBYD2_harbour_porpoise_corrected20260623_095707.RData")

D2mby <- WKMAMBYD2 %>% 
  mutate(year = as.integer(year),
         monitoringProgramType = as.numeric(monitoringProgramType)) %>%
  left_join(as.data.frame(geoinfo) %>%
              dplyr::select(ICESNAME, Area_27) %>%
              dplyr::rename(statRec = ICESNAME, areaCode = Area_27), by = "statRec") %>% 
  mutate(areaCode = paste0("27.", areaCode))
  
unique(D2mby$year)

### Load D3 for WGBYC and D3 for WKMAMBY
# WGBYC
load("data/WGBYCD3_for_WKMAMBY20260623_111931.RData")
D3byc <- WGBYCD3_for_WKMAMBY %>% 
  mutate(monitoringProgramType = as.numeric(monitoringProgramType))
unique(D3byc$year)
# WKMAMBY
load("data/WKMAMBYD3_harbour_porpoise_corrected20260623_095707.RData")
D3mby <- WKMAMBYD3 %>% 
  mutate(year = as.integer(year),
         monitoringProgramType = as.numeric(monitoringProgramType)) %>%
  left_join(as.data.frame(geoinfo) %>%
              dplyr::select(ICESNAME, Area_27) %>%
              dplyr::rename(statRec = ICESNAME, areaCode = Area_27), by = "statRec") %>% 
  mutate(areaCode = paste0("27.", areaCode))
  
unique(D3mby$year)


#### Now we need to merge WGBYC and WKMAMBY data
# Also add plain country names, monitoring method, time filter
D1 <- bind_rows(D1byc, D1mby) %>%
  left_join(ctrycodes, by = "country") %>%
  left_join(gearcodes, by = "metierL3") %>% 
  filter(statRec != -9 | is.na(statRec),
         metierL3 != "L3NK") %>% 
# Need to change GTN to GNS for PT
  mutate(metierL4 = case_when(country == "PT" & metierL4 == "GTN" ~ "GNS", TRUE ~ metierL4),
         daysAtSeaF = if_else(metierL4 %in% c("PTM", "PTB", "SPR") & !country %in% c("DK", "NL"),
                              daysAtSeaF / 2, daysAtSeaF))
# Divide fishing effort by 2 for every country except DK and NL : PTM, PTB, SPR 
  

D2_before_filters <- bind_rows(D2byc, D2mby) %>%
  left_join(ctrycodes, by = "country") %>%
  left_join(gearcodes, by = "metierL3") %>%
  left_join(monmeth, by = "monitoringMethod") %>%
  left_join(monprog, by = "monitoringProgramType") %>%
  filter(year %in% 2016:2024)

D2 <- D2_before_filters %>% 
  filter(!monitoringMethod %in% c("PO", "OTH"),
         statRec != -9 | is.na(statRec),
         metierL3 != "L3NK")

D3 <- bind_rows(D3byc, D3mby) %>%
  left_join(ctrycodes, by = "country") %>%
  left_join(gearcodes, by = "metierL3") %>%
  left_join(monmeth, by = "monitoringMethod") %>%
  left_join(monprog, by = "monitoringProgramType") %>%
  filter(year %in% 2016:2024,
         species %in% c("Phocoena phocoena", 
                        "Delphinus delphis",
                        "Halichoerus grypus",
                        "Phoca vitulina",
                        "Balaenoptera acutorostrata"),
         !monitoringMethod %in% c("PO", "OTH"),
         statRec != -9 | is.na(statRec),
         metierL3 != "L3NK")

# 2024 correction because Belgium, -9, and format definition of a data structure
# for WGBYC is only a dream
D3$individualsWithPingers[D3$individualsWithPingers == -9] <- NA
D3$incidentsWithPingers[D3$incidentsWithPingers == -9] <- NA
```

```{r Monitoring dataprep}
# Prepare data for figures (reported monitoring method)
# All
monmethall_before_filters <- D2_before_filters %>%
  group_by(monmethname, year) %>%
	summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>% 
	group_by(year) %>%
	mutate(tot = sum(das)) %>%
	ungroup() %>%
	mutate(reldas = das/tot*100) %>%
  mutate(monmethname = recode(monmethname,
                                   "SO"  = "At-sea observers",
                                   "EM"  = "Electronic monitoring",
                                   "LB"  = "Logbook",
                                   "OTH" = "Other",
                                   "PO"  = "Port observers",
                                   "VO"  = "Vessel crew observer"))
monmethall <- D2 %>%
  group_by(monmethname, year) %>%
	summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>% 
	group_by(year) %>%
	mutate(tot = sum(das)) %>%
	ungroup() %>%
	mutate(reldas = das/tot*100) %>%
  mutate(monmethname = recode(monmethname,
                                   "SO"  = "At-sea observers",
                                   "EM"  = "Electronic monitoring",
                                   "LB"  = "Logbook",
                                   "OTH" = "Other",
                                   "PO"  = "Port observers",
                                   "VO"  = "Vessel crew observer"))
# By country
monmethctry <- D2 %>%
  group_by(monmethname, ctryname, year) %>%
	summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
	group_by(year, ctryname) %>%
	mutate(tot = sum(das)) %>%
	ungroup() %>%
	mutate(reldas = das/tot*100) %>%
  mutate(monmethname = recode(monmethname,
                                   "SO"  = "At-sea observers",
                                   "EM"  = "Electronic monitoring",
                                   "LB"  = "Logbook",
                                   "OTH" = "Other",
                                   "PO"  = "Port observers",
                                   "VO"  = "Vessel crew observer"))

method_colors <- c(
  "At-sea observers"      = "#F8766D",
  "Electronic monitoring" = "#B79F00",
  "Logbook"               = "#00BA38",
  "Other"                 = "#00BFC4",
  "Port observers"        = "#619CFF",
  "Vessel crew observer"  = "#ED68ED"
)
```


## Overview of data and monitoring methods

```{r}
# Monitoring methods overview
# Annual totals
tot_before <- monmethall_before_filters %>%
  distinct(year, tot)

tot_torc <- monmethall %>%
  distinct(year, tot)

max_das <- max(c(tot_before$tot, tot_torc$tot), na.rm = TRUE)

# Main alluvial plot - before filters
plt_data_before_filters <- ggplot(
  monmethall_before_filters,
  aes(x = year,
    y = reldas,
    stratum = monmethname,
    alluvium = monmethname,
    fill = monmethname,
    label = monmethname)) +
  geom_flow(stat = "alluvium") +
  geom_stratum() +
  scale_fill_manual(values = method_colors) +
  scale_x_continuous(
    breaks = seq(2016, 2024, 1),
    limits = c(2015.5, 2024.5),
    expand = c(0, 0)) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(fill = "Monitoring method") +
  ylab("Proportion of annual DAS\nby monitoring method (%)")

# Main alluvial plot - ToR C
plt_data_torc <- ggplot(
  monmethall,
  aes(x = year,
    y = reldas,
    stratum = monmethname,
    alluvium = monmethname,
    fill = monmethname,
    label = monmethname)) +
  geom_flow(stat = "alluvium") +
  geom_stratum() +
  scale_fill_manual(values = method_colors, drop = FALSE) +
  scale_x_continuous(
    breaks = seq(2016, 2024, 1),
    limits = c(2015.5, 2024.5),
    expand = c(0, 0)) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(fill = "Monitoring method") +
  ylab("Proportion of annual DAS\nby monitoring method used in ToR C (%)")

# Small DAS barplot - before filters
bar_before <- ggplot(tot_before, aes(year, tot)) +
  geom_col(width = 0.4, col = "black") +
  scale_x_continuous(
    breaks = seq(2016, 2024, 1),
    limits = c(2015.5, 2024.5),
    expand = c(0, 0)) +
  scale_y_continuous(
    limits = c(0, max_das),
    expand = expansion(mult = c(0, 0.05))) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(5, 5, 0, 5)) +
  ylab("Total DAS\nmonitored") +
  annotate("text", x = -Inf, y = Inf, label = "A", 
           hjust = -0.5, vjust = 1.5, fontface = "bold", size = 6)

# Small DAS barplot - ToR C
bar_torc <- ggplot(tot_torc, aes(year, tot)) +
  geom_col(width = 0.4, col = "black") +
  scale_x_continuous(
    breaks = seq(2016, 2024, 1),
    limits = c(2015.5, 2024.5),
    expand = c(0, 0)) +
  scale_y_continuous(
    limits = c(0, max_das),
    expand = expansion(mult = c(0, 0.05))) +
  theme_bw() +
  theme(
    panel.grid = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(5, 5, 0, 5)
  ) +
  ylab("Total DAS\nmonitored") +
  annotate("text", x = -Inf, y = Inf, label = "B",
           hjust = -0.5, vjust = 1.5, fontface = "bold", size = 6)

# Stack barplots above alluvial plots
panel_before <- plot_grid(bar_before, plt_data_before_filters, ncol = 1,
                          align = "v", rel_heights = c(0.2, 1))

panel_torc <- plot_grid(bar_torc, plt_data_torc, ncol = 1, 
                        align = "v", rel_heights = c(0.2, 1))

# Extract legend from one plot
legend <- get_legend(
  plt_data_before_filters +
    theme( legend.position = "bottom", legend.box.margin = margin(0, 0, 0, 0)))

# Combine A and B panels
combined_plot <- plot_grid(panel_before, panel_torc, ncol = 2, align = "v")

# Add shared legend
final_plot <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.1))

print(final_plot)

taf.png("output/monitoring_method_overview.png")
print(final_plot)
dev.off()
```

```{r Monitoring method by country}
# Monitoring method by country 
plt <- ggplot(monmethctry, 
              aes(x = year, y = reldas,
                  stratum = monmethname,
                  alluvium = monmethname,
                  fill = monmethname,
                  label = monmethname)) +
	  scale_fill_manual(values = method_colors) +
	  geom_flow(stat = "alluvium") +
	  geom_stratum() +
	facet_wrap(~ctryname) +
	theme_bw() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_x_continuous(breaks = seq(2016,2024,1)) +
	labs(fill="Monitoring method") +
	ylab("Proportion of annual DAS by monitoring method (%)")

print(plt)

taf.png("output/monitoring_method_country.png")
print(plt)
dev.off()

```

```{r Monitoring program by country}
# Monitoring program by country 
monprogctry <- D2 %>% 
  group_by(monprog, ctryname, year) %>%
	summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
	group_by(year, ctryname) %>%
	mutate(tot = sum(das)) %>%
	ungroup() %>%
	mutate(reldas = das/tot*100)

plt <- ggplot(monprogctry, aes(x = year, y = reldas,
			    stratum = monprog,
			    alluvium = monprog,
			    fill = monprog,
			    label = monprog)) +
	geom_flow(stat = "alluvium") +
	geom_stratum() +
	facet_wrap(~ctryname) +
	theme_bw() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_x_continuous(breaks = seq(2016,2024,1)) +
  labs(fill = "Monitoring program") +
	ylab("Proportion of annual DAS by monitoring program (%)")

print(plt)

taf.png("output/monitoring_program_country.png")
print(plt)
dev.off()
```

```{r Incidents by monitoring program}
# Incidents by monitoring program and species class
incidents <- D3
incidents$nbinc <- apply(D3 %>% dplyr::select(incidentsWithPingers, incidentsWithoutPingers), 1, sum, na.rm = T)

incprog <- incidents %>%
  group_by(monmethname, year, species) %>%
	summarise(inc = sum(nbinc, na.rm = T), .groups = "drop") %>%
	ungroup()

plt <- ggplot(incprog, aes(x = as.factor(year), y = inc,
			    stratum = monmethname,
			    alluvium = monmethname,
			    fill = monmethname,
			    label = monmethname))+
	geom_flow(stat = "alluvium")+
	geom_stratum() +
	facet_wrap(~species, scale = "free") +
	theme_bw() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
	      legend.position = "inside",
	      legend.position.inside = c(.85,0.25)) +
  scale_x_discrete(breaks = seq(2016,2024,1)) +
	labs(fill = "Monitoring method") +
	ylab("Number of reported incidents") + xlab("")

print(plt)

taf.png("output/incidents_by_monprog_class.png")
print(plt)
dev.off()
```



## Prepare data for sample coverage by monitoring effort by gear by assessment unit




```{r Data by monitoring effort and AUs, echo = FALSE}
# Create correspondences table
species_info <- tibble::tibble(
  colname = c(
    "harbour_porpoise",
    "common_dolphin_aggregated",
    "common_dolphin_disaggregated",
    "minke_whale",
    "grey_seal_aggregated",
    "grey_seal_disaggregated",
    "harbour_seal_aggregated",
    "harbour_seal_disaggregated"
  ),
  layer = c(
    "harbour porpoise",
    "common dolphin",
    "common dolphin",
    "minke whale",
    "grey seal",
    "grey seal",
    "harbour seal",
    "harbour seal"
  ),
  aggregation = c(
    "single",
    "aggregated",
    "disaggregated",
    "single",
    "aggregated",
    "disaggregated",
    "aggregated",
    "disaggregated"
  ),
  AU_code = c(
    "AU_HP",
    "AU_CD",
    "AU_CD_dis",
    "AU_MK",
    "AU_GS",
    "AU_GS_dis",
    "AU_HS",
    "AU_HS_dis"
  ),
  species = c(
    "Harbour porpoise",
    "Common dolphin (aggregated)",
    "Common dolphin (disaggregated)",
    "Minke whale",
    "Grey seal (aggregated)",
    "Grey seal (disaggregated)",
    "Harbour seal (aggregated)",
    "Harbour seal (disaggregated)"
  )
)

species_info <- species_info %>%
  mutate(species_group = str_remove(species, " \\((aggregated|disaggregated)\\)$"),
         aggregation = case_when(
           str_detect(species, "\\(aggregated\\)$") ~ "Aggregated",
           str_detect(species, "\\(disaggregated\\)$") ~ "Disaggregated",
           TRUE ~ "Single"))

# Index containing the filenames of all plots
coverage_plot_index <- vector("list")
k <- 1

for (i in seq_len(nrow(species_info))) {
  
  sp <- species_info$colname[i]
  prefix <- species_info$colname[i]
  title_text <- species_info$species[i]
  species_group <- species_info$species_group[i]
  aggregation <- species_info$aggregation[i]
  
  monmethgear1 <- D2 %>%
    group_by(monmethname, .data[[sp]], L3name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = TRUE), .groups = "drop") %>%
    filter(!monmethname %in% c("Logbook", "Port observers"))
  
  monmethgear2 <- monmethgear1 %>%
    mutate(L3name = "All metier") %>%
    group_by(monmethname, .data[[sp]], L3name, year) %>%
    summarise(das = sum(das, na.rm = TRUE), .groups = "drop")
  
  monmethgear <- bind_rows(monmethgear1, monmethgear2) %>%
    mutate(monmethname = as.factor(monmethname))
  
  list_assessment_units <- na.omit(unique(monmethgear[[sp]]))

  
  for (au in list_assessment_units) {
    
    tmp <- monmethgear[monmethgear[[sp]] == au, ] %>%
      filter(!is.na(L3name))
    
    
    plt <- ggplot(tmp,
      aes(x = year, y = das, stratum = monmethname, alluvium = monmethname,
          fill = monmethname, label = monmethname)) +
      geom_flow(stat = "alluvium") +
      geom_stratum() +
      facet_wrap(~L3name, scale = "free") +
      theme_bw() +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
      scale_fill_manual(values = method_colors, limits = names(method_colors), drop = FALSE) +
      scale_x_continuous(breaks = seq(2016, 2024, 1)) +
      labs(fill = "Monitoring method") +
      ggtitle(paste(title_text, "/", au)) +
      ylab("Sum of annual DAS by monitoring method")
    
    filename <- paste0("output/monitoring_method_", prefix, "_", gsub(" ", "", au), ".png")
    
    ggsave(filename = filename,
      plot = plt,
      width = 12,
      height = 8,
      dpi = 120
    )
    
    # THIS is what creates the index used by the tab chunk
    coverage_plot_index[[k]] <- tibble(
      species_group = species_group,
      aggregation = aggregation,
      species = title_text,
      colname = sp,
      assessment_unit = as.character(au),
      filename = filename
    )
    
    k <- k + 1
  }
}

# Turn the list into a tibble
coverage_plot_index <- bind_rows(coverage_plot_index)
```


## Create the tabs outputs in html


```{r coverage_tabs, echo = FALSE, results = "asis"}

cat('
<style>
.coverage-plot img {
  width: 100%;
  height: auto;
  display: block;
  margin-top: 15px;
  margin-bottom: 25px;
}
</style>
')


# ============================================================
# LEVEL 1: SPECIES
# ============================================================

cat("\n## Monitoring effort by assessment unit {.tabset .tabset-pills}\n\n")


for (sp in unique(coverage_plot_index$species_group)) {
  
  dat_sp <- coverage_plot_index %>%
    filter(species_group == sp)
  
  
  cat("### ", sp, " {.tabset}\n\n", sep = "")
  
  
  # ==========================================================
  # LEVEL 2: AGGREGATED / DISAGGREGATED
  # ==========================================================
  
  for (agg in unique(dat_sp$aggregation)) {
    
    dat_agg <- dat_sp %>%
      filter(aggregation == agg)
    
    cat("#### ", agg, " {.tabset}\n\n", sep = "")
    
    # ========================================================
    # LEVEL 3: ASSESSMENT UNIT
    # ========================================================
    
    for (i in seq_len(nrow(dat_agg))) {
      
      x <- dat_agg[i, ]
      
      cat("##### ", x$assessment_unit, "\n\n", sep = "")
      
      
      cat('<div class="coverage-plot">', '<img src="', x$filename,
          '" alt="Monitoring effort plot for ', x$assessment_unit,
          '">', '</div>', "\n\n", sep = "")
      }
  }
}


# Close tabset
cat("\n## {-}\n\n")
```


## Maps of fishing effort, monitoring effort, and monitoring coverage

We need maps at both statistical rectangle and ICES divisions.
I should draw both the total and for instance the previous year.

Fishing over the years for each AU. Interesting for each division? Also check monitoring.

Add also at the ICES subdivisions levels

```{r Fishing effort monitoring effort and monitoring coverage: process data function, echo = FALSE}
# Function to process data
process_data <- function(d1, d2) {
  
  d_list = list()
  
  d1_l3 <- d1 %>% 
#    filter(year == analysis_year) %>%
#    group_by(!!sym(metierL3), areaCode) %>% # for WGBYC
    filter(statRec != -9) %>% 
    group_by(metierL3, statRec) %>%
    summarize(total_daysAtSeaF = sum(daysAtSeaF, na.rm = TRUE)) %>%
    dplyr::rename(ICESNAME = statRec) %>% 
    left_join(geoinfo, by = "ICESNAME")
  
  d2_l3 <- d2 %>% 
#    filter(year == analysis_year) %>%
#    group_by(!!sym(metierL3), areaCode) %>% # for WGBYC
    filter(statRec != -9) %>% 
    group_by(metierL3, statRec) %>%
    summarize(total_daysAtSeaOb = sum(daysAtSeaOb, na.rm = TRUE)) %>%
    dplyr::rename(ICESNAME = statRec) %>% 
    left_join(geoinfo, by = "ICESNAME")
  
  mc_l3 <- d1_l3 %>% 
    left_join(d2_l3) %>%
    mutate(monitoring_coverage = total_daysAtSeaOb / total_daysAtSeaF * 100,
           monitoring_coverage = replace_na(monitoring_coverage, 0))

  d1_l4 <- d1 %>% 
#    filter(year == analysis_year) %>%
#    group_by(!!sym(metierL4), areaCode) %>% # for WGBYC
      filter(statRec != -9) %>% 
      group_by(!!sym("metierL4"), statRec) %>%
      summarize(total_daysAtSeaF = sum(daysAtSeaF, na.rm = TRUE)) %>%
      dplyr::rename(ICESNAME = statRec) %>%
      left_join(geoinfo, by = "ICESNAME")
  
  d2_l4 <- d2 %>% 
#    filter(year == analysis_year) %>%
#    group_by(!!sym(metierL4), areaCode) %>% # for WGBYC
      filter(statRec != -9) %>% 
      group_by(!!sym("metierL4"), statRec) %>%
      summarize(total_daysAtSeaOb = sum(daysAtSeaOb, na.rm = TRUE)) %>%
      dplyr::rename(ICESNAME = statRec) %>% 
      left_join(geoinfo, by = "ICESNAME")
  
  mc_l4 <- d1_l4 %>% 
    left_join(d2_l4) %>%
    mutate(monitoring_coverage = total_daysAtSeaOb / total_daysAtSeaF * 100,
           monitoring_coverage = replace_na(monitoring_coverage, 0))
  
  d_list$metierL3 <- list(d1_l3, d2_l3, mc_l3)
  d_list$metierL4 <- list(d1_l4, d2_l4, mc_l4)
  
return(d_list)
}
```








```{r Fishing effort monitoring effort and monitoring coverage: launch function, echo = FALSE}
# ------------------------------------------------------------
# Data periods for maps
# ------------------------------------------------------------

map_data <- list(
  "2016-2024" = process_data(
    D1,
    D2
  ),

  "2024" = process_data(
    D1 %>% dplyr::filter(year == 2024),
    D2 %>% dplyr::filter(year == 2024)
  )
)
```







```{r Fishing effort monitoring effort and monitoring coverage: run, echo = FALSE}
# Rename metiers
rename_vector <- c(
  # L4 level
  "DRB" = "DRB - Boat dredge", "FPN" = "FPN - Stationary uncovered pound nets", "FPO" = "FPO - Pots and traps",
  "FWR" = "FWR - Fixed installations for fences and weirs", "FYK" = "FYK - Fyke nets", "GEL" = "GEL - Glass eel fishing",
  "GN" = "GN - Gill nets", "GNC" = "GNC - Encircling gillnets", "GND" = "GND - Driftnet", "GNS" = "GNS - Set gillnet",
  "GTN" = "GTN - Gillnets-trammel nets", "GTR" = "GTR - Trammel nets", "HMD" = "HMD - Mechanised/suction dredge",
  "LA" = "LA - Lampara nets", "LH" = "LH - Hand and pole lines", "LHM" = "LHM - Hand and pole lines",
  "LHP" = "LHP - Hand and pole lines (hand-operated)", "LLD" = "LLD - Drifting longlines", "LLS" = "LLS - Set longlines",
  "LTL" = "LTL - Trolling lines", "LX" = "LX - Hooks and lines", "MIS" = "MIS - Gear nei.", "OTB" = "OTB - Bottom otter trawl",
  "OTM" = "OTM - Midwater otter trawl", "OTT" = "OTT - Milti-rig otter trawl", "PS" = "PS - Purse seine",
  "PTB" = "PTB - Bottom pair trawl", "PTM" = "PTM - Pelagic pair trawl", "SB" = "SB - Beach Seine",
  "SBV" = "SBV - Beach and boat seine", "SDN" = "SDN - Anchored seine", "SPR" = "SPR - Pair seine",
  "SSC" = "SSC - Fly shooting seine", "TBB" = "TBB - Beam trawl",
  # L3 level
  "L3DR" = "DR - Dredges", "L3FIX" = "FIX - Traps", "L3GN" = "GN - Nets",
  "L3LH" = "LH - Rods and Lines", "L3LL" = "LL - Longlines", "L3PS" = "PS - Surrounding Nets",
  "L3PT" = "PT - Pelagic Trawls", "L3SX" = "SX - Seines", "L3TB" = "TB - Bottom Trawls"
)
```


```{r Load stuff, echo = FALSE}
# Load ICES areas, subareas and subdivisions
icesfile <- here("data/ICES_areas/ICES_Areas_20160601_cut_dense_3857.shp")
icesgeo <- st_read(icesfile)
icesgeo <- ms_simplify(icesgeo, keep = 0.01)

icesgeo_lab <- icesgeo %>%
  sf::st_transform(3035) %>%
  sf::st_point_on_surface() %>%
  sf::st_transform(st_crs(geoinfo)) %>%
  cbind(sf::st_coordinates(.))

icesgeo <- st_transform(icesgeo, crs = st_crs(geoinfo))

# Load and transform data
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")# %>%
  #st_make_valid() %>%
  #st_transform(st_crs(geoinfo))
```


```{r Fishing effort monitoring effort and monitoring coverage: prepare color palette, echo = FALSE}
## Plot the maps with fishing effort, monitoring effort, monitoring coverage by AU
cols <- brewer.pal(7, "Blues")
cols_2 <- brewer.pal(6, "Blues")

own_palette <- list(
  fishing_effort = c(
    "< 100" = cols[1],
    "100–1000" = cols[2],
    "1000–2000" = cols[3], 
    "2000–5000" = cols[4],
    "5000-10000" = cols[5],
    "10000-25000" = cols[6], 
    "≥ 25000" = cols[7]
    ),
  monitoring_effort = c(
    "< 5" = cols[1],
    "5–10" = cols[2],
    "10–20" = cols[3],
    "20–50" = cols[4],
    "50-100" = cols[5],
    "100-500" = cols[6],
    "≥ 500" = cols[7]
    ),
  monitoring_coverage = c(
    "< 0.25" = cols_2[1],
    "0.25-0.5" = cols_2[2],
    "0.5-1" = cols_2[3],
    "1–2" = cols_2[4],
    "2–5" = cols_2[5],
    "≥ 5" = cols_2[6]
    )
  )


```




```{r Test}
# ------------------------------------------------------------
# Load official MSFD regions/subregions from EEA
# ------------------------------------------------------------

msfd_url <- paste0(
  "https://marine.discomap.eea.europa.eu/",
  "arcgis/rest/services/Marine/Marine_regions_subregions/",
  "MapServer/0/query?",
  "where=1%3D1",
  "&outFields=*",
  "&returnGeometry=true",
  "&f=geojson"
)

msfd <- sf::st_read(
  msfd_url,
  quiet = TRUE
)

# Expected Common dolphin disaggregated AU names
cd_dis_names <- sort(
  unique(
    na.omit(
      as.character(
        geoinfo$common_dolphin_disaggregated
      )
    )
  )
)

# Select exactly those three polygons
cd_dis_sf <- msfd %>%
  dplyr::filter(
    name_text %in% cd_dis_names
  ) %>%
  dplyr::transmute(
    ID = "Common Dolphin",
    Name = name_text,
    geometry = geometry
  )

cd_dis_sf %>%
  sf::st_drop_geometry()

nrow(cd_dis_sf)

setdiff(
  cd_dis_names,
  cd_dis_sf$Name
)

cd_dis_sf_plot <- rmapshaper::ms_simplify(
  cd_dis_sf,
  keep = 0.01,
  keep_shapes = TRUE
)

plot(
  sf::st_geometry(cd_dis_sf_plot),
  border = "red",
  lwd = 2
)

plot(
  sf::st_geometry(
    geoinfo %>%
      dplyr::filter(!is.na(common_dolphin_disaggregated))
  ),
  add = TRUE,
  border = "grey40"
)

```








```{r Fishing effort monitoring effort and monitoring coverage: prepare data for maps function, echo = FALSE}
prepare_plot_species_maps <- function(list_d1_d2, colname, layer_name, species) {
  
  all_metiers <- c()
  all_metiers$metierL3 <- sort(unique(c(
    list_d1_d2$metierL3[[1]]$metierL3,
    list_d1_d2$metierL3[[2]]$metierL3,
    list_d1_d2$metierL3[[3]]$metierL3
    )))
  
  all_metiers$metierL4 <- sort(unique(c(
    list_d1_d2$metierL4[[1]]$metierL4,
    list_d1_d2$metierL4[[2]]$metierL4,
    list_d1_d2$metierL4[[3]]$metierL4
    )))
  
# Metier L3
  
  # Fishing effort
  list_d1_d2$metierL3[[1]] <- st_as_sf(list_d1_d2$metierL3[[1]]) %>%
    tidyr::drop_na(colname) %>% 
    mutate(
      total_daysAtSeaF_class = cut(
        total_daysAtSeaF,
        breaks = c(0, 100, 1000, 2000, 5000, 10000, 25000, Inf),
        labels = c("< 100", "100–1000", "1000–2000", "2000–5000", "5000-10000", "10000-25000", "≥ 25000"),
        include.lowest = TRUE),
      !!sym("metierL3") := recode(!!sym("metierL3"), !!!rename_vector))
  
  # Monitoring effort
  list_d1_d2$metierL3[[2]] <- st_as_sf(list_d1_d2$metierL3[[2]]) %>%
    tidyr::drop_na(colname) %>% 
    mutate(
      total_daysAtSeaOb_class = cut(
      total_daysAtSeaOb,
      breaks = c(0, 5, 10, 20, 50, 100, 500, Inf),
      labels = c("< 5", "5–10", "10–20", "20–50", "50-100", "100-500", "≥ 500"),
      include.lowest = TRUE),
      !!sym("metierL3") := recode(!!sym("metierL3"), !!!rename_vector))
  
  # Monitoring coverage
  list_d1_d2$metierL3[[3]] <- st_as_sf(list_d1_d2$metierL3[[3]]) %>%
    tidyr::drop_na(colname) %>% 
    mutate(monitoring_coverage_class = cut(
      monitoring_coverage,
      breaks = c(0, 0.25, 0.5, 1, 2, 5, Inf),
      labels = c("< 0.25", "0.25-0.5", "0.5-1", "1–2", "2–5", "≥ 5"),
      include.lowest = TRUE),
      !!sym("metierL3") := recode(!!sym("metierL3"), !!!rename_vector))
    
# Metier L4
  
  # Fishing effort
  list_d1_d2$metierL4[[1]] <- st_as_sf(list_d1_d2$metierL4[[1]]) %>%
    tidyr::drop_na(colname) %>% 
    mutate(
      total_daysAtSeaF_class = cut(
        total_daysAtSeaF,
        breaks = c(0, 100, 1000, 2000, 5000, 10000, 25000, Inf),
        labels = c("< 100", "100–1000", "1000–2000", "2000–5000", "5000-10000", "10000-25000", "≥ 25000"),
        include.lowest = TRUE),
      !!sym("metierL4") := recode(!!sym("metierL4"), !!!rename_vector))
  
  # Monitoring effort
  list_d1_d2$metierL4[[2]] <- st_as_sf(list_d1_d2$metierL4[[2]]) %>%
    tidyr::drop_na(colname) %>% 
    mutate(
      total_daysAtSeaOb_class = cut(
      total_daysAtSeaOb,
      breaks = c(0, 5, 10, 20, 50, 100, 500, Inf),
      labels = c("< 5", "5–10", "10–20", "20–50", "50-100", "100-500", "≥ 500"),
      include.lowest = TRUE),
      !!sym("metierL4") := recode(!!sym("metierL4"), !!!rename_vector))
  
  # Monitoring coverage
  list_d1_d2$metierL4[[3]] <- st_as_sf(list_d1_d2$metierL4[[3]]) %>%
    tidyr::drop_na(colname) %>% 
    mutate(monitoring_coverage_class = cut(
      monitoring_coverage,
      breaks = c(0, 0.25, 0.5, 1, 2, 5, Inf),
      labels = c("< 0.25", "0.25-0.5", "0.5-1", "1–2", "2–5", "≥ 5"),
      include.lowest = TRUE),
      !!sym("metierL4") := recode(!!sym("metierL4"), !!!rename_vector))
  
  
# ------------------------------------------------------------
# Select AU polygons corresponding to this AU definition
# ------------------------------------------------------------

# AU names expected for this particular definition
# e.g. grey_seal_aggregated vs grey_seal_disaggregated
au_names <- sort(
  unique(
    na.omit(
      as.character(geoinfo[[colname]])
    )
  )
)

# ------------------------------------------------------------
# Select the appropriate AU polygon source
# ------------------------------------------------------------

if (colname == "common_dolphin_disaggregated") {

  au_sf <- cd_dis_sf_map %>%
    dplyr::filter(Name %in% au_names)

} else {

  au_sf <- suppressMessages(
    sf::st_read(
      AUsfile,
      layer = layer_name,
      quiet = TRUE
    )
  )

  au_sf <- au_sf %>%
    dplyr::filter(Name %in% au_names)

  if (nrow(au_sf) > 0) {
    au_sf <- rmapshaper::ms_simplify(
      au_sf,
      keep = 0.01,
      keep_shapes = TRUE
    )
  }
}

# Check whether some expected AU polygons are missing
missing_au <- setdiff(
  au_names,
  as.character(au_sf$Name)
)

if (length(missing_au) > 0) {
  warning(
    "Missing AU polygon(s) for ", colname, ": ",
    paste(missing_au, collapse = " | ")
  )
}

# Bounding box
if (nrow(au_sf) > 0) {

  bbox <- sf::st_bbox(au_sf)

} else {

  bbox <- sf::st_bbox(
    list_d1_d2$metierL3[[1]]
  )
}

# Color selection
areas_totales <- au_names
areas_totales <- areas_totales[!is.na(areas_totales)]

colores <- setNames(
  grDevices::colorRampPalette(
    RColorBrewer::brewer.pal(8, "Set2")
  )(length(areas_totales)),
  areas_totales
)

  list(list_d1_d2 = list_d1_d2,
        au_sf = au_sf,
        bbox = bbox,
        colores = colores,
        all_metiers = all_metiers)
  }
```






```{r Fishing effort monitoring effort and monitoring coverage: function for mapping, echo = FALSE}
create_plot <- function(list_d1_d2,
                        au_sf,
                        bbox,
                        colores,
                        all_metiers,
                        species, 
                        colname, 
                        metier_lvl,
                        period) {
  
  # ------------------------------------------------------------
  # Helper: add AU polygons only when available
  # ------------------------------------------------------------
  
  add_au_layer <- function(p, au_sf, colores) {
    
    if (nrow(au_sf) > 0) {
      
      p +
        geom_sf(
          data = au_sf,
          aes(color = Name),
          size = 0.5,
          fill = NA
        ) +
        scale_color_manual(
          values = colores,
          breaks = names(colores),
          name = "AUs"
        )
      
    } else {
      
      p
    }
  }
  
  
  # ------------------------------------------------------------
  # Select métier level
  # ------------------------------------------------------------
  
  if (metier_lvl == "metierL3") {
    
    metiers_names <- all_metiers$metierL3
    list_d1_d2 <- list_d1_d2$metierL3
    
  } else {
    
    metiers_names <- all_metiers$metierL4
    list_d1_d2 <- list_d1_d2$metierL4
  }
  
  
  # Store only lightweight metadata
  plot_index <- vector(
    "list",
    length(metiers_names)
  )
  
  
  # ------------------------------------------------------------
  # Loop over métiers
  # ------------------------------------------------------------
  
  for (j in seq_along(metiers_names)) {
    
    metier <- metiers_names[j]
    
    
    # ------------------------------------------------------------
    # Subset data
    # ------------------------------------------------------------
    
    fishing_effort_data <- dplyr::filter(
      list_d1_d2[[1]],
      .data[[metier_lvl]] == rename_vector[[metier]]
    )
    
    monitoring_effort_data <- dplyr::filter(
      list_d1_d2[[2]],
      .data[[metier_lvl]] == rename_vector[[metier]]
    )
    
    monitoring_coverage_data <- dplyr::filter(
      list_d1_d2[[3]],
      .data[[metier_lvl]] == rename_vector[[metier]]
    )
    
    
    # ============================================================
    # Fishing effort
    # ============================================================
    
    plt_fishing_effort <- ggplot() +
      
      geom_sf(
        data = fishing_effort_data,
        aes(fill = total_daysAtSeaF_class),
        show.legend = TRUE
      ) +
      
      scale_fill_manual(
        values = own_palette$fishing_effort,
        limits = names(own_palette$fishing_effort),
        drop = FALSE,
        na.value = "white",
        name = "Fishing effort (days at sea)"
      ) +
      
      geom_sf(
        data = world,
        fill = "grey90",
        color = "black",
        size = 0.1
      ) +
      
      geom_sf(
        data = icesgeo,
        fill = NA,
        colour = "black",
        size = 0.5
      ) +
      
      geom_text_repel(
        data = icesgeo_lab,
        aes(
          x = X,
          y = Y,
          label = Area_27
        ),
        size = 2,
        force = 0.01
      ) +
      
      theme_bw() +
      
      ggtitle(
        paste0(
          species,
          " / Fishing effort"
          ),
        subtitle = paste0(
          rename_vector[[metier]],
          " / ",
          period
          )
        ) +
      
      xlab("Longitude") +
      ylab("Latitude")
    
    
    # Add correct AU polygons
    plt_fishing_effort <- add_au_layer(
      plt_fishing_effort,
      au_sf,
      colores
    )

    plt_fishing_effort <- plt_fishing_effort +
      coord_sf(
        xlim = as.numeric(
          bbox[c("xmin", "xmax")]
          ),
        ylim = as.numeric(
          bbox[c("ymin", "ymax")]
          ),
        crs = sf::st_crs(4326),
        default_crs = sf::st_crs(4326),
        expand = FALSE
        )
    
    # ============================================================
    # Monitoring effort
    # ============================================================
    
    plt_monitoring_effort <- ggplot() +
      
      geom_sf(
        data = monitoring_effort_data,
        aes(fill = total_daysAtSeaOb_class),
        show.legend = TRUE
      ) +
      
      scale_fill_manual(
        values = own_palette$monitoring_effort,
        limits = names(own_palette$monitoring_effort),
        drop = FALSE,
        na.value = "white",
        name = "Monitoring effort (days at sea)"
      ) +
      
      geom_sf(
        data = world,
        fill = "grey90",
        color = "black",
        size = 0.1
      ) +
      
      geom_sf(
        data = icesgeo,
        fill = NA,
        colour = "black",
        size = 0.5
      ) +
      
      geom_text_repel(
        data = icesgeo_lab,
        aes(
          x = X,
          y = Y,
          label = Area_27
        ),
        size = 2,
        force = 0.01
      ) +

      theme_bw() +
      
      ggtitle(
        paste0(
          species,
          " / Monitoring effort"
          ),
        subtitle = paste0(
          rename_vector[[metier]],
          " / ",
          period
          )
        ) +
      
      xlab("Longitude") +
      ylab("Latitude")
    
    
    # Add correct AU polygons
    plt_monitoring_effort <- add_au_layer(
      plt_monitoring_effort,
      au_sf,
      colores
    )
    
    plt_monitoring_effort <- plt_monitoring_effort +
      coord_sf(
        xlim = as.numeric(
          bbox[c("xmin", "xmax")]
          ),
        ylim = as.numeric(
          bbox[c("ymin", "ymax")]
          ),
        crs = sf::st_crs(4326),
        default_crs = sf::st_crs(4326),
        expand = FALSE
        )
    
    # ============================================================
    # Monitoring coverage
    # ============================================================
    
    plt_monitoring_coverage <- ggplot() +
      
      geom_sf(
        data = monitoring_coverage_data,
        aes(fill = monitoring_coverage_class),
        show.legend = TRUE
      ) +
      
      scale_fill_manual(
        values = own_palette$monitoring_coverage,
        limits = names(own_palette$monitoring_coverage),
        drop = FALSE,
        na.value = "white",
        name = "Monitoring coverage (%)"
      ) +
      
      geom_sf(
        data = world,
        fill = "grey90",
        color = "black",
        size = 0.1
      ) +
      
      geom_sf(
        data = icesgeo,
        fill = NA,
        color = "black",
        size = 0.5
      ) +
      
      geom_text_repel(
        data = icesgeo_lab,
        aes(
          x = X,
          y = Y,
          label = Area_27
        ),
        size = 2,
        force = 0.01
      ) +
      
      theme_bw() +
      
      ggtitle(
        paste0(
          species,
          " / Monitoring coverage"
          ),
        subtitle = paste0(
          rename_vector[[metier]],
          " / ",
          period
          )
        ) +
      
      xlab("Longitude") +
      ylab("Latitude")
    
    
    # Add correct AU polygons
    plt_monitoring_coverage <- add_au_layer(
      plt_monitoring_coverage,
      au_sf,
      colores
    )
    
    plt_monitoring_coverage <- plt_monitoring_coverage +
      coord_sf(
        xlim = as.numeric(
          bbox[c("xmin", "xmax")]
          ),
        ylim = as.numeric(
          bbox[c("ymin", "ymax")]
          ),
        crs = sf::st_crs(4326),
        default_crs = sf::st_crs(4326),
        expand = FALSE
        )
    
    # ------------------------------------------------------------
    # Filenames
    # ------------------------------------------------------------
    
    period_filename <- dplyr::recode(
      period,
      "2016-2024" = "all",
      "2024" = "2024"
      )
    
    fishing_file <- paste0(
      "output/map_fishing_effort_",
      colname, "_",
      metier_lvl, "_",
      metier, "_",
      period_filename,
      ".png"
    )
    
    monitoring_file <- paste0(
      "output/map_monitoring_effort_",
      colname, "_",
      metier_lvl, "_",
      metier, "_",
      period_filename,
      ".png"
    )
    
    coverage_file <- paste0(
      "output/map_monitoring_coverage_",
      colname, "_",
      metier_lvl, "_",
      metier, "_",
      period_filename,
      ".png"
    )
    
    
    # ------------------------------------------------------------
    # Save maps
    # ------------------------------------------------------------
    
    ggsave(
      filename = fishing_file,
      plot = plt_fishing_effort,
      width = 12,
      height = 8,
      dpi = 120
    )
    
    ggsave(
      filename = monitoring_file,
      plot = plt_monitoring_effort,
      width = 12,
      height = 8,
      dpi = 120
    )
    
    ggsave(
      filename = coverage_file,
      plot = plt_monitoring_coverage,
      width = 12,
      height = 8,
      dpi = 120
    )
    
    
    # ------------------------------------------------------------
    # Store metadata + filenames only
    # ------------------------------------------------------------
    
    plot_index[[j]] <- tibble(
      species = species,
      colname = colname,
      period = period,
      metier_level = metier_lvl,
      metier = metier,
      metier_name = rename_vector[[metier]],
      fishing_effort = fishing_file,
      monitoring_effort = monitoring_file,
      monitoring_coverage = coverage_file
    )
  }
  
  
  # ------------------------------------------------------------
  # Return lightweight index
  # ------------------------------------------------------------
  
  dplyr::bind_rows(plot_index)
}
```




```{r Launch plot generation, quiet = TRUE, echo = FALSE}
map_index <- purrr::pmap(
  species_info,
  function(colname,
           layer,
           AU_code,
           species,
           species_group,
           aggregation) {

    period_results <- lapply(
      names(map_data),
      function(period) {

        # Data for this period
        period_data <- map_data[[period]]

        # Prepare spatial data for this species / AU definition
        prep <- prepare_plot_species_maps(
          list_d1_d2 = period_data,
          colname = colname,
          layer_name = layer,
          species = species
        )

        # ------------------------------------------------------------
        # Métier L3
        # ------------------------------------------------------------

        index_L3 <- create_plot(
          list_d1_d2 = prep$list_d1_d2,
          au_sf = prep$au_sf,
          bbox = prep$bbox,
          colores = prep$colores,
          all_metiers = prep$all_metiers,
          species = species,
          colname = colname,
          metier_lvl = "metierL3",
          period = period
        )

        # ------------------------------------------------------------
        # Métier L4
        # ------------------------------------------------------------

        index_L4 <- create_plot(
          list_d1_d2 = prep$list_d1_d2,
          au_sf = prep$au_sf,
          bbox = prep$bbox,
          colores = prep$colores,
          all_metiers = prep$all_metiers,
          species = species,
          colname = colname,
          metier_lvl = "metierL4",
          period = period
        )

        dplyr::bind_rows(
          index_L3,
          index_L4
        )
      }
    )

    dplyr::bind_rows(period_results) %>%
      dplyr::mutate(
        species_group = species_group,
        aggregation = aggregation,
        AU_code = AU_code,
        layer = layer,
        .before = 1
      )
  }
) %>%
  purrr::list_rbind()
```




```{r map_tabs, echo = FALSE, results = "asis"}

# ------------------------------------------------------------
# Options
# ------------------------------------------------------------

cat('
<style>
.map-row {
  display: grid;
  grid-template-columns: 1fr;
  gap: 16px;
  width: 100%;
  margin-top: 15px;
  margin-bottom: 25px;
}

.map-row img {
  width: 100%;
  height: auto;
  display: block;
}
</style>
')


# ============================================================
# LEVEL 1: SPECIES
# ============================================================

cat("\n# Maps {.tabset .tabset-pills}\n\n")

for (sp in unique(map_index$species_group)) {

  dat_sp <- map_index %>%
    dplyr::filter(species_group == sp)

  cat(
    "## ", sp, " {.tabset}\n\n",
    sep = ""
  )


  # ==========================================================
  # LEVEL 2: AGGREGATED / DISAGGREGATED
  # ==========================================================

  for (agg in unique(dat_sp$aggregation)) {

    dat_agg <- dat_sp %>%
      dplyr::filter(aggregation == agg)

    cat(
      "### ", agg, " {.tabset}\n\n",
      sep = ""
    )


    # ========================================================
    # LEVEL 3: MÉTIER LEVEL
    # ========================================================

    for (lvl in unique(dat_agg$metier_level)) {

      dat_lvl <- dat_agg %>%
        dplyr::filter(metier_level == lvl)

      lvl_label <- dplyr::recode(
        lvl,
        "metierL3" = "L3",
        "metierL4" = "L4"
      )

      cat(
        "#### Métier ", lvl_label, " {.tabset}\n\n",
        sep = ""
      )


      # ======================================================
      # LEVEL 4: INDIVIDUAL MÉTIER
      # ======================================================

      for (met in unique(dat_lvl$metier)) {

        dat_met <- dat_lvl %>%
          dplyr::filter(metier == met)

        met_name <- unique(dat_met$metier_name)[1]

        cat(
          "##### ", met_name, " {.tabset}\n\n",
          sep = ""
        )


        # ====================================================
        # LEVEL 5: PERIOD
        # ====================================================

        for (per in c("2016-2024", "2024")) {

          x <- dat_met %>%
            dplyr::filter(period == per)

          if (nrow(x) == 0) {
            next
          }

          x <- x[1, ]

          cat(
            "###### ", per, "\n\n",
            sep = ""
          )

          cat(
            '<div class="map-row">',

            '<div>',
            '<img src="', x$fishing_effort,
            '" alt="Fishing effort map - ', per, '">',
            '</div>',

            '<div>',
            '<img src="', x$monitoring_effort,
            '" alt="Monitoring effort map - ', per, '">',
            '</div>',

            '<div>',
            '<img src="', x$monitoring_coverage,
            '" alt="Monitoring coverage map - ', per, '">',
            '</div>',

            '</div>',

            "\n\n",
            sep = ""
          )
        }
      }
    }
  }
}

# Close outer tabset
cat("\n# {-}\n\n")
