---
title: "Tor XX / QC WGBYC 2026"
subtittle: "subtitle" 
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: Laurent Dubroca
bibliography: /Users/moi/zotero/exportdb/MyLibrary.bib 
number-sections: true
toc: true
toc-depth: 3
toc-expand: 3
format:
  html:
    html-math-method: katex
    css: styles.css
    code-fold: true
    page-layout: full
execute:
    warning: false
---


```{r compile, include=FALSE,eval=FALSE}
#  bookdown::word_document2: default
rmarkdown::render("output_qc.Rmd",clean=T)#,output_dir="./output",intermediates_dir="../../../tmp",knit_root_dir="../../../tmp")
unlink("output_qc_cache",recursive=T)
unlink("output_qc_files",recursive=T)
  #officedown::rdocx_document: default
#bibliography: '/Users/moi/Zotero/exportdb/My Library.bib'
#csl: './ices.csl'
```

# Rationale

This report presents a quality check framework developped for the WGBYC
database. 

WGBYC data are accessed using the ICES API in order to minimize the
data manipultion time and to avoid the error associated with data formatting and
reading.

## Data and reproducibility

Data used in this report were accessed following the TAF guideline.

```{r renderstuff,eval=F,include=F}
#| cache: true
#| echo: false
#|
#describe dir contains in the readme.md
if(F){
library(here)
here::i_am("project.Rproj")
source(here("analyses", "printdir.R"))
printdir(here("data"), here("data", "README.md"))
printdir(here("outputs"), here("outputs", "README.md"))
quarto::quarto_render("report_data_read.qmd")
}
```


```{r libraryandco,eval=T,include=T}
#|
library(readxl)
library(openxlsx)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(arrow)
library(data.table)
library(tictoc)
library(knitr)
library(kableExtra)
library(DT)
library(sf)
library(httr)
library(jsonlite)
require(icesConnect)
library(officer)
library(flextable)
library(ggalluvial)
library(httr)
library(jsonlite)
library(grid)
library(glue)
library(rnaturalearth)
library(rnaturalearthdata)
library(viridis)
library(ggforce)
library(stringr)
library(gridExtra)
library(cowplot)

here::i_am("project.Rproj")
```

# Parameters

Some fixed parameters if needed here.

```{r parameters,eval=T,include=T}
#define the most recent year of the data provided (aka the present year minus one as a practical guess)
currentyear <- as.numeric(substr(Sys.Date(), 1, 4)) - 1
print(currentyear)
```

# Data

WGBYC wunderbar API use + ICES vocab request to ensure reproducibility (just
in case if, for example, some data are submitted again).

```{r dataco,eval=T,include=F}
#read and load the data
D1 <- fread(here("data", "D1_wgbyc.csv"))
D2 <- fread(here("data", "D2_wgbyc.csv"))
D3 <- fread(here("data", "D3_wgbyc.csv"))
 #read dowloaded data
    listvocab <- c(
        "ISO_3166",
        "GearType",
        "GearGroup",
        "BYC_MonProgr",
        "BYC_SamplingProtocol",
        "BYC_MonMethod",
        "BYC_Mon812Type"
    )
    f1 <- function(vocabname) {
        tmp <- fread(here("data", "vocab", paste0(vocabname, ".csv")))
        assign(vocabname, tmp, pos = ".GlobalEnv")
    }
    lapply(listvocab, f1)
#add human readable label with some basic test in case ICES vocab being not consistent
#D1
tmp <- D1 %>%
    left_join(
        ISO_3166 %>% transmute(country = Key, country_name = Description),
        by = c("country")
    ) %>%
    left_join(
        GearGroup %>% transmute(metierL3 = Key, metierL3_name = Description),
        by = c("metierL3")
    )
if (nrow(tmp) != nrow(D1)) {
    stop("pb D1 explo1")
} else {
    D1 <- tmp
    rm(tmp)
}
#D2
tmp <- D2 %>%
    left_join(
        ISO_3166 %>% transmute(country = Key, country_name = Description),
        by = c("country")
    ) %>%
    left_join(
        GearGroup %>% transmute(metierL3 = Key, metierL3_name = Description),
        by = c("metierL3")
    ) %>%
    left_join(
        BYC_MonMethod %>%
            transmute(monitoringMethod = Key, monitoringMethod_name = Description),
        by = c("monitoringMethod")
    ) %>%
    left_join(
        BYC_MonProgr %>%
            transmute(
                monitoringProgramType = Key,
                monitoringProgramType_name = Description
            ),
        by = c("monitoringProgramType")
    )
if (nrow(tmp) != nrow(D2)) {
    stop("pb D2 explo1")
} else {
    D2 <- tmp
    rm(tmp)
}
#D3
tmp <- D3 %>%
    left_join(
        ISO_3166 %>% transmute(country = Key, country_name = Description),
        by = c("country")
    ) %>%
    left_join(
        GearGroup %>% transmute(metierL3 = Key, metierL3_name = Description),
        by = c("metierL3")
    ) %>%
    left_join(
        BYC_MonMethod %>%
            transmute(monitoringMethod = Key, monitoringMethod_name = Description),
        by = c("monitoringMethod")
    ) %>%
    left_join(
        BYC_MonProgr %>%
            transmute(
                monitoringProgramType = Key,
                monitoringProgramType_name = Description
            ),
        by = c("monitoringProgramType")
    )
if (nrow(tmp) != nrow(D3)) {
    stop("pb D3 explo1")
} else {
    D3 <- tmp
    rm(tmp)
}
#assemble regional entities

```

# Merge regional entities to country level

```{r}
D1 <- D1 %>%
    mutate(country_name = ifelse(country_name == "Azores", "Portugal", country_name)) %>%
    mutate(country_name = ifelse( country_name == "Northern Ireland", "United Kingdom", country_name))
D2 <- D2 %>%
    mutate(country_name = ifelse(country_name == "Azores", "Portugal", country_name)) %>%
    mutate(country_name = ifelse( country_name == "Northern Ireland", "United Kingdom", country_name))
D3 <- D3 %>%
    mutate(country_name = ifelse(country_name == "Azores", "Portugal", country_name)) %>%
    mutate(country_name = ifelse( country_name == "Northern Ireland", "United Kingdom", country_name))
```


```{r generaltable,eval=T}
#{{{
greentableD1<-D1%>%
	group_by(ctryname,year)%>%
	summarise(n=n())%>%
	mutate(year=paste0("D1_",year))%>%
	tidyr::pivot_wider(values_from=n,names_from=year,names_sort=T)
greentableD2<-D2%>%
	group_by(ctryname,year)%>%
	summarise(n=n())%>%
	mutate(year=paste0("D2_",year))%>%
	tidyr::pivot_wider(values_from=n,names_from=year,names_sort=T)
greentableD3<-D3%>%
	group_by(ctryname,year)%>%
	summarise(n=n())%>%
	mutate(year=paste0("D3_",year))%>%
	tidyr::pivot_wider(values_from=n,names_from=year,names_sort=TRUE)
greentable<-full_join(full_join(greentableD1,greentableD2),greentableD3)%>%
	ungroup()%>%data.frame()

#build up a submission table
subtableD1<-D1%>%
	group_by(ctryname,year)%>%
	summarise(n=paste0(sort(unique(submissionYear)),collapse=","))%>%
	mutate(year=paste0("D1_",year))%>%
	tidyr::pivot_wider(values_from=n,names_from=year,names_sort=T)
subtableD2<-D2%>%
	group_by(ctryname,year)%>%
	summarise(n=paste0(sort(unique(submissionYear)),collapse=","))%>%
	mutate(year=paste0("D2_",year))%>%
	tidyr::pivot_wider(values_from=n,names_from=year,names_sort=T)
subtableD3<-D3%>%
	group_by(ctryname,year)%>%
	summarise(n=paste0(sort(unique(submissionYear)),collapse=","))%>%
	mutate(year=paste0("D3_",year))%>%
	tidyr::pivot_wider(values_from=n,names_from=year,names_sort=T)
subtable<-full_join(full_join(subtableD1,subtableD2),subtableD3)%>%
	ungroup()%>%data.frame()




nametab<-names(greentable)[-1]
nametab<-paste0(nametab,"='",substr(nametab,4,7),"'",collapse=",")
nametab<-eval(parse(text=paste0("list(",nametab,")")))
nametab<-c(list(ctryname="Year of data"),nametab)
bgcolormat<-ifelse(greentable[,-1]<0.00001,"white","darkgreen")
charcolormat<-matrix("white",nrow=nrow(subtable),ncol=ncol(subtable)-1)
charcolormat["2025"==subtable[,-1]]<-"yellow"
#charcolormat["2021,2023"==subtable[,-1]]<-"yellow"
big_border<-fp_border(color="black", width = 2)
#remove number
greentable[is.na(greentable)]<-0


#param to fix column formatting
nbcolyear<-(ncol(subtable)-1)/3
posbigline<-c(1,nbcolyear+1,1+nbcolyear*2)


ftgreentable<-flextable(greentable)%>%
	bg(j=2:ncol(greentable),bg=bgcolormat)%>%
	color(j=2:ncol(greentable),color=charcolormat) %>%
	set_header_labels(values=nametab) %>%
	add_header_row(values=c("","Fishing Effort (D1 table)",
				"Monitoring Effort (D2 table)",
				"Bycatch Events (D3 table)"),colwidths=c(1,nbcolyear,nbcolyear,nbcolyear))%>%
	theme_box()%>%
	align(align="center",part="header")%>%
	align(align="center",part="body")%>%
	vline(j=posbigline,border=big_border)%>%
	fontsize(part="all",size=6)%>%
	autofit(add_w=0,add_h=0)

pltD4<-flextable(D4)%>%
	theme_box()%>%
	align(align="center",part="header")%>%
	align(align="center",part="body")%>%
	fontsize(part="all",size=6)%>%
	autofit(add_w=0,add_h=0)

#some stat for the text
nbcountryD1<-D4%>%filter(submissionYear==2025,grepl("2024",yearsInFishingEffort))%>%nrow
nbcountryD2<-D4%>%filter(submissionYear==2025,yearsInByCatchEvent==2024)%>%nrow
yeardatacall<-unique(D4$submissionYear)


#save as docx object
#save_as_docx(ftgreentable,path='./output/greentable.docx',
#	     pr_section=prop_section(
#				page_size=page_size(orient="landscape"),
#				page_margins=page_mar(bottom=0.1,top=.1,right=.1,left=.1,header=0,footer=0,gutter=0)
#				)
#		)
#}}}
```

# Table of responses by country – 2018 – 2024


ICES/WGBYC requested data from `r print(nbcountryD1)` countries through the `r yeardatacall` data call. 
Bulgaria submitted updated data, providing data from 2019 to 2024. 
A data submission was considered achieved if 
at least one value of effort was reported in the fishing and monitoring effort tables. 
For the bycatch events, only the presence of data was considered, 
as zero values (e.g. absence of bycatch events) is not clearly defined in the datacall. 
The submission status and date are summarized in the table 
`r figcap("greentable",display="num")`^[https://github.com/ices-eg/wg_WGBYC]. 


```{r pltgreentable,fig.height=8,fig.width=7}
print(pltD4)
```
`r figcap("greentable",caption="Summary of the data submitted in 2025 in the WGBYC database.")`

# Observation methods evolution

The evolution of the monitoring methods reported in the WGBYC data call is
presented in the figure 
`r figcap("pltt33",display="num")`. This figure is a variation of the figure 3.3
of the TOR a.

```{r newt33}
#{{{
t33<-D2%>%
	group_by(year,monmethname)%>%
	summarise(das=sum(daysAtSeaOb,na.rm=T))%>%
	ungroup()%>%
	group_by(year)%>%
	mutate(tot=sum(das))%>%
	ungroup()%>%
	mutate(reldas=das/tot*100)%>%
	mutate(monmethname=as.factor(monmethname))
pltt33<-ggplot(t33,aes(x=year,
	       y=reldas,
	       stratum=monmethname,
	       alluvium=monmethname,
	       fill=monmethname,
	       label=monmethname))+
	geom_flow(stat="alluvium")+
	geom_stratum()+
	#geom_text(stat="stratum")+
	theme_bw()+
	ylab("Proportion of the annual days at sea by monitoring methods (%)")+
	labs(fill="Monitoring method")
print(pltt33)
#}}}
```

`r figcap("pltt33",caption="Annual proportion of monitoring methods in days at sea from
2017 to 2023.")`

# Stability checks

In this section, the temporal evolution of the fishing and observation effort
is analysed using basic xy plot. The rationale behind these exploratory analyses
is the hypotheses that observation program of bycatch should be linked in term
of effort to the fishing effort. Statistical summaries were computed by year,
ecoregion and countries to provided an overall view of the data.
Days at sea is the main effort analysed in this section.

# Fishing and observation effort by ecoregion

```{r exploeff}
#{{{
ef1<-D1%>%
	group_by(year,L3name,ecoregion)%>%
	summarise(das=sum(daysAtSeaF,na.rm=T))%>%
	ungroup()
ef1bis<-D1%>%
	group_by(year,L3name,ecoregion,month)%>%
	summarise(das=sum(daysAtSeaF,na.rm=T))%>%
	ungroup()
ef2<-D2%>%
	group_by(year,L3name,ecoregion,monmethname)%>%
	summarise(dasobs=sum(daysAtSeaOb,na.rm=T))%>%
	ungroup()
ef2bis<-D2%>%
	group_by(year,L3name,ecoregion)%>%
	summarise(dasobs=sum(daysAtSeaOb,na.rm=T))%>%
	ungroup()
ef12<-full_join(ef1,ef2)
#}}}
```

```{r exploeffallcountry}
#{{{
idyear<-2023
plte1<-ggplot(ef1,aes(x=year-2000,y=das))+
	geom_point()+
	facet_grid(fctshoreco(ecoregion)~sub(" ","\n",L3name),scale="free")+
	scale_y_log10()+
	geom_smooth(method="lm")+
	theme_bw()+
	theme(strip.text.y.right= element_text(angle = 0,size=8))+
	theme(strip.text.x.top= element_text(size=8))+
	ylab("Days at sea (log 10 scale)")+xlab("year")
plte1bis1<-ggplot(
		  ef1bis%>%group_by(year,month=as.factor(month),ecoregion)%>%summarise(das=sum(das)))+
	aes(x=month,y=das,fill=month)+
	geom_bar(stat="identity")+
	scale_y_log10()+
	facet_grid(fctshoreco(ecoregion)~year,scale="free")+
	theme_bw()+
	theme(strip.text.y.right= element_text(angle = 0,size=8))+
	ylab("Days at sea (log 10 scale)")+xlab("Month")

plte1bis2<-ggplot(ef1bis)+aes(x=month,y=das,color=year,group=year)+
	geom_line()+
	#geom_point(data= ef1bis%>%filter(is.na(month))%>%mutate(month=1),
	#	   aes(x=month,y=das,color=year,group=year,shape=20))%>%
	facet_grid(fctshoreco(ecoregion)~sub(" ","\n",L3name),scale="free")+
	scale_y_log10()+
	theme_bw()+
	theme(strip.text.y.right= element_text(angle = 0,size=8))+
	theme(strip.text.x.top= element_text(size=8))+
	ylab("Days at sea (log 10 scale)")+xlab("year")


plte2<-ggplot(ef2,aes(x=year-2000,y=dasobs,color=monmethname))+
	facet_grid(fctshoreco(ecoregion)~sub(" ","\n",L3name),scale="free")+
	scale_y_log10()+
	geom_smooth(method="lm")+
	theme_bw()+
	theme(strip.text.y.right= element_text(angle = 0,size=8))+
	theme(strip.text.x.top= element_text(size=8))+
	theme(legend.position="bottom")+
	ylab("Days at sea (log 10 scale)")+xlab("year")
plte12<-ggplot(ef12%>%filter(is.finite(das)&is.finite(dasobs)),
	       aes(x=das,y=dasobs,color=(year)))+
	geom_point()+
	facet_grid(fctshoreco(ecoregion)~sub(" ","\n",L3name),scale="free")+
	scale_y_log10()+
	scale_x_log10()+
	geom_smooth(method="lm")+
	geom_point(data=ef12%>%filter(is.finite(das)&is.finite(dasobs),year==idyear),color="red",shape="+",size=2)+
	theme_bw()+
	theme(strip.text.y.right= element_text(angle = 0,size=8))+
	theme(strip.text.x.top= element_text(size=8))+
	ylab("Observation effort (days at sea, lo10 scale)")+
	xlab("Fishing effort (days at sea, log10 scale)")
#}}}
```

## Fishing effort 

The figure `r figcap("plte1all",display="num")` presents the temporal evolution 
of the fishing effort by
gears (metier level 3) and ecoregion.

```{r plte1all,fig.height=8,fig.width=7}
print(plte1)
```
*`r figcap("plte1all",caption="Temporal evolution of the fishing effort by ecoregion and metier level 3 from 2017 to 2023. In blue, linear model highlight the temporal trends over the 5 years.")`*

```{r plte1bisi1,fig.height=8,fig.width=7}
print(plte1bis1)
```
*`r figcap("plte1all",caption="Monthly total of the fishing effort by ecoregion and year from 2017 to 2023.")`*

```{r plte1bis,fig.height=8,fig.width=7}
print(plte1bis2)
```
*`r figcap("plte1bis",caption="Month fishing effort by ecoregion and metier level 3 from 2017 to 2023. In blue, linear model highlight the temporal trends over the 5 years.")`*

## Observation effort 

The figure `r figcap("plte2all",display="num")` presents the temporal evolution 
of the observation effort in days at sea by
gears (metier level 3) and ecoregion, and monitoring programs.

```{r plte2all,fig.height=8,fig.width=7}
print(plte2)
```
*`r figcap("plte2all",caption="Temporal evolution of the fishing effort by ecoregion and metier level 3 from 2017 to 2023 by monitoring program. In blue, linear model highlight the temporal trends over the 5 years.")`*

## Fishing versus observation effort 

The figure `r figcap("plte12all",display="num")` shows the total effort of
fishing versus the total effort of observation by year for each ecoregion and
gears.

```{r plte12all,fig.height=8,fig.width=7}
print(plte12)
```
*`r figcap("plte12all",caption="Fishing effort (x-axis) versus observation effort (y-axis) total by years, ecoregion and gear.  In blue, linear model highlight the temporal trends over the 5 years.The red cross highlights the last year provided by the data call.")`*

# By catch events

Some test on D3 table in relation to D2 and D1.

```{r bycatchstuff}
ef1<-D1%>%
	group_by(year,L3name,ecoregion)%>%
	summarise(das=sum(daysAtSeaF,na.rm=T))%>%
	ungroup()
ef2<-D2%>%
	group_by(year,L3name,ecoregion,monmethname)%>%
	summarise(das=sum(daysAtSeaOb,na.rm=T))%>%
	ungroup()
b3<-D3%>%
	mutate(n1=ifelse(is.na(incidentsWithPingers),0,1),
		n2=ifelse(is.na(incidentsWithPingers),0,1))%>%
	group_by(species,year,L3name,ecoregion)%>%
	summarise(nb=n1+n2)%>%
	ungroup() 

b3%>%tidyr::pivot_wider(values_from=nb,names_from=species)

marmamspp<-D3%>%filter(cetaceanFlag=="Y")%>%select(species)%>%distinct()

b123<-full_join(full_join(ef1,ef2),b3)
```

# Fishing and observation effort by ecoregion and countries

This section reproduces the graphical analyses provided by the previous section,
by country.

```{r exploeffbycountries}
#{{{
ef1<-D1%>%
	group_by(ctryname,year,L3name,ecoregion)%>%
	summarise(das=sum(daysAtSeaF,na.rm=T))%>%
	ungroup()
ef2<-D2%>%
	group_by(ctryname,year,L3name,ecoregion,monmethname)%>%
	summarise(das=sum(daysAtSeaOb,na.rm=T))%>%
	ungroup()
ef2bis<-D2%>%
	group_by(ctryname,year,L3name,ecoregion)%>%
	summarise(dasobs=sum(daysAtSeaOb,na.rm=T))%>%
	ungroup()
ef12<-full_join(ef1,ef2bis)
#}}}
```

```{r prepexpand}
#{{{
listidcountry<-D1%>%transmute(id=ctryname)%>%distinct()%>%arrange(id)
rmd1<-knitr::knit_expand("expand_ef1.rmd", 
			 id=listidcountry$id[1],
			 idyear=2023)
rmd<-rmd1
    for(i in 2:nrow(listidcountry)){
    #for(i in 2:2){
    rmdtmp<-knitr::knit_expand("expand_ef1.Rmd",
			 id=listidcountry$id[i],
			 idyear=2023)
    rmd<-paste0(rmd,rmdtmp,collapse="\n")
    }
#}}}
```

```{r renderexpandconf,results = "asis",eval=T,cache=F}
rendered <- knitr::knit(text = rmd, quiet = TRUE)
cat(rendered, sep = "\n")
```




# Fishing effort

This section describes the effort reported in the fishing effort table.

Seven effort indicators can be reported to WGBYC, 2 of them being mandatory,
only one is used here.

\newpage

# References
