---
title: "Tor A / WGBYC 2026"
subtittle: "subtitle" 
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: Laurent Dubroca 
bibliography: /Users/moi/zotero/exportdb/MyLibrary.bib 
number-sections: true
toc: true
toc-depth: 3
toc-expand: 3
format:
  html:
    html-math-method: katex
    css: styles.css
    code-fold: true
    page-layout: full
execute:
    warning: false
---

# Framework 

Analyses and results in a reproducible format (code and output) following
reproducible framework [@powers2019].



```{r renderstuff,eval=F,include=F}
#| cache: true
#| echo: false
#|
#describe dir contains in the readme.md
here::i_am("project.Rproj")
source(here("analyses", "printdir.R"))
printdir(here("data"), here("data", "README.md"))
printdir(here("outputs"), here("outputs", "README.md"))
quarto::quarto_render("report_data_read.qmd")
```


```{r libraryandco,eval=T,include=T}
#|
library(readxl)
library(openxlsx)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(arrow)
library(data.table)
library(tictoc)
library(knitr)
library(kableExtra)
library(DT)
library(sf)
library(httr)
library(jsonlite)
require(icesConnect)
library(officer)
library(flextable)
library(ggalluvial)


here::i_am("project.Rproj")
```

# Parameters

Some fixed parameters if needed here.

```{r parameters,eval=T,include=T}
#define the most recent year of the data provided (aka the present year minus one as a practical guess)
currentyear <- as.numeric(substr(Sys.Date(), 1, 4)) - 1
print(currentyear)
```

# Data

WGBYC wunderbar API use + ICES vocab request to ensure reproducibility (just
in case if, for example, some data are submitted again).

```{r dataco,eval=T,include=F}
#|
#source keyring (token stuff)
keyringfile <- here("analyses", "keyring.R")
if (file.exists(keyringfile)) {
    source(keyringfile)
} else {
    stop("keyring file not found")
}
#keyring file is a short R code used to tokenize data exchange
#go to https://data.ices.dk/token to generate the token
#the sourced code is :
#require(icesConnect)
#icesConnect::set_username("john.doe")
#icesConnect:::token_set_from_keyring(
#"thelongcaracterchainofthetokenprovidedbyiceshastobypastedhere",
#    username = "john.doe"
#)

# data request

if (file.exists(here("data", "D1_wgbyc.csv"))) {
    #read dowloaded data
    D1 <- fread(here("data", "D1_wgbyc.csv"))
    D2 <- fread(here("data", "D2_wgbyc.csv"))
    D3 <- fread(here("data", "D3_wgbyc.csv"))
} else {
    # read the data from ices API
    linkD1 <- "https://bycatch.ices.dk/api/GetD1_Fishing_effort"
    linkD2 <- "https://bycatch.ices.dk/api/GetD2_Bycatch_monitoring_effort"
    linkD3 <- "https://bycatch.ices.dk/api/GetD3_BycatchEvent"
    options(icesSAG.use_token = T)
    resp1 <- ices_get_jwt(linkD1)
    D1 <- content(resp1, as = "text")
    D1 <- fromJSON(D1)
    resp2 <- ices_get_jwt(linkD2)
    D2 <- content(resp2, as = "text")
    D2 <- fromJSON(D2)
    #D3 request by year: API slow by construction
    D3 <- data.frame()
    for (idyear in 2017:2025) {
        print(idyear)
        resp3 <- ices_get_jwt(paste0(linkD3, "?year=", idyear))
        D3tmp <- content(resp3, as = "text")
        D3tmp <- fromJSON(D3tmp)
        D3 <- rbind(D3, D3tmp)
    }
    #filtering on year of interest
    D1 <- D1 %>% filter(year %in% 2017:2025)
    D2 <- D2 %>% filter(year %in% 2017:2025)
    D3 <- D3 %>% filter(year %in% 2017:2025)
    #save file in csv format into data dir
    write.csv(D1, file = here("data", "D1_wgbyc.csv"), row.names = F)
    write.csv(D2, file = here("data", "D2_wgbyc.csv"), row.names = F)
    write.csv(D3, file = here("data", "D3_wgbyc.csv"), row.names = F)
}

#get extra info to label stuff
if (file.exists(here("data", "vocab", "BYC_Mon812Type.csv"))) {
    #read dowloaded data
    listvocab <- c(
        "ISO_3166",
        "GearType",
        "GearGroup",
        "BYC_MonProgr",
        "BYC_SamplingProtocol",
        "BYC_MonMethod",
        "BYC_Mon812Type"
    )
    f1 <- function(vocabname) {
        tmp <- fread(here("data", "vocab", paste0(vocabname, ".csv")))
        assign(vocabname, tmp, pos = ".GlobalEnv")
    }
    lapply(listvocab, f1)
} else {
    #downloaded locally some ICES vocab
    listvocab <- c(
        "ISO_3166",
        "GearType",
        "GearGroup",
        "BYC_MonProgr",
        "BYC_SamplingProtocol",
        "BYC_MonMethod",
        "BYC_Mon812Type"
    )
    f1 <- function(vocabname) {
        pipo <- icesVocab::getCodeList(vocabname)
        write.csv(
            pipo,
            file = here("data/vocab", paste0(vocabname, ".csv")),
            row.names = F
        )
    }
    lapply(listvocab, f1)
    #manual fix country name
    tmp <- fread(here("data", "vocab", paste0("ISO_3166", ".csv")))
    tmp <- tmp %>%
        mutate(Description = ifelse(Key == "PT-20", "Azores", Description)) %>%
        mutate(Description = ifelse(Key == "NL", "Netherlands", Description)) %>%
        mutate(Description = ifelse(Key == "GB", "United Kingdom", Description))
    write.csv(
        tmp,
        file = here("data/vocab", paste0("ISO_3166", ".csv")),
        row.names = F
    )
}
```



# Labelling stuff 

Data is labelled for human.

```{r labeldata,cache=TRUE,eval=T,echo=T,include=T}
#add human readable label with some basic test in case ICES vocab being not consistent
#D1
tmp <- D1 %>%
    left_join(
        ISO_3166 %>% transmute(country = Key, country_name = Description),
        by = c("country")
    ) %>%
    left_join(
        GearGroup %>% transmute(metierL3 = Key, metierL3_name = Description),
        by = c("metierL3")
    )
if (nrow(tmp) != nrow(D1)) {
    stop("pb D1 explo1")
} else {
    D1 <- tmp
    rm(tmp)
}
#D2
tmp <- D2 %>%
    left_join(
        ISO_3166 %>% transmute(country = Key, country_name = Description),
        by = c("country")
    ) %>%
    left_join(
        GearGroup %>% transmute(metierL3 = Key, metierL3_name = Description),
        by = c("metierL3")
    ) %>%
    left_join(
        BYC_MonMethod %>%
            transmute(monitoringMethod = Key, monitoringMethod_name = Description),
        by = c("monitoringMethod")
    ) %>%
    left_join(
        BYC_MonProgr %>%
            transmute(
                monitoringProgramType = Key,
                monitoringProgramType_name = Description
            ),
        by = c("monitoringProgramType")
    )
if (nrow(tmp) != nrow(D2)) {
    stop("pb D2 explo1")
} else {
    D2 <- tmp
    rm(tmp)
}
#D3
tmp <- D3 %>%
    left_join(
        ISO_3166 %>% transmute(country = Key, country_name = Description),
        by = c("country")
    ) %>%
    left_join(
        GearGroup %>% transmute(metierL3 = Key, metierL3_name = Description),
        by = c("metierL3")
    ) %>%
    left_join(
        BYC_MonMethod %>%
            transmute(monitoringMethod = Key, monitoringMethod_name = Description),
        by = c("monitoringMethod")
    ) %>%
    left_join(
        BYC_MonProgr %>%
            transmute(
                monitoringProgramType = Key,
                monitoringProgramType_name = Description
            ),
        by = c("monitoringProgramType")
    )
if (nrow(tmp) != nrow(D3)) {
    stop("pb D3 explo1")
} else {
    D3 <- tmp
    rm(tmp)
}
```

# Figures

# Green table

Comment the correction of the country labelling, and the subnational
submission of Azores and Northern Ireland.

```{r greentable,cache=TRUE,eval=T,echo=T,include=T}
#| column: page
## green table output for WGBYC report

#green table
greentableD1 <- D1 %>%
    group_by(country_name, year) %>%
    summarise(n = n()) %>%
    mutate(year = paste0("D1_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
greentableD2 <- D2 %>%
    group_by(country_name, year) %>%
    summarise(n = n()) %>%
    mutate(year = paste0("D2_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
greentableD3 <- D3 %>%
    group_by(country_name, year) %>%
    summarise(n = n()) %>%
    mutate(year = paste0("D3_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = TRUE)
greentable <- full_join(full_join(greentableD1, greentableD2), greentableD3) %>%
    ungroup() %>%
    data.frame()

#build up a submission table
subtableD1 <- D1 %>%
    group_by(country_name, year) %>%
    summarise(n = paste0(sort(unique(submissionYear)), collapse = ",")) %>%
    mutate(year = paste0("D1_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
subtableD2 <- D2 %>%
    group_by(country_name, year) %>%
    summarise(n = paste0(sort(unique(submissionYear)), collapse = ",")) %>%
    mutate(year = paste0("D2_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
subtableD3 <- D3 %>%
    group_by(country_name, year) %>%
    summarise(n = paste0(sort(unique(submissionYear)), collapse = ",")) %>%
    mutate(year = paste0("D3_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
subtable <- full_join(full_join(subtableD1, subtableD2), subtableD3) %>%
    ungroup() %>%
    data.frame()
fulltable <- subtable
#add the number of record in subtable with a good old double loop du seigneur
for (i in 1:nrow(subtable)) {
    for (j in 2:ncol(subtable)) {
        fulltable[i, j] <- paste(
            subtable[i, j],
            paste0("(", greentable[i, j], ")"),
            sep = "\n"
        )
    }
}
#cosmetic
nametab <- names(greentable)[-1]
nametab <- paste0(nametab, "='", substr(nametab, 4, 7), "'", collapse = ",")
nametab <- eval(parse(text = paste0("list(", nametab, ")")))
nametab <- c(list(country_name = "Year of data"), nametab)
bgcolormat <- ifelse(greentable[, -1] < 0.00001, "white", "darkgreen")
charcolormat <- matrix("white", nrow = nrow(subtable), ncol = ncol(subtable) - 1)
charcolormat[(currentyear + 1) == subtable[, -1]] <- "black"
#charcolormat["2021,2023"==subtable[,-1]]<-"yellow"
big_border <- fp_border(color = "black", width = 2)
#remove number
#greentable[is.na(greentable)]<-0

#param to fix column formatting
nbcolyear <- (ncol(subtable) - 1) / 3
posbigline <- c(1, nbcolyear + 1, 1 + nbcolyear * 2)

textcaption <- "Summary table of countries providing data submissions to ICES WGBYC with data on fishing effort, observer effort (either days at sea or other measurement, e.g. effort per haul or set), and bycatch records. Green = Data submission received, White = no data received, Black = data submitted this year. The year of submission is also provided with the number of record (in bracket) for each table and country."

ftgreentable <- flextable(fulltable) %>%
    bg(j = 2:ncol(greentable), bg = bgcolormat) %>%
    color(j = 2:ncol(greentable), color = charcolormat) %>%
    set_header_labels(values = nametab) %>%
    add_header_row(
        values = c(
            "",
            "Fishing Effort (D1 table)",
            "Monitoring Effort (D2 table)",
            "Bycatch Events (D3 table)"
        ),
        colwidths = c(1, nbcolyear, nbcolyear, nbcolyear)
    ) %>%
    theme_box() %>%
    align(align = "center", part = "header") %>%
    align(align = "center", part = "body") %>%
    vline(j = posbigline, border = big_border) %>%
    fontsize(part = "all", size = 5) %>%
    autofit(add_w = 0, add_h = 0) %>%
    set_caption(caption = textcaption)
ftgreentable


save_as_docx(
    ftgreentable,
    path = here("outputs", "greentable.docx"),
    pr_section = prop_section(
        page_size = page_size(orient = "landscape"),
        page_margins = page_mar(
            bottom = 0.1,
            top = .1,
            right = .1,
            left = .1,
            header = 0,
            footer = 0,
            gutter = 0
        )
    )
)
```

# Monitoring methods overview 

By days at sea, the second plot use the BEAM filter.


```{r monitoringoverviewdas,cache=TRUE,eval=T,echo=T,include=T}
# prepare data for fig 3.3 (reported monitoring method)
#all
monmethall <- D2 %>%
    group_by(monitoringMethod_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    group_by(year) %>%
    mutate(tot = sum(das)) %>%
    ungroup() %>%
    mutate(reldas = das / tot * 100) %>%
    mutate(type = "WGBYC database")
#BEAM filtering only
monmethbeam <- D2 %>%
    filter(!is.na(ecoregion) & !(ecoregion %in% c("North West Atlantic"))) %>%
    filter(!(monitoringMethod == "LB" & country != "PT")) %>%
    filter(!(monitoringMethod == "LB" & country != "PT-20")) %>%
    filter(!(monitoringMethod == "OTH" & country != "NO")) %>%
    filter(monitoringMethod != "PO") %>%
    filter(!(monitoringMethod == "VO" & country == "EE")) %>%
    group_by(monitoringMethod_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    group_by(year) %>%
    mutate(tot = sum(das)) %>%
    ungroup() %>%
    mutate(reldas = das / tot * 100) %>%
    mutate(type = "WGBYC databse with BEAM filtering")
monmethall <- rbind(monmethall, monmethbeam)

#plt output here
#monitoring methods overview global + BEAM filtering
plt <- ggplot(
    monmethall,
    aes(
        x = year,
        y = reldas,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_grid(~type) +
    theme_bw() +
    labs(fill = "Monitoring method") +
    ylab("Proportion of annual DAS by monitoring method (%)")


#by ctry
monmethctry <- D2 %>%
    group_by(monitoringMethod_name, country_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    group_by(year, country_name) %>%
    mutate(tot = sum(das)) %>%
    ungroup() %>%
    mutate(reldas = das / tot * 100)


print(plt)
textcaption <- "Total monitored (observed) days at sea reported in the WGBYC database and in the data filtered to compute BEAM estimatess per monitoring method (2017-2025) at-sea-observers,
electronic monitoring, port observers, and vessel crew observers, logbooks, other. The BEAM 
filtered for use in ToR C (see specific section, BEAM): at-sea-observ-
ers, electronic monitoring, vessel crew observers except for Estonia, and
other only for Norway."
print(textcaption)
filename <- here("outputs", "monitoring_method_overview.png")
png(filename, res = 300, width = 18, height = 18, unit = "cm")
print(plt)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        height = 18,
        width = 18,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))

#by country
plt <- ggplot(
    monmethctry,
    aes(
        x = year,
        y = reldas,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~country_name) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring method") +
    ylab("Proportion of annual DAS by monitoring method (%)")
#keep 1.8 fix country
print(plt)
png(
    here("outputs", "monitoring_method_country.png"),
    res = 300,
    width = 1600,
    height = 1200,
    unit = "px"
)
print(plt)
dev.off()

#moniroting by program
monitoringProgramType_namectry <- D2 %>%
    group_by(monitoringProgramType_name, country_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    group_by(year, country_name) %>%
    mutate(tot = sum(das)) %>%
    ungroup() %>%
    mutate(reldas = das / tot * 100)
#a graph
plt <- ggplot(
    monitoringProgramType_namectry,
    aes(
        x = year,
        y = reldas,
        stratum = monitoringProgramType_name,
        alluvium = monitoringProgramType_name,
        fill = monitoringProgramType_name,
        label = monitoringProgramType_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~country_name) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring program") +
    ylab("Proportion of annual DAS by monitoring program (%)")

#keep 1.9
print(plt)
png(
    here("outputs", "monitoring_program_country.png"),
    res = 300,
    width = 1600,
    height = 1200,
    unit = "px"
)
print(plt)
dev.off()

#
# incidents by montoring program and species class
#
pipo <- D3
pipo$nbinc <- apply(
    D3 %>% select(incidentsWithPingers, incidentsWithoutPingers),
    1,
    sum,
    na.rm = T
)
pipo <- pipo %>%
    mutate(classname = ifelse(is.na(classname), "Reptiles", classname)) %>%
    mutate(
        newclass = ifelse(grepl("Aves|Elasmo|Mamma|Rept", classname), classname, "Other")
    )

incprog <- pipo %>%
    group_by(monitoringMethod_name, year, newclass) %>%
    summarise(inc = sum(nbinc, na.rm = T), .groups = "drop") %>%
    ungroup()
#a graph
plt <- ggplot(
    incprog,
    aes(
        x = as.factor(year),
        y = inc,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~newclass, scale = "free") +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "inside",
        legend.position.inside = c(.85, 0.25)
    ) +
    labs(fill = "Monitoring method") +
    ylab("Number of reported incidents") +
    xlab("")

#keep 1.10 keep
print(plt)
png(
    here("outputs", "incidents_by_monitoringProgram.png"),
    res = 300,
    width = 1600,
    height = 1200,
    unit = "px"
)
print(plt)
dev.off()


#prepare data for sample coverage by monitoring effort by gear by georegions
monmethgear <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop")

#monitoring methods: ecoregion/gear view
monmethgear1 <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    filter(!monitoringMethod_name %in% c("Logbook", "Port observers"))
monmethgear2 <- monmethgear1 %>%
    mutate(metierL3_name = "All metier") %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(das = sum(das, na.rm = T), .groups = "drop")
monmethgear <- rbind(monmethgear1, monmethgear2) %>%
    mutate(monitoringMethod_name = as.factor(monitoringMethod_name))


listecoregion <- na.omit(unique(monmethgear$ecoregion))
for (i in listecoregion) {
    tmp <- monmethgear[monmethgear$ecoregion == i, ] %>%
        filter(!is.na(metierL3_name))
    plt <- ggplot(
        tmp,
        aes(
            x = year,
            y = das,
            stratum = monitoringMethod_name,
            alluvium = monitoringMethod_name,
            fill = monitoringMethod_name,
            label = monitoringMethod_name
        )
    ) +
        geom_flow(stat = "alluvium") +
        geom_stratum() +
        facet_wrap(~metierL3_name, scale = "free") +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
        labs(fill = "Monitoring method") +
        ggtitle(i) +
        ylab("Sum of annual DAS by monitoring method")
    print(plt)
    filename <- here("outputs", paste0("monitoring_method_", gsub(" ", "", i), ".png"))
    png(filename, res = 300, width = 1600, height = 1200, unit = "px")
    print(plt)
    dev.off()
}

#prepare data for sample coverage by monitoring effort by georegions
monmethecoregion <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    filter(!is.na(ecoregion))

plt <- ggplot(
    monmethecoregion %>% mutate(ecoregion = sub(" ", "\n", ecoregion)),
    aes(
        x = year,
        y = das,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~ecoregion, scale = "free") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring method") +
    ggtitle("Monitoring program by ecoregion") +
    ylab("Sum of annual DAS by monitoring method")
filename <- here("outputs", "monitoring_method_ecoregion.png")
print(plt)
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()
```


# Monitoring methods overview trips

By trips : last year figures were wrong (according to the script).

```{r monitoringoverviewtrips,cache=TRUE,eval=T,echo=T,include=T}
# prepare data for fig 3.3 (reported monitoring method)
#all
monmethall <- D2 %>%
    group_by(monitoringMethod_name, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    group_by(year) %>%
    mutate(tot = sum(trips)) %>%
    ungroup() %>%
    mutate(reltrips = trips / tot * 100)
#by ctry
monmethctry <- D2 %>%
    group_by(monitoringMethod_name, country_name, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    group_by(year, country_name) %>%
    mutate(tot = sum(trips)) %>%
    ungroup() %>%
    mutate(reltrips = trips / tot * 100)
#plt output here
#monitoring methods overview
plt <- ggplot(
    monmethall,
    aes(
        x = year,
        y = reltrips,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    theme_bw() +
    labs(fill = "Monitoring method") +
    ylab("Proportion of annual trips by monitoring method (%)")
print(plt)
filename <- here("outputs", "monitoring_method_overview_trips.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()

#clean
plt <- ggplot(
    monmethctry,
    aes(
        x = year,
        y = reltrips,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~country_name) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring method") +
    ylab("Proportion of annual trips by monitoring method (%)")
print(plt)
filename <- here("outputs", "monitoring_method_country_trips.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()

#moniroting by program
monprogctry <- D2 %>%
    group_by(monitoringProgramType_name, country_name, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    group_by(year, country_name) %>%
    mutate(tot = sum(trips)) %>%
    ungroup() %>%
    mutate(reltrips = trips / tot * 100)
#a graph
plt <- ggplot(
    monprogctry,
    aes(
        x = year,
        y = reltrips,
        stratum = monitoringProgramType_name,
        alluvium = monitoringProgramType_name,
        fill = monitoringProgramType_name,
        label = monitoringProgramType_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~country_name) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring program") +
    ylab("Proportion of annual trips by monitoring program (%)")

print(plt)
filename <- here("outputs", "monitoring_program_country_trips.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()

#prepare data for sample coverage by monitoring effort by gear by georegions
monmethgear <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(trips = sum(daysAtSeaOb, na.rm = T), .groups = "drop")

#monitoring methods: ecoregion/gear view
monmethgear1 <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    filter(!monitoringMethod_name %in% c("Logbook", "Port observers"))
monmethgear2 <- monmethgear1 %>%
    mutate(metierL3_name = "All metier") %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(trips = sum(trips, na.rm = T), .groups = "drop")
monmethgear <- rbind(monmethgear1, monmethgear2) %>%
    mutate(monitoringMethod_name = as.factor(monitoringMethod_name))


listecoregion <- na.omit(unique(monmethgear$ecoregion))
for (i in listecoregion) {
    tmp <- monmethgear[monmethgear$ecoregion == i, ] %>%
        filter(!is.na(metierL3_name))
    plt <- ggplot(
        tmp,
        aes(
            x = year,
            y = trips,
            stratum = monitoringMethod_name,
            alluvium = monitoringMethod_name,
            fill = monitoringMethod_name,
            label = monitoringMethod_name
        )
    ) +
        geom_flow(stat = "alluvium") +
        geom_stratum() +
        facet_wrap(~metierL3_name, scale = "free") +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
        labs(fill = "Monitoring method") +
        ggtitle(i) +
        ylab("Sum of annual trips by monitoring method")
    print(plt)
    filename <- here(
        "outputs",
        paste0("monitoring_method_", gsub(" ", "", i), "_trips.png")
    )
    png(filename, res = 300, width = 1600, height = 1200, unit = "px")
    print(plt)
    dev.off()
}

#prepare data for sample coverage by monitoring effort by georegions
monmethecoregion <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    filter(!is.na(ecoregion))

plt <- ggplot(
    monmethecoregion %>% mutate(ecoregion = sub(" ", "\n", ecoregion)),
    aes(
        x = year,
        y = trips,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~ecoregion, scale = "free") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring method") +
    ggtitle(i) +
    ylab("Sum of annual trips by monitoring method")
print(plt)
filename <- here("outputs", "monitoring_method_ecoregion_trips.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()
```

# EM figures 


```{r EMfig,cache=TRUE,eval=T,echo=T,include=T}

#### calculate percentages for text ####

#"Along with at-sea observers, this method yet represents the source of a majority of bycatch
#records reported (Figure 5), with 40% of reported mammal and 30% of seabird bycatch incidents recorded with EM.

## Pivot the info
NofIncidents <- D3 %>%
    group_by(year, monitoringMethod_name, classname) %>%
    summarize(
        sum_incidents = sum(incidentsWithPingers, na.rm = TRUE) +
            sum(incidentsWithoutPingers, na.rm = TRUE),
    )

#Use only Electronic Monitoring
Aves <- NofIncidents %>%
    filter(classname == "Aves")

Aves <- Aves %>%
    group_by(year, monitoringMethod_name) %>%
    summarise(sum_incidents = sum(sum_incidents)) %>%
    mutate(total_incidents_per_year = sum(sum_incidents, na.rm = TRUE)) %>%
    mutate(percentage = (sum_incidents / total_incidents_per_year) * 100)

Aves

#Use only Electronic Monitoring
Mammalia <- NofIncidents %>%
    filter(classname == "Mammalia")

Mammalia <- Mammalia %>%
    group_by(year, monitoringMethod_name) %>%
    summarise(sum_incidents = sum(sum_incidents)) %>%
    mutate(total_incidents_per_year = sum(sum_incidents, na.rm = TRUE)) %>%
    mutate(percentage = (sum_incidents / total_incidents_per_year) * 100)

Mammalia

#### Trips per country per year ####

#Use only Electronic Monitoring
EM_monitoringeffort <- D2 %>%
    filter(monitoringMethod == "EM")
## Pivot the info
EM_Grouped <- EM_monitoringeffort %>%
    group_by(year, country_name) %>%
    summarize(total_tripsOb = sum(tripsOb, na.rm = TRUE))

# PLOT EM total trips observed
plt <- ggplot(EM_Grouped, aes(x = year, y = total_tripsOb, fill = country_name)) +
    geom_bar(stat = "identity", position = "stack") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Country") +
    ylab("Total trips observed using EM")

print(plt)
filename <- here("outputs", "EM_trips_obs_country.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()

#### N of incidents per year per taxa, stack countries ####

#Use only Electronic Monitoring
EM_taxa_bycatch <- D3 %>%
    filter(monitoringMethod == "EM")

# Drop na values and drop teleostei (Because only 3 incidents! Check this later if needed!)
EM_taxa_bycatch <- EM_taxa_bycatch %>%
    filter(!is.na(classname) & classname != "Teleostei")

## Pivot the info
EM_Taxa_Grouped <- EM_taxa_bycatch %>%
    group_by(year, country_name, classname) %>%
    summarize(
        sum_incidents = sum(incidentsWithPingers, na.rm = TRUE) +
            sum(incidentsWithoutPingers, na.rm = TRUE),
        sum_individuals = sum(individualsWithPingers, na.rm = TRUE) +
            sum(individualsWithoutPingers, na.rm = TRUE)
    )

#Plot individuals per taxa
plt <- ggplot(EM_Taxa_Grouped, aes(x = year, y = sum_individuals, fill = country_name)) +
    geom_bar(stat = "identity", position = "stack") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Country") +
    ylab("Total number of individuals") +
    facet_wrap(~classname, scales = "free") +
    scale_x_continuous(limits = c(2018.5, max(EM_Taxa_Grouped$year + 0.5))) #this is just to make all x-axes the same

print(plt)
filename <- here("outputs", "EM_individuals_country.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()

#Plot incidents per taxa
plt <- ggplot(EM_Taxa_Grouped, aes(x = year, y = sum_incidents, fill = country_name)) +
    geom_bar(stat = "identity", position = "stack") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Country") +
    ylab("Total number of incidents") +
    facet_wrap(~classname, scales = "free") +
    scale_x_continuous(limits = c(2018.5, max(EM_Taxa_Grouped$year + 0.5))) #this is just to make all x-axes the same
plot(plt)

print(plt)
filename <- here("outputs", "EM_incidents_country.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()
```

# Long tables production 

Text

```{r longtab,cache=TRUE,eval=T,echo=T,include=T}
#long-table-container style="max-height: 500px; overflow-y: auto;"

#big long useless table
#Annex_table
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, areaCode, metierL3, monitoringMethod) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3, areaCode) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, areaCode, metierL3, monitoringMethod, species, classname) %>%
    summarise(
        nb = sum(individualsWithPingers, individualsWithoutPingers, na.rm = T),
        inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)
    ) %>%
    ungroup()

tab123 <- left_join(left_join(tab2, tab1), tab3) %>%
    filter(!is.na(species)) %>%
    transmute(
        ecoregion,
        areaCode,
        metierL3,
        dastot,
        monitoringMethod,
        dasobs,
        moncov = 100 * (dasobs / dastot),
        classname,
        species,
        nb,
        inc
    ) %>%
    mutate(dasobs = ifelse(dasobs == 0, NA, dasobs)) %>%
    mutate(moncov = ifelse(dasobs == 0, NA, moncov)) %>%
    transmute(
        Ecoregion = ecoregion,
        AreaCode = areaCode,
        MetierL3 = metierL3,
        `Fishing Effort (das)` = dastot,
        MonitoringMethod = monitoringMethod,
        `Total Observed Effort (das)` = dasobs,
        `Monitoring Coverage (%)` = moncov,
        classname,
        Species = species,
        Total_No_Specimens = nb,
        Incidents = inc
    )

head(tab123)

#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(
    wb,
    file = here("outputs", "TOR_A_long_table_bycatch_only.xlsx"),
    overwrite = T
)

#big long useless table with all info
#Annex_table
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, areaCode, metierL3, monitoringMethod) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3, areaCode) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, areaCode, metierL3, monitoringMethod, species, classname) %>%
    summarise(
        nb = sum(individualsWithPingers, individualsWithoutPingers, na.rm = T),
        inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)
    ) %>%
    ungroup()

tab123 <- full_join(full_join(tab2, tab1), tab3) %>% #filter(!is.na(species))%>%
    transmute(
        ecoregion,
        areaCode,
        metierL3,
        dastot,
        monitoringMethod,
        dasobs,
        moncov = 100 * (dasobs / dastot),
        classname,
        species,
        nb,
        inc
    ) %>%
    mutate(dasobs = ifelse(dasobs == 0, NA, dasobs)) %>%
    mutate(moncov = ifelse(dasobs == 0, NA, moncov)) %>%
    transmute(
        Ecoregion = ecoregion,
        AreaCode = areaCode,
        MetierL3 = metierL3,
        `Fishing Effort (das)` = dastot,
        MonitoringMethod = monitoringMethod,
        `Total Observed Effort (das)` = dasobs,
        `Monitoring Coverage (%)` = moncov,
        classname,
        Species = species,
        Total_No_Specimens = nb,
        Incidents = inc
    )

#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(wb, file = here("outputs", "TOR_A_long_table.xlsx"), overwrite = T)

#short table
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3, monitoringMethod) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3, monitoringMethod, species, classname) %>%
    summarise(
        nb = sum(individualsWithPingers, individualsWithoutPingers, na.rm = T),
        inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)
    ) %>%
    ungroup()

tab123 <- left_join(left_join(tab2, tab1), tab3) %>%
    filter(!is.na(species)) %>%
    transmute(
        ecoregion,
        metierL3,
        dastot,
        monitoringMethod,
        dasobs,
        moncov = 100 * (dasobs / dastot),
        classname,
        species,
        nb,
        inc
    ) %>%
    mutate(dasobs = ifelse(dasobs == 0, NA, dasobs)) %>%
    mutate(moncov = ifelse(dasobs == 0, NA, moncov)) %>%
    transmute(
        Ecoregion = ecoregion,
        MetierL3 = metierL3,
        `Fishing Effort (das)` = dastot,
        MonitoringMethod = monitoringMethod,
        `Total Observed Effort (das)` = dasobs,
        `Monitoring Coverage (%)` = moncov,
        classname,
        Species = species,
        Total_No_Specimens = nb,
        Incidents = inc
    )

#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(wb, file = here("outputs", "TOR_A_short_table.xlsx"), overwrite = T)


#shorter table
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    mutate(classname = ifelse(is.na(classname), "Reptiles", classname)) %>%
    group_by(Ecoregion = ecoregion, classname) %>%
    summarise(
        nbspp = n_distinct(species),
        nb = sum(individualsWithPingers, individualsWithoutPingers, na.rm = T),
        inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)
    ) %>%
    ungroup() %>%
    tidyr::pivot_longer(nbspp:inc, names_to = "variable") %>%
    tidyr::pivot_wider(values_from = value, names_from = classname)
tab3$variable[tab3$variable == "nbspp"] <- "Species"
tab3$variable[tab3$variable == "nb"] <- "Individuals"
tab3$variable[tab3$variable == "inc"] <- "Incidents"


tab123 <- left_join(tab1, tab2) %>%
    filter(!is.na(ecoregion)) %>%
    transmute(
        Ecoregion = ecoregion,
        `Fishing Effort (das)` = dastot,
        `Total Observed Effort (das)` = dasobs,
        `Monitoring Coverage (%)` = 100 * (dasobs / dastot)
    ) %>%
    left_join(tab3) %>%
    filter(Ecoregion != "")


#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(wb, file = here("outputs", "TOR_A_shorter_table.xlsx"), overwrite = T)

pipo <- flextable(tab123) %>%
    merge_v(
        j = ~ Ecoregion +
            `Fishing Effort (das)` +
            `Total Observed Effort (das)` +
            `Monitoring Coverage (%)`
    ) %>%
    hline(i = seq(3, nrow(tab123), 3)) %>%
    valign(j = 1:4, valign = "top") %>%
    colformat_double(j = 4, digits = 2) %>%
    colformat_double(j = c(2:3, 5:ncol(tab123)), digits = 0) %>%
    bg(part = "header", bg = "grey30") %>%
    color(part = "header", color = "ghostwhite") %>%
    italic(j = 1) %>%
    fit_to_width(max_width = 11.7)
pipo

#save as docx object
save_as_docx(
    pipo,
    path = here("outputs", "TOR_A_shorter_table.docx"),
    pr_section = prop_section(
        page_size = page_size(orient = "landscape"),
        page_margins = page_mar(
            bottom = 0,
            top = 0,
            right = 0,
            left = 0,
            header = 0,
            footer = 0,
            gutter = 0
        )
    )
)


#shorter table detailed effort
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3_name) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3_name) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab12 <- full_join(tab1, tab2) %>%
    mutate(moncov = 100 * (dasobs / dastot)) %>%
    mutate(moncov = ifelse(is.finite(moncov), moncov, 0)) %>%
    mutate(info = paste0(round(dastot, 0), " das (", round(moncov, 3), "% cov.)")) %>%
    select(ecoregion, metierL3_name, info)
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3_name) %>%
    summarise(inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)) %>%
    mutate(inc = ifelse(is.finite(inc), inc, 0)) %>%
    ungroup()
tab123 <- full_join(tab12, tab3) %>%
    mutate(inc = ifelse(is.finite(inc), inc, 0)) %>%
    mutate(info = paste0(info, " das/ ", inc, " inc.")) %>%
    select(ecoregion, metierL3_name, info) %>%
    pivot_wider(values_from = info, names_from = metierL3_name) %>%
    filter(!is.na(ecoregion))


flextable(tab123) %>%
    autofit() %>%
    set_table_properties(
        opts_html = list(
            scroll = list(
                height = "400px", # Limits height and enables vertical scroll
                freeze_first_column = TRUE # Keeps first column visible when scrolling right
            )
        )
    )
#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(
    wb,
    file = here("output", "TOR_A_cool_effort_shorter_table.xlsx"),
    overwrite = T
)
```

# Stuff 

Text

```{r anotherchunk,cache=TRUE,eval=T,echo=T,include=T}
#code
```



# R session information

```{r Rsession,cache=TRUE,eval=T,echo=T,include=T}
sessioninfo::session_info()
```

# Bibliography
