---
title: "Tor A / WGBYC 2026"
subtittle: "subtitle" 
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: Laurent Dubroca 
bibliography: /Users/moi/zotero/exportdb/MyLibrary.bib 
number-sections: true
toc: true
toc-depth: 3
toc-expand: 3
format:
  html:
    html-math-method: katex
    css: styles.css
    code-fold: true
    page-layout: full
execute:
    warning: false
---

# Framework 

Analyses and results in a reproducible format (code and output) following
reproducible framework [@powers2019].



```{r renderstuff,eval=F,include=F}
#| cache: true
#| echo: false
#|
#describe dir contains in the readme.md
here::i_am("project.Rproj")
source(here("analyses", "printdir.R"))
printdir(here("data"), here("data", "README.md"))
printdir(here("outputs"), here("outputs", "README.md"))
quarto::quarto_render("report_data_read.qmd")
```


```{r libraryandco,eval=T,include=T}
#|
library(readxl)
library(openxlsx)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(arrow)
library(data.table)
library(tictoc)
library(knitr)
library(kableExtra)
library(DT)
library(sf)
library(httr)
library(jsonlite)
require(icesConnect)
library(officer)
library(flextable)
library(ggalluvial)
library(httr)
library(jsonlite)
library(grid)
library(glue)
library(rnaturalearth)
library(rnaturalearthdata)
library(lwgeom)
library(viridis)
library(ggforce)
library(stringr)
library(gridExtra)

here::i_am("project.Rproj")
```

# Parameters

Some fixed parameters if needed here.

```{r parameters,eval=T,include=T}
#define the most recent year of the data provided (aka the present year minus one as a practical guess)
currentyear <- as.numeric(substr(Sys.Date(), 1, 4)) - 1
print(currentyear)
```

# Data

WGBYC wunderbar API use + ICES vocab request to ensure reproducibility (just
in case if, for example, some data are submitted again).

```{r dataco,eval=T,include=F}
#|
#source keyring (token stuff)
keyringfile <- here("analyses", "keyring.R")
if (file.exists(keyringfile)) {
    source(keyringfile)
} else {
    stop("keyring file not found")
}
#keyring file is a short R code used to tokenize data exchange
#go to https://data.ices.dk/token to generate the token
#the sourced code is :
#require(icesConnect)
#icesConnect::set_username("john.doe")
#icesConnect:::token_set_from_keyring(
#"thelongcaracterchainofthetokenprovidedbyiceshastobypastedhere",
#    username = "john.doe"
#)

# data request

if (file.exists(here("data", "D1_wgbyc.csv"))) {
    #read dowloaded data
    D1 <- fread(here("data", "D1_wgbyc.csv"))
    D2 <- fread(here("data", "D2_wgbyc.csv"))
    D3 <- fread(here("data", "D3_wgbyc.csv"))
} else {
    # read the data from ices API
    linkD1 <- "https://bycatch.ices.dk/api/GetD1_Fishing_effort"
    linkD2 <- "https://bycatch.ices.dk/api/GetD2_Bycatch_monitoring_effort"
    linkD3 <- "https://bycatch.ices.dk/api/GetD3_BycatchEvent"
    options(icesSAG.use_token = T)
    resp1 <- ices_get_jwt(linkD1)
    D1 <- content(resp1, as = "text")
    D1 <- fromJSON(D1)
    resp2 <- ices_get_jwt(linkD2)
    D2 <- content(resp2, as = "text")
    D2 <- fromJSON(D2)
    #D3 request by year: API slow by construction
    D3 <- data.frame()
    for (idyear in 2017:2025) {
        print(idyear)
        resp3 <- ices_get_jwt(paste0(linkD3, "?year=", idyear))
        D3tmp <- content(resp3, as = "text")
        D3tmp <- fromJSON(D3tmp)
        D3 <- rbind(D3, D3tmp)
    }
    #filtering on year of interest
    D1 <- D1 %>% filter(year %in% 2017:2025)
    D2 <- D2 %>% filter(year %in% 2017:2025)
    D3 <- D3 %>% filter(year %in% 2017:2025)
    #save file in csv format into data dir
    write.csv(D1, file = here("data", "D1_wgbyc.csv"), row.names = F)
    write.csv(D2, file = here("data", "D2_wgbyc.csv"), row.names = F)
    write.csv(D3, file = here("data", "D3_wgbyc.csv"), row.names = F)
}

#get extra info to label stuff
if (file.exists(here("data", "vocab", "BYC_Mon812Type.csv"))) {
    #read dowloaded data
    listvocab <- c(
        "ISO_3166",
        "GearType",
        "GearGroup",
        "BYC_MonProgr",
        "BYC_SamplingProtocol",
        "BYC_MonMethod",
        "BYC_Mon812Type"
    )
    f1 <- function(vocabname) {
        tmp <- fread(here("data", "vocab", paste0(vocabname, ".csv")))
        assign(vocabname, tmp, pos = ".GlobalEnv")
    }
    lapply(listvocab, f1)
} else {
    #downloaded locally some ICES vocab
    listvocab <- c(
        "ISO_3166",
        "GearType",
        "GearGroup",
        "BYC_MonProgr",
        "BYC_SamplingProtocol",
        "BYC_MonMethod",
        "BYC_Mon812Type"
    )
    f1 <- function(vocabname) {
        pipo <- icesVocab::getCodeList(vocabname)
        write.csv(
            pipo,
            file = here("data/vocab", paste0(vocabname, ".csv")),
            row.names = F
        )
    }
    dir.create(here("data", "vocab"))
    lapply(listvocab, f1)
    #manual fix country name
    tmp <- fread(here("data", "vocab", paste0("ISO_3166", ".csv")))
    tmp <- tmp %>%
        mutate(Description = ifelse(Key == "PT-20", "Azores", Description)) %>%
        mutate(Description = ifelse(Key == "NL", "Netherlands", Description)) %>%
        mutate(Description = ifelse(Key == "GB", "United Kingdom", Description))
    write.csv(
        tmp,
        file = here("data/vocab", paste0("ISO_3166", ".csv")),
        row.names = F
    )
}
```



# Labelling stuff 

Data is labelled for human.

```{r labeldata,cache=TRUE,eval=T,echo=T,include=T}
#add human readable label with some basic test in case ICES vocab being not consistent
#D1
tmp <- D1 %>%
    left_join(
        ISO_3166 %>% transmute(country = Key, country_name = Description),
        by = c("country")
    ) %>%
    left_join(
        GearGroup %>% transmute(metierL3 = Key, metierL3_name = Description),
        by = c("metierL3")
    )
if (nrow(tmp) != nrow(D1)) {
    stop("pb D1 explo1")
} else {
    D1 <- tmp
    rm(tmp)
}
#D2
tmp <- D2 %>%
    left_join(
        ISO_3166 %>% transmute(country = Key, country_name = Description),
        by = c("country")
    ) %>%
    left_join(
        GearGroup %>% transmute(metierL3 = Key, metierL3_name = Description),
        by = c("metierL3")
    ) %>%
    left_join(
        BYC_MonMethod %>%
            transmute(monitoringMethod = Key, monitoringMethod_name = Description),
        by = c("monitoringMethod")
    ) %>%
    left_join(
        BYC_MonProgr %>%
            transmute(
                monitoringProgramType = Key,
                monitoringProgramType_name = Description
            ),
        by = c("monitoringProgramType")
    )
if (nrow(tmp) != nrow(D2)) {
    stop("pb D2 explo1")
} else {
    D2 <- tmp
    rm(tmp)
}
#D3
tmp <- D3 %>%
    left_join(
        ISO_3166 %>% transmute(country = Key, country_name = Description),
        by = c("country")
    ) %>%
    left_join(
        GearGroup %>% transmute(metierL3 = Key, metierL3_name = Description),
        by = c("metierL3")
    ) %>%
    left_join(
        BYC_MonMethod %>%
            transmute(monitoringMethod = Key, monitoringMethod_name = Description),
        by = c("monitoringMethod")
    ) %>%
    left_join(
        BYC_MonProgr %>%
            transmute(
                monitoringProgramType = Key,
                monitoringProgramType_name = Description
            ),
        by = c("monitoringProgramType")
    )
if (nrow(tmp) != nrow(D3)) {
    stop("pb D3 explo1")
} else {
    D3 <- tmp
    rm(tmp)
}
```

# Figures table 

```{r}
list_table <- list(
  #  ID |    Figure   | Descritpion 
  plt_fishing_effort = c("Figure 1.3", ""), 
  plt_monitoring_effort = c("Figure 1.4", ""), 
  plt_monitoring_coverage = c("Figure 1.5", ""), 
  data_toPlot_allPETS_plot = c("Figure 1.6", ""),
  plt_monitoring_method = c("Figure 1.7", ""), 
  plt_monitoring_program_country = c("Figure 1.8", ""),
  plt_incident_programm = c("Figure 1.9", ""), 
  plt_monitoring_method_country = c("Figure 1.10", ""),
  plt_em_total_das_trip = c("Figure 1.11", ""),
  plt_em_plt_incidents = c("Figure 1.12", "")
  

  
)

```


# Green table

Comment the correction of the country labelling, and the subnational
submission of Azores and Northern Ireland.

```{r greentable,cache=TRUE,eval=T,echo=T,include=T}
#| column: page
## green table output for WGBYC report

#green table
greentableD1 <- D1 %>%
    group_by(country_name, year) %>%
    summarise(n = n()) %>%
    mutate(year = paste0("D1_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
greentableD2 <- D2 %>%
    group_by(country_name, year) %>%
    summarise(n = n()) %>%
    mutate(year = paste0("D2_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
greentableD3 <- D3 %>%
    group_by(country_name, year) %>%
    summarise(n = n()) %>%
    mutate(year = paste0("D3_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = TRUE)
greentable <- full_join(full_join(greentableD1, greentableD2), greentableD3) %>%
    ungroup() %>%
    data.frame()

#build up a submission table
subtableD1 <- D1 %>%
    group_by(country_name, year) %>%
    summarise(n = paste0(sort(unique(submissionYear)), collapse = ",")) %>%
    mutate(year = paste0("D1_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
subtableD2 <- D2 %>%
    group_by(country_name, year) %>%
    summarise(n = paste0(sort(unique(submissionYear)), collapse = ",")) %>%
    mutate(year = paste0("D2_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
subtableD3 <- D3 %>%
    group_by(country_name, year) %>%
    summarise(n = paste0(sort(unique(submissionYear)), collapse = ",")) %>%
    mutate(year = paste0("D3_", year)) %>%
    tidyr::pivot_wider(values_from = n, names_from = year, names_sort = T)
subtable <- full_join(full_join(subtableD1, subtableD2), subtableD3) %>%
    ungroup() %>%
    data.frame()
fulltable <- subtable
#add the number of record in subtable with a good old double loop du seigneur
for (i in 1:nrow(subtable)) {
    for (j in 2:ncol(subtable)) {
        fulltable[i, j] <- paste(
            subtable[i, j],
            paste0("(", greentable[i, j], ")"),
            sep = "\n"
        )
    }
}
#cosmetic
nametab <- names(greentable)[-1]
nametab <- paste0(nametab, "='", substr(nametab, 4, 7), "'", collapse = ",")
nametab <- eval(parse(text = paste0("list(", nametab, ")")))
nametab <- c(list(country_name = "Year of data"), nametab)
bgcolormat <- ifelse(greentable[, -1] < 0.00001, "white", "darkgreen")
charcolormat <- matrix("white", nrow = nrow(subtable), ncol = ncol(subtable) - 1)
charcolormat[(currentyear + 1) == subtable[, -1]] <- "black"
#charcolormat["2021,2023"==subtable[,-1]]<-"yellow"
big_border <- fp_border(color = "black", width = 2)
#remove number
#greentable[is.na(greentable)]<-0

#param to fix column formatting
nbcolyear <- (ncol(subtable) - 1) / 3
posbigline <- c(1, nbcolyear + 1, 1 + nbcolyear * 2)

textcaption <- "Summary table of countries providing data submissions to ICES WGBYC with data on fishing effort, observer effort (either days at sea or other measurement, e.g. effort per haul or set), and bycatch records. Green = Data submission received, White = no data received, Black = data submitted this year. The year of submission is also provided with the number of record (in bracket) for each table and country."

ftgreentable <- flextable(fulltable) %>%
    bg(j = 2:ncol(greentable), bg = bgcolormat) %>%
    color(j = 2:ncol(greentable), color = charcolormat) %>%
    set_header_labels(values = nametab) %>%
    add_header_row(
        values = c(
            "",
            "Fishing Effort (D1 table)",
            "Monitoring Effort (D2 table)",
            "Bycatch Events (D3 table)"
        ),
        colwidths = c(1, nbcolyear, nbcolyear, nbcolyear)
    ) %>%
    theme_box() %>%
    align(align = "center", part = "header") %>%
    align(align = "center", part = "body") %>%
    vline(j = posbigline, border = big_border) %>%
    fontsize(part = "all", size = 5) %>%
    autofit(add_w = 0, add_h = 0) %>%
    set_caption(caption = textcaption)
ftgreentable


save_as_docx(
    ftgreentable,
    path = here("outputs", "greentable.docx"),
    pr_section = prop_section(
        page_size = page_size(orient = "landscape"),
        page_margins = page_mar(
            bottom = 0.1,
            top = .1,
            right = .1,
            left = .1,
            header = 0,
            footer = 0,
            gutter = 0
        )
    )
)
```


# Merge country 

```{r}
D1 <- D1 %>%
    mutate(country_name = ifelse(country_name == "Azores", "Portugal", country_name)) %>%
    mutate(country_name = ifelse( country_name == "Northern Ireland", "United Kingdom", country_name))
D2 <- D2 %>%
    mutate(country_name = ifelse(country_name == "Azores", "Portugal", country_name)) %>%
    mutate(country_name = ifelse( country_name == "Northern Ireland", "United Kingdom", country_name))
D3 <- D3 %>%
    mutate(country_name = ifelse(country_name == "Azores", "Portugal", country_name)) %>%
    mutate(country_name = ifelse( country_name == "Northern Ireland", "United Kingdom", country_name))
```


# Monitoring methods overview 

By days at sea, the second plot use the BEAM filter.


```{r monitoringoverviewdas,cache=TRUE,eval=T,echo=T,include=T}
# prepare data for fig 3.3 (reported monitoring method)
#all
monmethall <- D2 %>%
    group_by(monitoringMethod_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    group_by(year) %>%
    mutate(tot = sum(das)) %>%
    ungroup() %>%
    mutate(reldas = das / tot * 100) %>%
    mutate(type = "WGBYC database")
#BEAM filtering only
monmethbeam <- D2 %>%
    filter(!is.na(ecoregion) & !(ecoregion %in% c("North West Atlantic"))) %>%
    filter(!(monitoringMethod == "LB" & country != "PT")) %>%
    filter(!(monitoringMethod == "LB" & country != "PT-20")) %>%
    filter(!(monitoringMethod == "OTH" & country != "NO")) %>%
    filter(monitoringMethod != "PO") %>%
    filter(!(monitoringMethod == "VO" & country == "EE")) %>%
    group_by(monitoringMethod_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    group_by(year) %>%
    mutate(tot = sum(das)) %>%
    ungroup() %>%
    mutate(reldas = das / tot * 100) %>%
    mutate(type = "WGBYC databse with BEAM filtering")
monmethall <- rbind(monmethall, monmethbeam)

#plt output here
#monitoring methods overview global + BEAM filtering
plt_monitoring_method <- ggplot(
    monmethall,
    aes(
        x = year,
        y = reldas,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_grid(~type) +
    theme_bw() +
    labs(fill = "Monitoring method") +
    ylab("Proportion of annual DAS by monitoring method (%)")


#by ctry
monmethctry <- D2 %>%
    group_by(monitoringMethod_name, country_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    group_by(year, country_name) %>%
    mutate(tot = sum(das)) %>%
    ungroup() %>%
    mutate(reldas = das / tot * 100)


print(plt_monitoring_method)
textcaption <- "Total monitored (observed) days at sea reported in the WGBYC database and in the data filtered to compute BEAM estimatess per monitoring method (2017-2025) at-sea-observers, electronic monitoring, port observers, and vessel crew observers, logbooks, other. The BEAM filtered for use in ToR C (see specific section, BEAM): at-sea-observers, electronic monitoring, vessel crew observers except for Estonia, and other only for Norway."
print(textcaption)
filename <- here("outputs", glue(list_table$plt_monitoring_method[1], "_monitoring_method_overview.png"))
png(filename, res = 300, width = 18, height = 18, unit = "cm")
print(plt_monitoring_method)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        height = 18,
        width = 18,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```
# Monitoring methods overview by country

```{r monitoringoverviewdas_by_country, cache=TRUE,eval=T,echo=T,include=T}
#by country
plt_monitoring_method_country <- ggplot(
    monmethctry,
    aes(
        x = year,
        y = reldas,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~country_name) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring method") +
    ylab("Proportion of annual DAS by monitoring method (%)")
#keep 1.8 fix country
print(plt_monitoring_method_country)
textcaption <- "test"
print(textcaption)
filename <- here("outputs", glue(list_table$plt_monitoring_method_country[1], "_monitoring_method_country.png"))
png(filename, res = 300, width = 22, height = 16, unit = "cm")
print(plt_monitoring_method_country)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        height = 16,
        width = 18,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))

```

# Monitoring program overview by country


```{r monitoringoverviewdas_by_country_program ,cache=TRUE,eval=T,echo=T,include=T}
#moniroting by program
monitoringProgramType_namectry <- D2 %>%
    group_by(monitoringProgramType_name, country_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    group_by(year, country_name) %>%
    mutate(tot = sum(das)) %>%
    ungroup() %>%
    mutate(reldas = das / tot * 100)
#a graph
plt_monitoring_program_country <- ggplot(
    monitoringProgramType_namectry,
    aes(
        x = year,
        y = reldas,
        stratum = monitoringProgramType_name,
        alluvium = monitoringProgramType_name,
        fill = monitoringProgramType_name,
        label = monitoringProgramType_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~country_name) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring program") +
    ylab("Proportion of annual DAS by monitoring program (%)")
textcaption <- "Total monitored (observed) days at sea reported by each country for each monitoring program (2017-2025)"
print(textcaption)
filename <- here("outputs", glue(list_table$plt_monitoring_program_country[1], "monitoring_program_country.png"))
#keep 1.9
print(plt_monitoring_program_country)
png(filename, res = 300, width = 21, height = 16, unit = "cm")
print(plt_monitoring_program_country)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        height = 16,
        width = 21,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered", ) |>
    print(target = sub(".png", ".docx", filename))
```


# Incident by monitoring program

```{r incidents_by_monitoringProgram ,cache=TRUE,eval=T,echo=T,include=T}

pipo <- D3
pipo$nbinc <- apply(
    D3 %>% select(incidentsWithPingers, incidentsWithoutPingers),
    1,
    sum,
    na.rm = T
)
pipo <- pipo %>%
    mutate(classname = ifelse(is.na(classname), "Reptilia", classname)) %>% 
    mutate(
        newclass = ifelse(grepl("Aves|Elasmo|Mamma|Rept", classname), classname, "Other")
    ) 

incprog <- pipo %>%
    group_by(monitoringMethod_name, year, newclass) %>%
    summarise(inc = sum(nbinc, na.rm = T), .groups = "drop") %>%
    ungroup()
#a graph
plt_incident_programm <- ggplot(
    incprog,
    aes(
        x = as.factor(year),
        y = inc,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~newclass, scale = "free") +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "inside",
        legend.position.inside = c(.85, 0.25)
    ) +
    labs(fill = "Monitoring method") +
    ylab("Number of reported incidents") +
    xlab("")
textcaption <- "Total number of bycatch incidents for each taxa (birds, elasmobranchs, mammals, other fish species, and rep-tiles) reported by each monitoring method (2017-2025): at-sea-observers, electronic monitoring, port observers, and vessel crew observers, logbooks, other."
print(textcaption)
filename <- here("outputs", glue(list_table$plt_incident_programm[1], "_incidents_by_monitoringProgram.png"))
#keep 1.10 keep
print(plt_incident_programm)
png(
    filename,
    res = 300,
    width = 16,
    height = 12,
    unit = "cm"
)
print(plt_incident_programm)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        height = 12,
        width = 16,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```


# Monitoring methods overview by Ecoregion


```{r monitoring_method_ecoregion ,cache=TRUE,eval=T,echo=T,include=T}
#prepare data for sample coverage by monitoring effort by gear by georegions
monmethgear <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop")

#monitoring methods: ecoregion/gear view
monmethgear1 <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    filter(!monitoringMethod_name %in% c("Logbook", "Port observers"))
monmethgear2 <- monmethgear1 %>%
    mutate(metierL3_name = "All metier") %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(das = sum(das, na.rm = T), .groups = "drop")
monmethgear <- rbind(monmethgear1, monmethgear2) %>%
    mutate(monitoringMethod_name = as.factor(monitoringMethod_name))


listecoregion <- na.omit(unique(monmethgear$ecoregion))
for (i in listecoregion) {
    tmp <- monmethgear[monmethgear$ecoregion == i, ] %>%
        filter(!is.na(metierL3_name))
    plt <- ggplot(
        tmp,
        aes(
            x = year,
            y = das,
            stratum = monitoringMethod_name,
            alluvium = monitoringMethod_name,
            fill = monitoringMethod_name,
            label = monitoringMethod_name
        )
    ) +
        geom_flow(stat = "alluvium") +
        geom_stratum() +
        facet_wrap(~metierL3_name, scale = "free") +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
        labs(fill = "Monitoring method") +
        ggtitle(i) +
        ylab("Sum of annual DAS by monitoring method")
    print(plt)
    filename <- here("outputs", paste0("monitoring_method_", gsub(" ", "", i), ".png"))
    png(filename, res = 300, width = 1600, height = 1200, unit = "px")
    print(plt)
    dev.off()
}

#prepare data for sample coverage by monitoring effort by georegions
monmethecoregion <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, year) %>%
    summarise(das = sum(daysAtSeaOb, na.rm = T), .groups = "drop") %>%
    filter(!is.na(ecoregion))

plt <- ggplot(
    monmethecoregion %>% mutate(ecoregion = sub(" ", "\n", ecoregion)),
    aes(
        x = year,
        y = das,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~ecoregion, scale = "free") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring method") +
    ggtitle("Monitoring program by ecoregion") +
    ylab("Sum of annual DAS by monitoring method")
textcaption <- "test"
print(textcaption)
filename <- here("outputs", "monitoring_method_ecoregion.png")
print(plt)
png(filename, res = 300, width = 18, height = 14, unit = "cm")
print(plt)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        height = 14,
        width = 16,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```


# Monitoring methods overview trips

By trips : last year figures were wrong (according to the script).

```{r monitoringoverviewtrips,cache=TRUE,eval=T,echo=T,include=T}
# prepare data for fig 3.3 (reported monitoring method)
#all
monmethall <- D2 %>%
    group_by(monitoringMethod_name, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    group_by(year) %>%
    mutate(tot = sum(trips)) %>%
    ungroup() %>%
    mutate(reltrips = trips / tot * 100)
#by ctry
monmethctry <- D2 %>%
    group_by(monitoringMethod_name, country_name, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    group_by(year, country_name) %>%
    mutate(tot = sum(trips)) %>%
    ungroup() %>%
    mutate(reltrips = trips / tot * 100)
#plt output here
#monitoring methods overview
plt <- ggplot(
    monmethall,
    aes(
        x = year,
        y = reltrips,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    theme_bw() +
    labs(fill = "Monitoring method") +
    ylab("Proportion of annual trips by monitoring method (%)")
print(plt)
textcaption <- ""
print(textcaption)
filename <- here("outputs", "monitoring_method_overview_trips.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        width = 16,
        height = 12,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```


# Monitoring methods overview trips and country

```{r monitoringoverviewtrips_country,cache=TRUE,eval=T,echo=T,include=T}
#clean
plt <- ggplot(
    monmethctry,
    aes(
        x = year,
        y = reltrips,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~country_name) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring method") +
    ylab("Proportion of annual trips by monitoring method (%)")
print(plt)
textcaption <- "Total monitored (observed) days at sea reported by each country for each monitoring method (2017-2025); at-sea-observers, electronic monitoring, port observers, vessel crew observers, logbooks, other."
print(textcaption)
filename <- here("outputs", "monitoring_method_country_trips.png")
png(filename, res = 300, width = 30, height = 20, unit = "cm")
print(plt)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        width = 10,
        height = 8,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```


# Monitoring program overview trips and country


```{r monitoring_program_overviewtrips_country_trip,cache=TRUE,eval=T,echo=T,include=T}

#moniroting by program
monprogctry <- D2 %>%
    group_by(monitoringProgramType_name, country_name, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    group_by(year, country_name) %>%
    mutate(tot = sum(trips)) %>%
    ungroup() %>%
    mutate(reltrips = trips / tot * 100)
#a graph
plt <- ggplot(
    monprogctry,
    aes(
        x = year,
        y = reltrips,
        stratum = monitoringProgramType_name,
        alluvium = monitoringProgramType_name,
        fill = monitoringProgramType_name,
        label = monitoringProgramType_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~country_name) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring program") +
    ylab("Proportion of annual trips by monitoring program (%)")
print(plt)
filename <- here("outputs", "monitoring_program_country_trips.png")
textcaption <- "test"
print(textcaption)
png(filename, res = 300, width = 30, height = 20, unit = "cm")
print(plt)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        width = 12,
        height = 10,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```


```{r monitoring_method_ecoregion_trips, cache=TRUE, eval=T, echo=T, include=T}
#prepare data for sample coverage by monitoring effort by gear by georegions
monmethgear <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(trips = sum(daysAtSeaOb, na.rm = T), .groups = "drop")

#monitoring methods: ecoregion/gear view
monmethgear1 <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    filter(!monitoringMethod_name %in% c("Logbook", "Port observers"))
monmethgear2 <- monmethgear1 %>%
    mutate(metierL3_name = "All metier") %>%
    group_by(monitoringMethod_name, ecoregion, metierL3_name, year) %>%
    summarise(trips = sum(trips, na.rm = T), .groups = "drop")
monmethgear <- rbind(monmethgear1, monmethgear2) %>%
    mutate(monitoringMethod_name = as.factor(monitoringMethod_name))


listecoregion <- na.omit(unique(monmethgear$ecoregion))
for (i in listecoregion) {
    tmp <- monmethgear[monmethgear$ecoregion == i, ] %>%
        filter(!is.na(metierL3_name))
    plt <- ggplot(
        tmp,
        aes(
            x = year,
            y = trips,
            stratum = monitoringMethod_name,
            alluvium = monitoringMethod_name,
            fill = monitoringMethod_name,
            label = monitoringMethod_name
        )
    ) +
        geom_flow(stat = "alluvium") +
        geom_stratum() +
        facet_wrap(~metierL3_name, scale = "free") +
        theme_bw() +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
        labs(fill = "Monitoring method") +
        ggtitle(i) +
        ylab("Sum of annual trips by monitoring method")
    print(plt)
    filename <- here(
        "outputs",
        paste0("monitoring_method_", gsub(" ", "", i), "_trips.png")
    )
    png(filename, res = 300, width = 1600, height = 1200, unit = "px")
    print(plt)
    dev.off()
}

#prepare data for sample coverage by monitoring effort by georegions
monmethecoregion <- D2 %>%
    group_by(monitoringMethod_name, ecoregion, year) %>%
    summarise(trips = sum(tripsOb, na.rm = T), .groups = "drop") %>%
    filter(!is.na(ecoregion))

plt <- ggplot(
    monmethecoregion %>% mutate(ecoregion = sub(" ", "\n", ecoregion)),
    aes(
        x = year,
        y = trips,
        stratum = monitoringMethod_name,
        alluvium = monitoringMethod_name,
        fill = monitoringMethod_name,
        label = monitoringMethod_name
    )
) +
    geom_flow(stat = "alluvium") +
    geom_stratum() +
    facet_wrap(~ecoregion, scale = "free") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Monitoring method") +
    ggtitle(i) +
    ylab("Sum of annual trips by monitoring method")
print(plt)
textcaption <- "test"
print(textcaption)
filename <- here("outputs", "monitoring_method_ecoregion_trips.png")
png(filename, res = 300, width = 30, height = 20, unit = "cm")
print(plt)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        width = 12,
        height = 10,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```

# EM figures 

```{r, em_total_das, cache=TRUE, eval=T, echo=T, include=T}
#Use only Electronic Monitoring
EM_monitoringeffort <- D2 %>%
    filter(monitoringMethod == "EM")
## Pivot the info
EM_Grouped <- EM_monitoringeffort %>%
    group_by(year, country) %>%
    summarize(
        total_tripsOb = sum(tripsOb, na.rm = TRUE),
        total_daysAtSeaOb = sum(daysAtSeaOb, na.rm = TRUE)
    )

## 4.4 PLOT EM total trips observed =======
plt_em_total_das_trip <- ggplot(EM_Grouped, aes(x = year, y = total_daysAtSeaOb, fill = country)) +
    geom_bar(stat = "identity", position = "stack", colour = "black") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    scale_x_continuous(breaks = unique(EM_Grouped$year)) +
    labs(fill = "Country") +
    xlab('') +
    ylab("Total DAS observed using EM")
print(plt_em_total_das_trip)
textcaption <- "Total monitored (observed) number of days at sea reported by each country with electronic monitoring."
print(textcaption)
filename <- here("outputs", glue(list_table$plt_em_total_das_trip[1], "_em_total_das.png"))
png(filename, res = 300, width = 8, height = 10, unit = "cm")
print(plt_em_total_das_trip)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 8,
        height = 10,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))

plt_B <- ggplot(EM_Grouped, aes(x = year, y = total_tripsOb, fill = country)) +
    geom_bar(stat = "identity", position = "stack", colour = "black") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    scale_x_continuous(breaks = unique(EM_Grouped$year)) +
    labs(fill = "Country") +
    xlab('') +
    ylab("Total trips observed using EM")
print(plt_B)
textcaption <- "Total monitored (observed) number of trip reported by each country with electronic monitoring."
print(textcaption)
filename <- here("outputs", "em_total_das_trip.png")
png(filename, res = 300, width = 30, height = 20, unit = "cm")
print(plt_B)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 8,
        height = 10,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))

AB <- plot_grid(
    plt_em_total_das_trip + theme(legend.position = "none"),
    plt_B + theme(legend.position = "right"),
    rel_widths = c(1, 1.2),
    nrow = 1
)
textcaption <- "Total monitored (observed) number of trip reported by each country with electronic monitoring."
print(textcaption)
filename <- here("outputs", "em_total_das_and_trip.png")
png(filename, res = 300, width = 30, height = 20, unit = "cm")
print(AB)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 12,
        height = 10,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```
```{r, em_plt_incident, cache=TRUE, eval=T, echo=T, include=T}
EM_taxa_bycatch <- D3 %>%
    filter(monitoringMethod == "EM")

# Drop na values and drop teleostei (Because only 3 incidents! Check this later if needed!)
EM_taxa_bycatch <- EM_taxa_bycatch %>%
    filter(!is.na(classname) & classname != "Teleostei") %>%
    filter(!classname == "Chondrostei")

## Pivot the info
EM_Taxa_Grouped <- EM_taxa_bycatch %>%
    group_by(year, country, classname) %>%
    summarize(
        sum_incidents = sum(incidentsWithPingers, na.rm = TRUE) +
            sum(incidentsWithoutPingers, na.rm = TRUE),
        sum_individuals = sum(individualsWithPingers, na.rm = TRUE) +
            sum(individualsWithoutPingers, na.rm = TRUE)
    )


#Plot incidents per taxa
plt_em_plt_incidents <- ggplot(EM_Taxa_Grouped, aes(x = year, y = sum_incidents, fill = country)) +
    geom_bar(stat = "identity", position = "stack", colour = "black") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    scale_x_continuous(breaks = unique(EM_Taxa_Grouped$year)) +
    labs(fill = "Country") +
    ylab("Total number of incidents") +
    xlab('') +
    facet_wrap(~classname, scales = "free", nrow = 1)

textcaption <- "Total number of bycatch incidents reported by each country through the WGBYC data calls, observed with electronic monitoring. Sweden did not report elasmobranch bycatch with EM in 2022-2025 due to difficulties in taxo-nomic identification. France did not report bycatch of seabirds and elasmobranchs with EM because data collection is not authorized for these taxa unless specified by the crew. Spain has ongoing EM research programs but did not submit EM data through the WGBYC data calls. "
print(textcaption)
filename <- here("outputs", glue(list_table$plt_em_plt_incidents[1], "_em_plt_incidents.png"))
png(filename, res = 300, width = 30, height = 20, unit = "cm")
print(plt_em_plt_incidents)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 12,
        height = 10,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```


```{r EMfig_country,cache=TRUE,eval=T,echo=T,include=T}

#### calculate percentages for text ####

#"Along with at-sea observers, this method yet represents the source of a majority of bycatch
#records reported (Figure 5), with 40% of reported mammal and 30% of seabird bycatch incidents recorded with EM.

## Pivot the info
NofIncidents <- D3 %>%
    group_by(year, monitoringMethod_name, classname) %>%
    summarize(
        sum_incidents = sum(incidentsWithPingers, na.rm = TRUE) +
            sum(incidentsWithoutPingers, na.rm = TRUE),
    )

#Use only Electronic Monitoring
Aves <- NofIncidents %>%
    filter(classname == "Aves")

Aves <- Aves %>%
    group_by(year, monitoringMethod_name) %>%
    summarise(sum_incidents = sum(sum_incidents)) %>%
    mutate(total_incidents_per_year = sum(sum_incidents, na.rm = TRUE)) %>%
    mutate(percentage = (sum_incidents / total_incidents_per_year) * 100)

Aves

#Use only Electronic Monitoring
Mammalia <- NofIncidents %>%
    filter(classname == "Mammalia")

Mammalia <- Mammalia %>%
    group_by(year, monitoringMethod_name) %>%
    summarise(sum_incidents = sum(sum_incidents)) %>%
    mutate(total_incidents_per_year = sum(sum_incidents, na.rm = TRUE)) %>%
    mutate(percentage = (sum_incidents / total_incidents_per_year) * 100)

Mammalia

#### Trips per country per year ####

#Use only Electronic Monitoring
EM_monitoringeffort <- D2 %>%
    filter(monitoringMethod == "EM")
## Pivot the info
EM_Grouped <- EM_monitoringeffort %>%
    group_by(year, country_name) %>%
    summarize(total_tripsOb = sum(tripsOb, na.rm = TRUE))

# PLOT EM total trips observed
plt <- ggplot(EM_Grouped, aes(x = year, y = total_tripsOb, fill = country_name)) +
    geom_bar(stat = "identity", position = "stack") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Country") +
    ylab("Total trips observed using EM")

print(plt)
filename <- here("outputs", "EM_trips_obs_country.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()

#### N of incidents per year per taxa, stack countries ####

#Use only Electronic Monitoring
EM_taxa_bycatch <- D3 %>%
    filter(monitoringMethod == "EM")

# Drop na values and drop teleostei (Because only 3 incidents! Check this later if needed!)
EM_taxa_bycatch <- EM_taxa_bycatch %>%
    filter(!is.na(classname) & classname != "Teleostei")

## Pivot the info
EM_Taxa_Grouped <- EM_taxa_bycatch %>%
    group_by(year, country_name, classname) %>%
    summarize(
        sum_incidents = sum(incidentsWithPingers, na.rm = TRUE) +
            sum(incidentsWithoutPingers, na.rm = TRUE),
        sum_individuals = sum(individualsWithPingers, na.rm = TRUE) +
            sum(individualsWithoutPingers, na.rm = TRUE)
    )

#Plot individuals per taxa
plt <- ggplot(EM_Taxa_Grouped, aes(x = year, y = sum_individuals, fill = country_name)) +
    geom_bar(stat = "identity", position = "stack") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Country") +
    ylab("Total number of individuals") +
    facet_wrap(~classname, scales = "free") +
    scale_x_continuous(limits = c(2018.5, max(EM_Taxa_Grouped$year + 0.5))) #this is just to make all x-axes the same

print(plt)
textcaption <- "test"
print(textcaption)
filename <- here("outputs", "EM_individuals_country.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        width = 12,
        height = 10,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```


```{r EMfig_incident_country,cache=TRUE,eval=T,echo=T,include=T}

#Plot incidents per taxa
plt <- ggplot(EM_Taxa_Grouped, aes(x = year, y = sum_incidents, fill = country_name)) +
    geom_bar(stat = "identity", position = "stack") +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    labs(fill = "Country") +
    ylab("Total number of incidents") +
    facet_wrap(~classname, scales = "free") +
    scale_x_continuous(limits = c(2018.5, max(EM_Taxa_Grouped$year + 0.5))) #this is just to make all x-axes the same
plot(plt)
print(plt)
textcaption <- "test"
print(textcaption)
filename <- here("outputs", "EM_incidents_country.png")
png(filename, res = 300, width = 1600, height = 1200, unit = "px")
print(plt)
dev.off()
#save as docx object
read_docx() |>
    body_add_img(
        src = filename,
        width = 12,
        height = 10,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```

# Long tables production 

Text

```{r longtab,cache=TRUE,eval=T,echo=T,include=T}
#long-table-container style="max-height: 500px; overflow-y: auto;"

#big long useless table
#Annex_table
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, areaCode, metierL3, monitoringMethod) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3, areaCode) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, areaCode, metierL3, monitoringMethod, species, classname) %>%
    summarise(
        nb = sum(individualsWithPingers, individualsWithoutPingers, na.rm = T),
        inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)
    ) %>%
    ungroup()

tab123 <- left_join(left_join(tab2, tab1), tab3) %>%
    filter(!is.na(species)) %>%
    transmute(
        ecoregion,
        areaCode,
        metierL3,
        dastot,
        monitoringMethod,
        dasobs,
        moncov = 100 * (dasobs / dastot),
        classname,
        species,
        nb,
        inc
    ) %>%
    mutate(dasobs = ifelse(dasobs == 0, NA, dasobs)) %>%
    mutate(moncov = ifelse(dasobs == 0, NA, moncov)) %>%
    transmute(
        Ecoregion = ecoregion,
        AreaCode = areaCode,
        MetierL3 = metierL3,
        `Fishing Effort (das)` = dastot,
        MonitoringMethod = monitoringMethod,
        `Total Observed Effort (das)` = dasobs,
        `Monitoring Coverage (%)` = moncov,
        classname,
        Species = species,
        Total_No_Specimens = nb,
        Incidents = inc
    )

head(tab123)

#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(
    wb,
    file = here("outputs", "TOR_A_long_table_bycatch_only.xlsx"),
    overwrite = T
)

#big long useless table with all info
#Annex_table
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, areaCode, metierL3, monitoringMethod) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3, areaCode) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, areaCode, metierL3, monitoringMethod, species, classname) %>%
    summarise(
        nb = sum(individualsWithPingers, individualsWithoutPingers, na.rm = T),
        inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)
    ) %>%
    ungroup()

tab123 <- full_join(full_join(tab2, tab1), tab3) %>% #filter(!is.na(species))%>%
    transmute(
        ecoregion,
        areaCode,
        metierL3,
        dastot,
        monitoringMethod,
        dasobs,
        moncov = 100 * (dasobs / dastot),
        classname,
        species,
        nb,
        inc
    ) %>%
    mutate(dasobs = ifelse(dasobs == 0, NA, dasobs)) %>%
    mutate(moncov = ifelse(dasobs == 0, NA, moncov)) %>%
    transmute(
        Ecoregion = ecoregion,
        AreaCode = areaCode,
        MetierL3 = metierL3,
        `Fishing Effort (das)` = dastot,
        MonitoringMethod = monitoringMethod,
        `Total Observed Effort (das)` = dasobs,
        `Monitoring Coverage (%)` = moncov,
        classname,
        Species = species,
        Total_No_Specimens = nb,
        Incidents = inc
    )

#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(wb, file = here("outputs", "TOR_A_long_table.xlsx"), overwrite = T)

#short table
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3, monitoringMethod) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3, monitoringMethod, species, classname) %>%
    summarise(
        nb = sum(individualsWithPingers, individualsWithoutPingers, na.rm = T),
        inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)
    ) %>%
    ungroup()

tab123 <- left_join(left_join(tab2, tab1), tab3) %>%
    filter(!is.na(species)) %>%
    transmute(
        ecoregion,
        metierL3,
        dastot,
        monitoringMethod,
        dasobs,
        moncov = 100 * (dasobs / dastot),
        classname,
        species,
        nb,
        inc
    ) %>%
    mutate(dasobs = ifelse(dasobs == 0, NA, dasobs)) %>%
    mutate(moncov = ifelse(dasobs == 0, NA, moncov)) %>%
    transmute(
        Ecoregion = ecoregion,
        MetierL3 = metierL3,
        `Fishing Effort (das)` = dastot,
        MonitoringMethod = monitoringMethod,
        `Total Observed Effort (das)` = dasobs,
        `Monitoring Coverage (%)` = moncov,
        classname,
        Species = species,
        Total_No_Specimens = nb,
        Incidents = inc
    )

#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(wb, file = here("outputs", "TOR_A_short_table.xlsx"), overwrite = T)


#shorter table
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    mutate(classname = ifelse(is.na(classname), "Reptiles", classname)) %>%
    group_by(Ecoregion = ecoregion, classname) %>%
    summarise(
        nbspp = n_distinct(species),
        nb = sum(individualsWithPingers, individualsWithoutPingers, na.rm = T),
        inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)
    ) %>%
    ungroup() %>%
    tidyr::pivot_longer(nbspp:inc, names_to = "variable") %>%
    tidyr::pivot_wider(values_from = value, names_from = classname)
tab3$variable[tab3$variable == "nbspp"] <- "Species"
tab3$variable[tab3$variable == "nb"] <- "Individuals"
tab3$variable[tab3$variable == "inc"] <- "Incidents"


tab123 <- left_join(tab1, tab2) %>%
    filter(!is.na(ecoregion)) %>%
    transmute(
        Ecoregion = ecoregion,
        `Fishing Effort (das)` = dastot,
        `Total Observed Effort (das)` = dasobs,
        `Monitoring Coverage (%)` = 100 * (dasobs / dastot)
    ) %>%
    left_join(tab3) %>%
    filter(Ecoregion != "")


#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(wb, file = here("outputs", "TOR_A_shorter_table.xlsx"), overwrite = T)

pipo <- flextable(tab123) %>%
    merge_v(
        j = ~ Ecoregion +
            `Fishing Effort (das)` +
            `Total Observed Effort (das)` +
            `Monitoring Coverage (%)`
    ) %>%
    hline(i = seq(3, nrow(tab123), 3)) %>%
    valign(j = 1:4, valign = "top") %>%
    colformat_double(j = 4, digits = 2) %>%
    colformat_double(j = c(2:3, 5:ncol(tab123)), digits = 0) %>%
    bg(part = "header", bg = "grey30") %>%
    color(part = "header", color = "ghostwhite") %>%
    italic(j = 1) %>%
    fit_to_width(max_width = 11.7)
pipo

#save as docx object
save_as_docx(
    pipo,
    path = here("outputs", "TOR_A_shorter_table.docx"),
    pr_section = prop_section(
        page_size = page_size(orient = "landscape"),
        page_margins = page_mar(
            bottom = 0,
            top = 0,
            right = 0,
            left = 0,
            header = 0,
            footer = 0,
            gutter = 0
        )
    )
)


#shorter table detailed effort
tab2 <- D2 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3_name) %>%
    summarise(dasobs = sum(daysAtSeaOb, na.rm = T)) %>%
    ungroup()
tab1 <- D1 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3_name) %>%
    summarise(dastot = sum(daysAtSeaF, na.rm = T)) %>%
    ungroup()
tab12 <- full_join(tab1, tab2) %>%
    mutate(moncov = 100 * (dasobs / dastot)) %>%
    mutate(moncov = ifelse(is.finite(moncov), moncov, 0)) %>%
    mutate(info = paste0(round(dastot, 0), " das (", round(moncov, 3), "% cov.)")) %>%
    select(ecoregion, metierL3_name, info)
tab3 <- D3 %>%
    filter(year == currentyear) %>%
    group_by(ecoregion, metierL3_name) %>%
    summarise(inc = sum(incidentsWithPingers, incidentsWithoutPingers, na.rm = T)) %>%
    mutate(inc = ifelse(is.finite(inc), inc, 0)) %>%
    ungroup()
tab123 <- full_join(tab12, tab3) %>%
    mutate(inc = ifelse(is.finite(inc), inc, 0)) %>%
    mutate(info = paste0(info, " das/ ", inc, " inc.")) %>%
    select(ecoregion, metierL3_name, info) %>%
    pivot_wider(values_from = info, names_from = metierL3_name) %>%
    filter(!is.na(ecoregion))


flextable(tab123) %>%
    autofit() %>%
    set_table_properties(
        opts_html = list(
            scroll = list(
                height = "400px", # Limits height and enables vertical scroll
                freeze_first_column = TRUE # Keeps first column visible when scrolling right
            )
        )
    )
#write excel file
wb <- createWorkbook()
addWorksheet(wb, 1)
writeData(wb, 1, tab123)
saveWorkbook(
    wb,
    file = here("output", "TOR_A_cool_effort_shorter_table.xlsx"),
    overwrite = T
)
```

# Effort Monitoring

## Select parameters

```{r affort_monitoring_parameters, cache=TRUE, eval=T, echo=T, include=T}

# Select year
analysis_year = 2025

# Rename vector
rename_vector <- c(
    # L4 level
    "DRB" = "DRB - Boat dredge",
    "FPN" = "FPN - Stationary uncovered pound nets",
    "FPO" = "FPO - Pots and traps",
    "FWR" = "FWR - Fixed installations for fences and weirs",
    "FYK" = "FYK - Fyke nets",
    "GEL" = "GEL - Glass eel fishing",
    "GN" = "GN - Gill nets",
    "GNC" = "GNC - Encircling gillnets",
    "GND" = "GND - Driftnet",
    "GNS" = "GNS - Set gillnet",
    "GTN" = "GTN - Gillnets-trammel nets",
    "GTR" = "GTR - Trammel nets",
    "HMD" = "HMD - Mechanised/suction dredge",
    "LA" = "LA - Lampara nets",
    "LH" = "LH - Hand and pole lines",
    "LHM" = "LHM - Hand and pole lines",
    "LHP" = "LHP - Hand and pole lines (hand-operated)",
    "LLD" = "LLD - Drifting longlines",
    "LLS" = "LLS - Set longlines",
    "LTL" = "LTL - Trolling lines",
    "LX" = "LX - Hooks and lines",
    "MIS" = "MIS - Gear nei.",
    "OTB" = "OTB - Bottom otter trawl",
    "OTM" = "OTM - Midwater otter trawl",
    "OTT" = "OTT - Milti-rig otter trawl",
    "PS" = "PS - Purse seine",
    "PTB" = "PTB - Bottom pair trawl",
    "PTM" = "PTM - Pelagic pair trawl",
    "SB" = "SB - Beach Seine",
    "SBV" = "SBV - Beach and boat seine",
    "SDN" = "SDN - Anchored seine",
    "SPR" = "SPR - Pair seine",
    "SSC" = "SSC - Fly shooting seine",
    "TBB" = "TBB - Beam trawl",
    # L3 level
    "L3DR" = "DR - Dredges",
    "L3FIX" = "Fix - Traps",
    "L3GN" = "GN - Nets",
    "L3LH" = "LH - Rods and Lines",
    "L3LL" = "LL - Longlines",
    "L3PS" = "PS - Surrounding Nets",
    "L3PT" = "PT - Pelagic Trawls",
    "L3SX" = "SX - Seines",
    "L3TB" = "TB - Bottom Trawls"
)

# Do we include NAFO or not
include_NAFO_areas <- FALSE

## Prepare Map
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf") %>%
    st_make_valid() %>%
    st_transform(crs = 3035)

# ICES area and GFCM
# The shape file come from Laurent Dubroca, we need to found a wya to source it correctly
ices_gfcm <- readRDS(here("data", "vocab", "allgeo.rds"))
ices_gfcm <- ices_gfcm %>% filter(type %in% c("ices_subdivision", "ices_division"))

# NAFO
if (include_NAFO_areas) {
    NAFOareas <- st_read("../../ICES_areas_map/NAFO_Divisions.shp") %>%
        st_transform(crs = st_crs(ices_gfcm))
    combined_areas <- bind_rows(
        combined_areas,
        NAFOareas %>%
            select(geometry, icesdiv = Label) %>%
            mutate(
                Area_Full = paste0(
                    "21.",
                    str_replace(as.character(Label), "([0-9])([A-Z])", "\\1.\\2")
                )
            )
    )
} else {
    combined_areas <- ices_gfcm
}

# ransform shapefile =====
ICESareas_laea <- st_transform(combined_areas, crs = 3035)
```


## Data processing

```{r effort_monitoring_prepare_data,cache=TRUE,eval=T,echo=T,include=T}
# 3. Data processing D1 =====================================================================

## 3.1 Data D1 =============================
D1_data <- D1 %>%
    filter(!areaCode %in% c("27.1", "27.6", "27.1~27.2.a~27.2.b", "21.1")) %>%
    mutate(
        areaCode = recode(
            areaCode,
            "27.3.a" = "27.3.a.20",
            "27.5.a" = "27.5.a.2",
            "27.3.d.28" = "27.3.d.28.2",
            "11.1" = "111",
            "11.2" = "112",
            "21.3.l" = "21.3.L",
            "21.3.m" = "21.3.M",
            "21.3.n" = "21.3.N",
            "21.3.k" = "21.3.K",
            "21.3.o" = "21.3.O",
            "21.4.V.s" = "21.4.Vs"
        )
    )
if (include_NAFO_areas) {
    D1_data <- D1_data %>% filter(areaType != "NAFOArea")
}


## 3.2 function to process data ==========================================
process_data <- function(data, group_var, rename_vector) {
    data %>%
        filter(year == analysis_year) %>%
        group_by(!!sym(group_var), areaCode) %>%
        summarize(total_daysAtSeaF = sum(daysAtSeaF, na.rm = TRUE)) %>%
        left_join(ICESareas_laea, by = c("areaCode" = "offname")) %>%
        mutate(
            daysAtSeaF_cat = cut(
                total_daysAtSeaF,
                breaks = c(-Inf, 100, 1000, 2000, 5000, 10000, 25000, Inf),
                labels = c(
                    "< 100",
                    "100-1000",
                    "1000-2000",
                    "2000-5000",
                    "5000-10000",
                    "10000-25000",
                    "> 25000"
                )
            ),
            !!sym(group_var) := recode(!!sym(group_var), !!!rename_vector)
        ) %>%
        filter(!is.na(daysAtSeaF_cat) & !!sym(group_var) != "L3NK")
}


## 3.3 process L3 and L4 data ==========================================
daysAtSea_spatial <- process_data(D1_data, "metierL3", rename_vector)
daysAtSea_L4_spatial <- process_data(D1_data, "metierL4", rename_vector)

hist(
    daysAtSea_spatial$total_daysAtSeaF,
    breaks = seq(from = 0, to = 1000000, by = 100),
    xlim = c(0, 10000)
)
hist(
    daysAtSea_L4_spatial$total_daysAtSeaF,
    breaks = seq(from = 0, to = 1000000, by = 100),
    xlim = c(0, 10000)
)


# 3. Data processing D2 =====================================================================

## 3.1 Data D2 ==================
D2_data <- D2 %>%
    filter(!areaCode %in% c("27.1", "27.6", "27.1~27.2.a~27.2.b", "21.1")) %>%
    mutate(
        areaCode = recode(
            areaCode,
            "27.3.a" = "27.3.a.20",
            "27.5.a" = "27.5.a.2",
            "27.3.d.28" = "27.3.d.28.2",
            "11.1" = "111",
            "11.2" = "112",
            "21.3.l" = "21.3.L",
            "21.3.m" = "21.3.M",
            "21.3.n" = "21.3.N",
            "21.3.k" = "21.3.K",
            "21.3.o" = "21.3.O",
            "21.4.V.s" = "21.4.Vs"
        )
    )

## exclude all log books (LB), port observers (PO), and all log books (except for Norway)
indRef <- which(
    D2_data$monitoringMethod == 'SO' | #all observer data
        D2_data$monitoringMethod == 'VO' & D2_data$country != 'EE' | #include vessel crew except Estonia
        D2_data$monitoringMethod == 'EM' | #all electronic monitoring data
        D2_data$monitoringMethod == 'OTH' & D2_data$country == 'NO'
) #include other only for Norway

D2_data <- D2_data[indRef, ]
rm(indRef)

## 3.2 Filtert Data
daysAtSea_Ob <- D2_data %>%
    filter(year == analysis_year) %>%
    group_by(metierL3, areaCode) %>%
    summarize(daysAtSea_Ob = sum(daysAtSeaOb, na.rm = TRUE)) %>%
    left_join(ICESareas_laea, by = c("areaCode" = "offname"))


## 3.2 assign breaks for plotting and sort data accordingly ====
daysAtSea_Ob <- daysAtSea_Ob %>%
    mutate(
        daysAtSea_Ob_cat = cut(
            daysAtSea_Ob,
            breaks = c(-Inf, 5, 10, 20, 50, 100, 500, Inf),
            labels = c("<5", "5-10", "10-20", "20-50", "50-100", "100-500", ">500")
        ),
        metierL3 = recode(metierL3, !!!rename_vector)
    ) %>%
    filter(!is.na(daysAtSea_Ob_cat) & metierL3 != "L3NK")


## 3.3 monitoring effort Metier L4 ==============
daysAtSea_L4_Ob <- D2_data %>%
    filter(year == analysis_year) %>%
    group_by(metierL4, areaCode) %>%
    summarize(daysAtSea_Ob = sum(daysAtSeaOb, na.rm = TRUE)) %>%
    left_join(ICESareas_laea, by = c("areaCode" = "offname")) %>%
    mutate(
        daysAtSea_L4_Ob_cat = cut(
            daysAtSea_Ob,
            breaks = c(-Inf, 5, 10, 20, 50, 100, 500, Inf),
            labels = c("<5", "5-10", "10-20", "20-50", "50-100", "100-500", ">500")
        ),
        metierL4 = recode(metierL4, !!!rename_vector)
    ) %>%
    filter(!is.na(daysAtSea_L4_Ob_cat) & metierL4 != "L4NK")


## 3.4 observer coverage L3 =========================
matchedDaysAtSea <- full_join(
    daysAtSea_spatial,
    daysAtSea_Ob,
    by = c("areaCode", "metierL3")
)
hist(
    (matchedDaysAtSea$daysAtSea_Ob / matchedDaysAtSea$total_daysAtSeaF) * 100,
    breaks = seq(from = 0, to = 100, by = 0.1),
    xlim = c(0, 2)
)

spatialCoverage <- left_join(
    ICESareas_laea,
    matchedDaysAtSea,
    by = c("offname" = "areaCode")
) %>%
    mutate(
        Obs_Coverage = (daysAtSea_Ob / total_daysAtSeaF) * 100,
        Obs_Coverage_cat = cut(
            Obs_Coverage,
            breaks = c(-Inf, 0.25, 0.5, 0.75, 1, 2, 4, Inf),
            labels = c(
                "< 0.25",
                "0.25-0.5",
                "0.5-0.75",
                "0.75-1.0",
                "1.0-2.0",
                "2.0-4.0",
                "> 4.0"
            )
        )
    ) %>%
    filter(!is.na(Obs_Coverage_cat) & metierL3 != "L3NK") %>%
    mutate(metierL3 = recode(metierL3, !!!rename_vector))


## 3.5 Observer coverage L4 ==================================
matchedDaysAtSea_L4 <- full_join(
    daysAtSea_L4_spatial,
    daysAtSea_L4_Ob,
    by = c("areaCode", "metierL4")
)
spatialCoverage_L4 <- left_join(
    ICESareas_laea,
    matchedDaysAtSea_L4,
    by = c("offname" = "areaCode")
) %>%
    mutate(
        Obs_Coverage = (daysAtSea_Ob / total_daysAtSeaF) * 100,
        Obs_Coverage_cat = cut(
            Obs_Coverage,
            breaks = c(-Inf, 0.1, 0.25, 0.5, 0.75, 1, 2, Inf),
            labels = c(
                "< 0.1",
                "0.1-0.25",
                "0.25-0.5",
                "0.5-0.75",
                "0.75-1.0",
                "1.0-2.0",
                "> 2.0"
            )
        )
    ) %>%
    filter(!is.na(Obs_Coverage_cat) & metierL4 != "L4NK") %>%
    mutate(metierL4 = recode(metierL4, !!!rename_vector))
```

## Plot the map

```{r effort_monitoring_plot_map,cache=TRUE,eval=T,echo=T,include=T}

## color palette options to avoid a highly controversial continuous log scale
c("#FFCCCC", "#FF9999", "#FF6666", "#FF3333", "#FF0000", "#CC0000") # old and ugly

c(
    "#440154FF",
    "#46337EFF",
    "#365C8DFF",
    "#277F8EFF",
    "#1FA187FF",
    "#4AC16DFF",
    "#9FDA3AFF",
    "#FDE725FF"
) # 8 levels

c(
    "#440154FF",
    "#443A83FF",
    "#31688EFF",
    "#21908CFF",
    "#35B779FF",
    "#8FD744FF",
    "#FDE725FF"
) # 7 levels

c("#440154FF", "#414487FF", "#2A788EFF", "#22A884FF", "#7AD151FF", "#FDE725FF") # 6 levels

## define color palette for plot with shades of red
ownPalette <- list(
    fishingEffort = c(
        "< 100" = "#440154FF",
        "100-1000" = "#443A83FF",
        "1000-2000" = "#31688EFF",
        "2000-5000" = "#21908CFF",
        "5000-10000" = "#35B779FF",
        "10000-25000" = "#8FD744FF",
        "> 25000" = "#FDE725FF"
    ),
    observer = c(
        "<5" = "#440154FF",
        "5-10" = "#443A83FF",
        "10-20" = "#31688EFF",
        "20-50" = "#21908CFF",
        "50-100" = "#35B779FF",
        "100-500" = "#8FD744FF",
        ">500" = "#FDE725FF"
    ),
    observerPerc = c(
        "< 0.25" = "#440154FF",
        "0.25-0.5" = "#443A83FF",
        "0.5-0.75" = "#31688EFF",
        "0.75-1.0" = "#21908CFF",
        "1.0-2.0" = "#35B779FF",
        "2.0-4.0" = "#8FD744FF",
        "> 4.0" = "#FDE725FF"
    )
)


# Function to create plots =================
create_plot <- function(
    data = XXXX,
    fill_var = "xxxx",
    plotTitle = 'xxxx',
    paletteBreaksColours = ownPalette$xxxx,
    paletteTitle = 'xxxx',
    facet_var = "xxxx",
    facet_ncol = 3,
    facet_nrow = 3
) {
    total_pages <- ceiling(length(unique(data[[facet_var]])) / (facet_ncol * facet_nrow))
    for (page in 1:total_pages) {
        plot <- ggplot(data = data) +
            geom_sf(data = ICESareas_laea, fill = "white", color = "lightgray") +
            geom_sf(
                aes_string(geometry = "geometry", fill = fill_var),
                color = "darkgrey"
            ) +
          geom_sf(data = world, fill = "lightgray", color = "darkgrey", size = 0.1) +

            # scale_fill_viridis(name = paletteBreaksTitle, direction = 1, option = "D",
            #                     trans = "pseudo_log", breaks = paletteBreaks,
            #                    labels = c(paletteBreakLabels),
            #                    na.value = "white",
            #                     discrete = FALSE,
            #                    guide = guide_colourbar(title.position = "top", barwidth = 18, barheight = 1)) +
            scale_fill_manual(
                name = paletteTitle,
                values = paletteBreaksColours,
                na.value = "white",
                guide = guide_legend(
                    title.position = "top",
                    label.position = "bottom",
                    nrow = 1,
                    byrow = TRUE
                )
            ) +
            coord_sf(
                crs = st_crs(3035),
                xlim = c(-40000, 6800000),
                ylim = c(1100000, 7000000)
            ) +
            theme_minimal() +
            labs(title = plotTitle) +
            facet_wrap_paginate(
                as.formula(paste("~", facet_var)),
                ncol = facet_ncol,
                nrow = facet_nrow,
                page = page,
                drop = TRUE
            ) +
            theme(
                strip.text = element_text(size = 6),
                panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                axis.text = element_blank(),
                axis.ticks = element_blank(),
                axis.title = element_blank(),
                legend.position = "bottom",
                legend.text = element_text(
                    size = 8,
                    angle = -67,
                    vjust = 0.33,
                    hjust = 0
                ),
                legend.title = element_text(size = 10),
                panel.background = element_blank(),
                plot.background = element_blank()
            )
        return(plot)
    }
}


## 4.3 Plots for Metier Level 3 ===================

## days at sea
plt_fishing_effort <- create_plot(
    data = daysAtSea_spatial,
    fill_var = "daysAtSeaF_cat",
    plotTitle = 'Fishing effort',
    paletteBreaksColours = ownPalette$fishingEffort,
    paletteTitle = 'Fishing effort (days at sea)',
    facet_var = "metierL3",
    facet_ncol = 3,
    facet_nrow = 3
)
print(plt_fishing_effort)
textcaption <- "Metier level 3 fishing effort for 2025 (days at sea) submitted to the WGBYC database for monitoring methods retained for use under ToR C for the calculation of bycatch estimates for 2025 This includes all scientific observer data, all electronic mon-itoring data, and all vessel crew data (except for Estonia) but excludes reports from logbooks and port observers. Transparent = no data or zero."
print(textcaption)
filename <- here("outputs", glue(list_table$plt_fishing_effort[1], "_fishingEffort_L3.png"))
png(filename, res = 300, width = 30, height = 20, unit = "cm")
print(plt_fishing_effort)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 18,
        height = 14,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))


## monitoring effort days at sea
plt_monitoring_effort <- create_plot(
    data = daysAtSea_Ob,
    fill_var = "daysAtSea_Ob_cat",
    plotTitle = 'Monitoring effort',
    paletteBreaksColours = ownPalette$observer,
    paletteTitle = 'Monitoring effort (days at sea)',
    facet_var = "metierL3",
    facet_ncol = 3,
    facet_nrow = 3
)
print(plt_monitoring_effort)
textcaption <- "Metier level 3 monitoring effort for 2025 (days at sea) submitted to the WGBYC database for monitoring methods retained for use under ToR C for the calculation of bycatch estimates for 2025 This includes all scientific observer data, all elec-tronic monitoring data, and all vessel crew data (except for Estonia) but excludes reports from logbooks and port observers. Transparent = no data or zero."
print(textcaption)
filename <- here("outputs", glue(list_table$plt_monitoring_effort[1], "_monitoringEffort_L3.png"))
png(filename, res = 300, width = 30, height = 20, unit = "cm")
print(plt_monitoring_effort)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 18,
        height = 14,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))


## monitoring coverage (%)
plt_monitoring_coverage <- create_plot(
    data = spatialCoverage,
    fill_var = "Obs_Coverage_cat",
    plotTitle = 'Monitoring coverage',
    paletteBreaksColours = ownPalette$observerPerc,
    paletteTitle = 'Monitoring coverage (%)',
    facet_var = "metierL3",
    facet_ncol = 3,
    facet_nrow = 3
)
print(plt_monitoring_coverage)
textcaption <- "Metier level 3 monitoring coverage of 2025 data submitted to the WGBYC database for monitoring methods retained for use under ToR C for the calculation of bycatch estimates for 2025 This includes all scientific observer data, all electronic mon-itoring data, and all vessel crew data (except for Estonia) but excludes reports from logbooks and port observers. Transparent = no data or zero. Coverage calculated as monitoring effort as a percentage of fishing effort (using numbers presented in Figures X3 & X4 above). Transparent = no data or zero."
print(textcaption)
filename <- here("outputs", glue(list_table$plt_monitoring_coverage[1], "_monitoringCoverage_L3.png"))
png(filename, res = 300, width = 30, height = 20, unit = "cm")
print(plt_monitoring_coverage)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 18,
        height = 14,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```


# Bycatch reporting 

## Prepare data

```{r bycatch_reporting_preparation, cache=TRUE,eval=T,echo=T,include=T}
# Select parameters =======================================

ecoregions_unique <- unique(D3$ecoregion)
years_unique <- c(2017:2025)

# Clean D3 ===============================================

D3 <- D3 %>%
    mutate(
        classname = case_when(
            superClass == "Reptilia" & is.na(D3$classname) ~ 'Reptilia',
            .default = classname
        )
    )

# Create a summary data empty ==============================
summaryData <- expand.grid(years_unique, ecoregions_unique) %>%
    rename(year = Var1, ecoregion = Var2) %>%
    data.frame() %>%
    mutate(
        Aves = NA,
        Elasmobranchii = NA,
        Holocephali = NA,
        Mammalia = NA,
        Reptilia = NA,
        Teleostei = NA,
        Chondrostei = NA,
        Petromyzonti = NA
    )

# Count the number of bycatch by Year / Ecoregion / Order
# The loop goes row by row to fill the summurized information

for (II in 1:nrow(summaryData)) {
    dataRef <- D3[
        which(
            D3$ecoregion == summaryData$ecoregion[II] & D3$year == summaryData$year[II]
        ),
    ]

    summaryData$Aves[II] <- sum(
        dataRef$individualsWithPingers[which(dataRef$classname == "Aves")],
        na.rm = TRUE
    ) +
        sum(
            dataRef$individualsWithoutPingers[which(dataRef$classname == "Aves")],
            na.rm = TRUE
        )

    summaryData$Elasmobranchii[II] <- sum(
        dataRef$individualsWithPingers[which(dataRef$classname == "Elasmobranchii")],
        na.rm = TRUE
    ) +
        sum(
            dataRef$individualsWithoutPingers[which(
                dataRef$classname == "Elasmobranchii"
            )],
            na.rm = TRUE
        )

    summaryData$Holocephali[II] <- sum(
        dataRef$individualsWithPingers[which(dataRef$classname == "Holocephali")],
        na.rm = TRUE
    ) +
        sum(
            dataRef$individualsWithoutPingers[which(dataRef$classname == "Holocephali")],
            na.rm = TRUE
        )

    summaryData$Mammalia[II] <- sum(
        dataRef$individualsWithPingers[which(dataRef$classname == "Mammalia")],
        na.rm = TRUE
    ) +
        sum(
            dataRef$individualsWithoutPingers[which(dataRef$classname == "Mammalia")],
            na.rm = TRUE
        )

    summaryData$Reptilia[II] <- sum(
        dataRef$individualsWithPingers[which(dataRef$classname == "Reptilia")],
        na.rm = TRUE
    ) +
        sum(
            dataRef$individualsWithoutPingers[which(dataRef$classname == "Reptilia")],
            na.rm = TRUE
        )

    summaryData$Teleostei[II] <- sum(
        dataRef$individualsWithPingers[which(dataRef$classname == "Teleostei")],
        na.rm = TRUE
    ) +
        sum(
            dataRef$individualsWithoutPingers[which(dataRef$classname == "Teleostei")],
            na.rm = TRUE
        )

    summaryData$Chondrostei[II] <- sum(
        dataRef$individualsWithPingers[which(dataRef$classname == "Chondrostei")],
        na.rm = TRUE
    ) +
        sum(
            dataRef$individualsWithoutPingers[which(dataRef$classname == "Chondrostei")],
            na.rm = TRUE
        )

    summaryData$Petromyzonti[II] <- sum(
        dataRef$individualsWithPingers[which(dataRef$classname == "Petromyzonti")],
        na.rm = TRUE
    ) +
        sum(
            dataRef$individualsWithoutPingers[which(dataRef$classname == "Petromyzonti")],
            na.rm = TRUE
        )

    rm(dataRef)
}


# Restructure data to plot
# transform the data to have classname as a column
data_toPlot <- data.frame(
    ecoregion = rep(summaryData$ecoregion, 8),
    year = rep(summaryData$year, 8),
    className = c(
        rep('Aves', nrow(summaryData)),
        rep('Elasmobranchii', nrow(summaryData)),
        rep('Holocephali', nrow(summaryData)),
        rep('Mammalia', nrow(summaryData)),
        rep('Reptilia', nrow(summaryData)),
        rep('Teleostei', nrow(summaryData)),
        rep('Chondrostei', nrow(summaryData)),
        rep('Petromyzonti', nrow(summaryData))
    ),
    bycatch_numInd = c(
        summaryData$Aves,
        summaryData$Elasmobranchii,
        summaryData$Holocephali,
        summaryData$Mammalia,
        summaryData$Reptilia,
        summaryData$Teleostei,
        summaryData$Chondrostei,
        summaryData$Petromyzonti
    )
)

# shorten some names =========================================
data_toPlot <-
    data_toPlot %>%
    mutate(
        ecoregion = case_when(
            ecoregion == 'Aegean-Levantine Sea' ~ 'Aegean-Levantine Sea',
            ecoregion ==
                'Bay of Biscay and the Iberian Coast' ~ 'Bay Biscay & Iberian Coast',
            ecoregion ==
                'Ionian Sea and the Central Mediterranean Sea' ~ 'Ionian & Central Mediterranean',
            ecoregion == 'Western Mediterranean Sea' ~ 'Western Mediterranean',
            ecoregion == 'Oceanic Northeast Atlantic' ~ 'Oceanic NE Atlantic',
            .default = ecoregion
        )
    )

# From this table, create sub-dataset
# birds, mammals, turtles
data_toPlot_noFish <- data_toPlot[
    which(
        data_toPlot$className == 'Aves' |
            data_toPlot$className == 'Mammalia' |
            data_toPlot$className == 'Reptilia'
    ),
]
# sharks
data_toPlot_elasmo <- data_toPlot[which(data_toPlot$className == 'Elasmobranchii'), ]
# All pets
data_toPlot_allPETS <- data_toPlot[
    which(
        data_toPlot$className == 'Aves' |
            data_toPlot$className == 'Mammalia' |
            data_toPlot$className == 'Reptilia' |
            data_toPlot$className == 'Elasmobranchii'
    ),
]
## split data in two
data_toPlot_allPETS_a <- data_toPlot_allPETS[
    which(
        data_toPlot_allPETS$ecoregion == 'Adriatic Sea' |
            data_toPlot_allPETS$ecoregion == 'Aegean-Levantine Sea' |
            data_toPlot_allPETS$ecoregion == 'Azores' |
            data_toPlot_allPETS$ecoregion == 'Baltic Sea' |
            data_toPlot_allPETS$ecoregion == 'Barents Sea'
    ),
]

data_toPlot_allPETS_b <- data_toPlot_allPETS[
    which(
        data_toPlot_allPETS$ecoregion == 'Bay Biscay & Iberian Coast' |
            data_toPlot_allPETS$ecoregion == 'Black Sea' |
            data_toPlot_allPETS$ecoregion == 'Celtic Seas' |
            data_toPlot_allPETS$ecoregion == 'Faroes' |
            data_toPlot_allPETS$ecoregion == 'Greater North Sea' |
            data_toPlot_allPETS$ecoregion == 'Greenland Sea'
    ),
]
data_toPlot_allPETS_b$bycatch_numInd[which(
    data_toPlot_allPETS_b$bycatch_numInd > 10000
)] <- 10000


data_toPlot_allPETS_c <- data_toPlot_allPETS[
    which(
        data_toPlot_allPETS$ecoregion == 'Icelandic Waters' |
            data_toPlot_allPETS$ecoregion == 'Ionian & Central Mediterranean' |
            data_toPlot_allPETS$ecoregion == 'North West Atlantic' |
            data_toPlot_allPETS$ecoregion == 'Norwegian Sea' |
            data_toPlot_allPETS$ecoregion == 'Oceanic NE Atlantic' |
            data_toPlot_allPETS$ecoregion == 'Western Mediterranean'
    ),
]

data_toPlot_allPETS_list = list(
    "data_toPlot_allPETS_a" = data_toPlot_allPETS_a,
    "data_toPlot_allPETS_b" = data_toPlot_allPETS_b,
    "data_toPlot_allPETS_c" = data_toPlot_allPETS_c
)
```

## plot 

```{r bycatch_reporting_plot,cache=TRUE,eval=T,echo=T,include=T}

## Number bycatch birds / mammals / turtles

plt <- ggplot(
    data_toPlot_noFish,
    aes(x = year, y = bycatch_numInd, fill = className, label = className)
) +
    geom_col(stat = 'identity', position = 'dodge', width = 0.75) +
    scale_fill_manual(
        name = '',
        values = c('Aves' = '#F8766D', 'Mammalia' = '#00BA38', 'Reptilia' = '#619CFF')
    ) +

    facet_wrap(~ecoregion, ncol = 3, nrow = 6) +

    labs(fill = " ") +
    ylab("Number reported bycatch") +
    xlab(" ") +

    scale_x_continuous(breaks = years_unique) +

    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom",
        panel.grid.minor.x = element_blank(),
        panel.background = element_blank()
    ) +
    guides(fill = guide_legend(nrow = 1))
print(plt)
textcaption <- ""
print(textcaption)
filename <- here("outputs", "numBycatch_birdsMammalTurtle.png")
png(filename, res = 300, width = 14, height = 18, unit = "cm")
print(plt)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 14,
        height = 18,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))


# Elasmo ==================================================================

plt <- ggplot(
    data_toPlot_elasmo,
    aes(x = year, y = bycatch_numInd, fill = className, label = className)
) +
    geom_col(stat = 'identity', position = 'dodge', width = 0.5) +
    scale_fill_manual(name = '', values = c('Elasmobranchii' = "#E76BF3")) +

    facet_wrap(~ecoregion, ncol = 3, nrow = 6) +

    labs(fill = " ") +
    ylab("Number reported bycatch") +
    xlab(" ") +

    scale_x_continuous(breaks = years_unique) +
    # scale_y_continuous(trans=scales::pseudo_log_trans(base = 10), breaks = c(0, 100, 1000, 10000, 60000)) +

    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        legend.position = "bottom",
        panel.grid.minor.x = element_blank(),
        panel.background = element_blank()
    ) +
    guides(fill = guide_legend(nrow = 1))
print(plt)
textcaption <- ""
print(textcaption)
filename <- here("outputs", "numBycatch_elasmo.png")
png(filename, res = 300, width = 14, height = 18, unit = "cm")
print(plt)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 14,
        height = 18,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))


# All pets ==================================================================

# fix y-axis

range(data_toPlot_allPETS$bycatch_numInd[which(data_toPlot_allPETS$className == 'Aves')])
range(data_toPlot_allPETS$bycatch_numInd[which(
    data_toPlot_allPETS$className == 'Mammalia'
)])
range(data_toPlot_allPETS$bycatch_numInd[which(
    data_toPlot_allPETS$className == 'Reptilia'
)])
range(data_toPlot_allPETS$bycatch_numInd[which(
    data_toPlot_allPETS$className == 'Elasmobranchii'
)])

dummyData_yaxis_a <- data.frame(
    ecoregion = rep(
        c('Adriatic Sea', 'Aegean-Levantine Sea', 'Azores', 'Baltic Sea', 'Barents Sea'),
        4
    ),
    year = rep(2017, (5 * 4)),
    className = c(
        rep('Aves', 5),
        rep('Mammalia', 5),
        rep('Reptilia', 5),
        rep('Elasmobranchii', 5)
    ),
    bycatch_numInd = c(rep(1050, 5), rep(1000, 5), rep(170, 5), rep(10000, 5))
) #set y-axis max for each species

dummyData_yaxis_b <- data.frame(
    ecoregion = rep(
        c(
            'Bay Biscay & Iberian Coast',
            'Black Sea',
            'Celtic Seas',
            'Faroes',
            'Greater North Sea',
            'Greenland Sea'
        ),
        4
    ),
    year = rep(2017, (6 * 4)),
    className = c(
        rep('Aves', 6),
        rep('Mammalia', 6),
        rep('Reptilia', 6),
        rep('Elasmobranchii', 6)
    ),
    bycatch_numInd = c(rep(1050, 6), rep(1000, 6), rep(170, 6), rep(10000, 6))
) #set y-axis max for each species

dummyData_yaxis_c <- data.frame(
    ecoregion = rep(
        c(
            'Icelandic Waters',
            'Ionian & Central Mediterranean',
            'North West Atlantic',
            'Norwegian Sea',
            'Oceanic NE Atlantic',
            'Western Mediterranean'
        ),
        4
    ),
    year = rep(2017, (6 * 4)),
    className = c(
        rep('Aves', 6),
        rep('Mammalia', 6),
        rep('Reptilia', 6),
        rep('Elasmobranchii', 6)
    ),
    bycatch_numInd = c(rep(1050, 6), rep(1000, 6), rep(170, 6), rep(10000, 6))
) #set y-axis max for each species

dummyData_yaxis_list = list(
    "dummyData_yaxis_a" = dummyData_yaxis_a,
    "dummyData_yaxis_b" = dummyData_yaxis_b,
    "dummyData_yaxis_c" = dummyData_yaxis_c
)


# Make plot =================
# Loop to produce the figure a/b/c

lapply(1:3, function(x) {
    plt <- ggplot(
        data_toPlot_allPETS_list[[x]],
        aes(x = year, y = bycatch_numInd, fill = className, label = className)
    ) +
        geom_col(stat = 'identity', position = 'dodge', width = 0.75) +
        scale_fill_manual(
            name = '',
            values = c(
                'Aves' = '#F8766D',
                'Mammalia' = '#00BA38',
                'Reptilia' = '#619CFF',
                'Elasmobranchii' = '#E76BF3'
            ),
            labels = c(
                'Aves' = 'Aves',
                'Mammalia' = 'Mammalia',
                'Reptilia' = 'Reptilia',
                'Elasmobranchii' = 'Elasmobranchii'
            )
        ) +

        facet_grid(className ~ ecoregion, scales = "free") +
        geom_point(data = dummyData_yaxis_list[[x]], col = 'white', alpha = 0) +

        labs(fill = " ") +
        ylab(" ") +
        xlab(" ") +

        scale_x_continuous(breaks = years_unique) +

        theme_bw() +
        theme(
            axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
            legend.position = "right",
            panel.grid.minor.x = element_blank(),
            panel.background = element_blank()
        ) +
        guides(fill = guide_legend(nrow = 4))
}) -> data_toPlot_allPETS_plot

textcaption <- "Bycatch reporting (individual numbers) by ecoregion from 2017 to 2024, across all metiers. Note that in 2017 and 2018 numbers are capped at 14,000 from 66,524 and 19,206 respectively to improve visualization"
print(textcaption)
filename <- here("outputs", glue(list_table$data_toPlot_allPETS_plot[1], "_numBycatch_allPETSplots.png"))
png(filename, res = 300, width = 35, height = 37, unit = "cm")
plotLayout = rbind(c(1), c(2), c(3))
grid.arrange(
    data_toPlot_allPETS_plot[[1]],
    data_toPlot_allPETS_plot[[2]],
    data_toPlot_allPETS_plot[[3]],
    ncol = 1,
    layout_matrix = plotLayout
)
dev.off()
read_docx() |>
    body_add_img(
        src = filename,
        width = 15,
        height = 15,
        unit = "cm",
        style = "centered"
    ) |>
    body_add_par(value = textcaption, style = "centered") |>
    print(target = sub(".png", ".docx", filename))
```

# Bycatch reporting per country

## Prepare data per country

```{r bycatch_report_by_country_data, cache=TRUE, eval=T, echo=T,include=T}
country_unique <- unique(D3$country)
years_unique <- c(2017:2025)

D3 <- 
  D3 %>% 
  mutate(classname = case_when(superClass == "Reptilia" & is.na(D3$classname) ~ 'Reptilia', 
                               .default = classname))
# Create a summary data empty 
summaryData <- expand.grid(years_unique, country_unique) %>% 
  rename(year = Var1, country = Var2) %>% 
  data.frame() %>% 
  mutate(Aves = NA, 
         Elasmobranchii = NA, 
         Holocephali = NA, 
         Mammalia = NA, 
         Reptilia = NA, 
         Teleostei = NA, 
         Chondrostei = NA, 
         Petromyzonti = NA) 

# Count the number of bycatch by Year / Ecoregion / Order =======
# The loop goes row by row to fill the summurized information

for(II in 1:nrow(summaryData)){
  dataRef <- D3[which(D3$country == summaryData$country[II] & D3$year == summaryData$year[II]), ]
  
  summaryData$Aves[II] <- sum(dataRef$individualsWithPingers[which(dataRef$classname == "Aves")], na.rm = TRUE) + 
    sum(dataRef$individualsWithoutPingers[which(dataRef$classname == "Aves")], na.rm = TRUE)
  
  summaryData$Elasmobranchii[II] <- sum(dataRef$individualsWithPingers[which(dataRef$classname == "Elasmobranchii")], na.rm = TRUE) + 
    sum(dataRef$individualsWithoutPingers[which(dataRef$classname == "Elasmobranchii")], na.rm = TRUE)
  
  summaryData$Holocephali[II] <- sum(dataRef$individualsWithPingers[which(dataRef$classname == "Holocephali")], na.rm = TRUE) + 
    sum(dataRef$individualsWithoutPingers[which(dataRef$classname == "Holocephali")], na.rm = TRUE)
  
  summaryData$Mammalia[II] <- sum(dataRef$individualsWithPingers[which(dataRef$classname == "Mammalia")], na.rm = TRUE) + 
    sum(dataRef$individualsWithoutPingers[which(dataRef$classname == "Mammalia")], na.rm = TRUE)
  
  summaryData$Reptilia[II] <- sum(dataRef$individualsWithPingers[which(dataRef$classname == "Reptilia")], na.rm = TRUE) + 
    sum(dataRef$individualsWithoutPingers[which(dataRef$classname == "Reptilia")], na.rm = TRUE)
  
  summaryData$Teleostei[II] <- sum(dataRef$individualsWithPingers[which(dataRef$classname == "Teleostei")], na.rm = TRUE) + 
    sum(dataRef$individualsWithoutPingers[which(dataRef$classname == "Teleostei")], na.rm = TRUE)
  
  summaryData$Chondrostei[II] <- sum(dataRef$individualsWithPingers[which(dataRef$classname == "Chondrostei")], na.rm = TRUE) + 
    sum(dataRef$individualsWithoutPingers[which(dataRef$classname == "Chondrostei")], na.rm = TRUE)
  
  summaryData$Petromyzonti[II] <- sum(dataRef$individualsWithPingers[which(dataRef$classname == "Petromyzonti")], na.rm = TRUE) + 
    sum(dataRef$individualsWithoutPingers[which(dataRef$classname == "Petromyzonti")], na.rm = TRUE)
  
  rm(dataRef)
}


# Restructure data to plot =================================================
# transform the data to have classname as a column
data_toPlot <- data.frame(country = rep(summaryData$country, 8),
                          year = rep(summaryData$year, 8),
                          className = c(rep('Aves', nrow(summaryData)),
                                        rep('Elasmobranchii', nrow(summaryData)),
                                        rep('Holocephali', nrow(summaryData)),
                                        rep('Mammalia', nrow(summaryData)),
                                        rep('Reptilia', nrow(summaryData)),
                                        rep('Teleostei', nrow(summaryData)),
                                        rep('Chondrostei', nrow(summaryData)),
                                        rep('Petromyzonti', nrow(summaryData))),
                          bycatch_numInd = c(summaryData$Aves, 
                                             summaryData$Elasmobranchii, 
                                             summaryData$Holocephali, 
                                             summaryData$Mammalia,
                                             summaryData$Reptilia, 
                                             summaryData$Teleostei,
                                             summaryData$Chondrostei, 
                                             summaryData$Petromyzonti))

# From this table, create sub-dataset ===========================
#birds, mammals, turtles ====
data_toPlot_noFish <- data_toPlot[which(data_toPlot$className == 'Aves' |
                                          data_toPlot$className == 'Mammalia' |
                                          data_toPlot$className == 'Reptilia'), ]
# sharks and shit ========
data_toPlot_elasmo <- data_toPlot[which(data_toPlot$className == 'Elasmobranchii'), ]
# All pets ===========
data_toPlot_allPETS <- data_toPlot[which(data_toPlot$className == 'Aves' |
                                           data_toPlot$className == 'Mammalia' |
                                           data_toPlot$className == 'Reptilia' |
                                           data_toPlot$className == 'Elasmobranchii'), ]
```


# Plot 

```{r bycatch_report_by_country_plot, cache=TRUE, eval=T, echo=T, include=T}

## number bycatch birds / mammals / turtles =============

plotOut_bmt <- ggplot(data_toPlot_noFish,aes(x = year,y = bycatch_numInd,
                                             fill = className,
                                             label = className)) +
  geom_col(stat = 'identity', position = 'dodge', width = 0.75) +
  scale_fill_manual(name = '', values = c('Aves' = '#F8766D', 
                                          'Mammalia' = '#00BA38',
                                          'Reptilia' = '#619CFF')) +
  facet_wrap(~country) +
  labs(fill = " ") +
  ylab("Number reported bycatch") +
  xlab(" ") +
  scale_x_continuous(breaks = years_unique) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom",
        panel.grid.minor.x = element_blank(),
        panel.background = element_blank()) +
  guides(fill = guide_legend(nrow = 1))
filename <- here("outputs", "numBycatch_birdsMammalTurtle_country.png")
png(filename = filename, width = 75, height = 38, res = 300, units = "cm")
plotOut_bmt
dev.off()


# Elasmo ==================================================================
plotOut_elasmo <- ggplot(data_toPlot_elasmo,aes(x = year,y = bycatch_numInd,
                                                fill = className,
                                                label = className)) +
  geom_col(stat = 'identity', position = 'dodge', width = 0.5) +
  scale_fill_manual(name = '', values = c('Elasmobranchii' = "#E76BF3")) +
  facet_wrap(~country) +
  labs(fill = " ") +
  ylab("Number reported bycatch") +
  xlab(" ") +
  scale_x_continuous(breaks = years_unique) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        legend.position = "bottom",
        panel.grid.minor.x = element_blank(),
        panel.background = element_blank()) +
  guides(fill = guide_legend(nrow = 1))
filename <- here("outputs", "numBycatch_elasmo_country.png")
png(filename = filename, width = 75, height = 38, res = 300, units = "cm")
plotOut_elasmo
dev.off()

# All pets ==================================================================

plotOut_all_a <- ggplot(data_toPlot_allPETS, 
                          aes(x = year,
                              y = bycatch_numInd,
                              fill = className,
                              label = className)) +
    geom_col(stat = 'identity', position = 'dodge', width = 0.75) +
    scale_fill_manual(name = '', values = c('Aves' = '#F8766D', 
                                            'Mammalia' = '#00BA38',
                                            'Reptilia' = '#619CFF',
                                            'Elasmobranchii' = '#E76BF3'),
                      labels = c('Aves' = 'Aves', 
                                 'Mammalia' = 'Mammalia',
                                 'Reptilia' = 'Reptilia',
                                 'Elasmobranchii' = 'Elasmobranchii')) +
    
    facet_grid(className ~ country, scales = "free") +
    labs(fill = " ") +
    ylab(" ") +
    xlab(" ") +
    scale_x_continuous(breaks = years_unique) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
          legend.position = "right",
          panel.grid.minor.x = element_blank(),
          panel.background = element_blank()) +
    guides(fill = guide_legend(nrow = 4))
filename <- here("outputs", "numBycatch_allPETSplots_country.png")
png(filename = filename, width = 75, height = 38, res = 300, units = "cm")
plotOut_all_a
dev.off()

```


# Stuff 

Text

```{r anotherchunk,cache=TRUE,eval=T,echo=T,include=T}
#code
```



# R session information

```{r Rsession,cache=TRUE,eval=T,echo=T,include=T}
sessioninfo::session_info()
```

# Bibliography
