---
title: "Data Checks WGBYC 2026 - Country: `r params$country` "
author: "ICES WGBYC Database Subgroup"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: html_document
params:
  country: "UNKNOWN"

---


<style>

h1 {
  background-color: #d9f2d9;
  padding: 10px;
  border-radius: 6px;
}

h2 {
  background-color: #d9f2d9;
  padding: 10px;
  border-radius: 6px;
}

</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, results = "asis")
# Load libraries
library(tidyr)
library(dplyr)
library(tidyverse)
library(sf)
library(ggplot2)

# Check icon function
check_icon <- function(x) {
  if (isTRUE(x)) {
    "<span style='color: green; font-weight: bold;'>✅</span>"
  } else {
    "<span style='color: red; font-weight: bold;'>❌</span>"
  }
}

# Load gis info
ices_rect <- st_read("../data/ICES_areas/ICES_Areas_20160601_cut_dense_3857.shp")

# Load metier hierarchy, extracted from Decision 2008/949/EC
metier_hierarchy <- readxl::read_excel(path = "../data/ICES_metiers_L3_L4_L5_hierarchy_Decision_2008_949_ICES_RECO.xlsx",
                                       sheet = "Hierarchy")

```


## D1 Table - Fishing effort
### All vessel sizes are reported, they do not overlap & there is no unknown (NK) reported
```{r , echo = FALSE, results = "asis", warning = FALSE}
lengths <- gsub("^VL", "", sort(unique(D1_country$vesselLengthRange)))

lower <- as.numeric(substr(lengths, 1, 2))
upper_raw <- substr(lengths, 3, 4)
upper <- ifelse(upper_raw == "XX", Inf, as.numeric(upper_raw))

vl_ranges <- data.frame(
    code = sort(unique(D1_country$vesselLengthRange)),
    lower = lower,
    upper = upper)

# Add frequencies
freq <- as.data.frame(table(D1_country$vesselLengthRange))
names(freq) <- c("code", "n")

vl_ranges <- merge(vl_ranges, freq, by = "code", all.x = TRUE)

overlap <- !any(vl_ranges$lower[-1] < vl_ranges$upper[-nrow(vl_ranges)], na.rm = T)
no_NK <- !"NK" %in% D1_country$vesselLengthRange

continuous <- all(
  vl_ranges$upper[-nrow(vl_ranges)] == vl_ranges$lower[-1])

cat(check_icon(overlap), "No overlapping categories.", "<br>", check_icon(no_NK),
    "No unknown (NK) categories.")

knitr::kable(vl_ranges, caption = "Vessel length ranges reported")

```

### Check effort (DaS) units and range
```{r , echo = FALSE}
range(D1_country$daysAtSeaF)
```

### Information on MetierL6 is provided (not mandatory)
```{r , echo = FALSE, results = "asis"}
missing_codes <- c("", "NK", "Unknown")

metier_check <- D1_country %>%
  mutate(MetierL4_check = trimws(as.character(metierL4)),
         MetierL6_check = trimws(as.character(metierL6)),
         L4_provided = !is.na(MetierL4_check) &
            !MetierL4_check %in% missing_codes,
         L6_provided = !is.na(MetierL6_check) &
            !MetierL6_check %in% missing_codes)

metier_summary <- metier_check %>%
  filter(L4_provided) %>%
  group_by(MetierL4 = MetierL4_check) %>%
  summarise(
    n_rows = n(),
    n_L6_provided = sum(L6_provided),
    n_L6_missing = sum(!L6_provided),
    pct_L6_provided = 100 * n_L6_provided / n_rows,

    n_distinct_L6 = n_distinct(
      MetierL6_check[L6_provided]
    ),
    .groups = "drop") %>%
  arrange(MetierL4)

metier_table <- metier_summary %>%
  knitr::kable(
    format = "html",
    caption = "MetierL6 information by MetierL4, if provided"
  ) %>%
  kableExtra::kable_styling(
    bootstrap_options = c(
      "striped",
      "hover",
      "condensed",
      "responsive"
    ),
    full_width = TRUE,
    position = "left"
  )
```

### Information on VesselsF is provided (not mandatory)
```{r , echo = FALSE, results = "asis"}
missing_codes <- c("", "NK", "Unknown")

VesselsF_check <- trimws(as.character(D1_country$vesselsF))
VesselsF_provided <- !is.na(VesselsF_check) &
    !VesselsF_check %in% missing_codes

n_total <- length(VesselsF_check)
n_VesselsF_provided <- sum(VesselsF_provided)
n_VesselsF_missing <- sum(!VesselsF_provided)

pct_VesselsF_provided <- if (n_total > 0L) {
  100 * n_VesselsF_provided / n_total
} else {NA_real_}

any_VesselsF_provided <- n_VesselsF_provided > 0L

cat(
  check_icon(any_VesselsF_provided),
  sprintf(
    paste0(
      "Information on VesselsF is provided for %s of %s records ",
      "(%.1f%%). VesselsF is not mandatory."
    ),
    format(n_VesselsF_provided, big.mark = ","),
    format(n_total, big.mark = ","),
    pct_VesselsF_provided
  ),
  "<br>"
)

```

### Check unique ICES area codes & no "9999"|"-9" are reported
```{r , echo = FALSE, results = "asis"}
no_invalid_area <- !any(D1_country$areaCode %in% c(-9, 9999, "-9", "9999"))

cat(check_icon(no_invalid_area), "Valid areaCode values (-9 / 9999) have been reported.<br>")

cat("Data for", length(unique(D1_country$areaCode)), "ICES areaCode has been reported.")

# Proportion of the data missing because -9/9999
cat(" Missing ICES areaCode information represents ", round((length(which(D1_country$areaCode %in% c(-9, 9999, "-9", "9999")))/length(D1_country$areaCode))*100, 5), "% of the fishing effort data.")

#sort(unique(D1_country$statRec))
```


### Check reported metier levels are consistent (3, 4 and 5) - Commission Decision 2008/949/EC
```{r , echo = FALSE, results = "asis"}
# Inspect the column names
#names(metier_hierarchy)

# Rename hierarchy columns to match those from D's tables
valid_metiers <- metier_hierarchy %>%
  transmute(metierL3 = `L3 code`, metierL4 = `L4 code`, metierL5 = `L5 code`) %>%
  distinct()

# Make sure both df's metier codes upper cases!

# Simplest approach:
# datos_validated <- D1_country %>%
#   left_join(valid_metiers %>%
#               mutate(valid_combination = TRUE), 
#             by = c("metierL3", "metierL4", "metierL5")) %>%
#   mutate(valid_combination = coalesce(valid_combination, FALSE))
# 
# # Get the number of cases
# datos_validated %>%
#   count(valid_combination, sort = TRUE)

# More specific validation
valid_l3 <- valid_metiers %>%
  distinct(metierL3) %>%
  mutate(valid_L3 = TRUE)

valid_l3_l4 <- valid_metiers %>%
  distinct(metierL3, metierL4) %>%
  mutate(valid_L3_L4 = TRUE)

valid_l3_l4_l5 <- valid_metiers %>%
  distinct(metierL3, metierL4, metierL5) %>%
  mutate(valid_L3_L4_L5 = TRUE)

datos_validated <- D1_country %>%
  left_join(valid_l3, by = "metierL3") %>%
  left_join(valid_l3_l4, by = c("metierL3", "metierL4")) %>%
  left_join(valid_l3_l4_l5, by = c("metierL3", "metierL4", "metierL5")) %>%
  mutate(
    across(c(valid_L3, valid_L3_L4, valid_L3_L4_L5), ~ coalesce(.x, FALSE)),
    validation_status = case_when(
      is.na(metierL3) | is.na(metierL4) | is.na(metierL5) ~ "Missing métier information",
      !valid_L3 ~ "Invalid L3 code",
      !valid_L3_L4 ~ "L4 code is not valid for the stated L3 code",
      !valid_L3_L4_L5 ~ "L5 code is not valid for the stated L3-L4 combination",
      TRUE ~ "Valid combination"))

# Print
#datos_validated %>% count(validation_status)

validation_summary <- datos_validated %>%
  dplyr::count(validation_status, name = "n") %>%
  dplyr::mutate(percentage = 100 * n / sum(n))

knitr::kable(validation_summary,
             col.names = c("Validation status", "Number of records", "Percentage (%)"),
             digits = 2, align = c("l", "r", "r"),
             caption = "Validation of Level 3, Level 4 and Level 5 métier combinations")

# TRUE only when every record has a valid métier combination
no_invalid_metiers_D1 <- all(datos_validated$validation_status == "Valid combination", na.rm = FALSE)

cat(check_icon(no_invalid_metiers_D1), "All records contain a valid metier L3-L4-L5 combination.<br>")


```


## D2 Table - Monitoring effort
### Monitoring type and monitoring method are provided
```{r , echo = FALSE}
tab_unique <- D2_country %>%
  distinct(monitoringMethod, monitoringProgramType)
knitr::kable(tab_unique)
```

### Number of trips is not higher than DaS
```{r , echo = FALSE, results='asis'}
check_ok <- all(D2_country$tripsOb <= D2_country$daysAtSeaOb, na.rm = TRUE)
cat(check_icon(check_ok), "TripsOb ≤ daysAtSeaOb <br>")

cat("Registers with trips <= DaS observed = ", length(which(D2_country$tripsOb <= D2_country$daysAtSeaOb)),", out of ", nrow(D2_country), " cases.")


```

### Number of trips is not equal to DaS observed
```{r , echo = FALSE, results = "asis"}
check_ok_trips <- !any(D2_country$tripsOb != D2_country$daysAtSeaOb, na.rm = TRUE)
cat(check_icon(check_ok_trips), "Number of trips is not equal to DaS observed.<br>")
cat("There is ", sum(!(D2_country$tripsOb == D2_country$daysAtSeaOb)), " out of ", nrow(D2_country), " cases where trips is not equal to DaS observed.")
```

### All metiers with monitoring effort have reported fishing effort
This is already checked during data submission, just double-checking.
```{r , echo = FALSE, results = "asis"}
metiers_obs <- D2_country %>% 
  group_by(metierL4) %>%
  summarize(das_obs = sum(daysAtSeaOb))

obs_fish <- D1_country %>%
  group_by(metierL4) %>%
  summarise(das_fish = sum(daysAtSeaF, na.rm = TRUE), .groups = "drop" )

summary_obs_fish <- metiers_obs %>% 
  left_join(obs_fish, by = c("metierL4")) %>%
  mutate(obs_less_fish = das_obs <= das_fish)
#summary_obs_fish

more_das_obs <- summary_obs_fish %>% 
  filter(obs_less_fish == FALSE) %>%
  distinct(metierL4) %>% 
  summarise(metiers = list(unique(metierL4)), .groups = "drop")

check_ok_match <- all(summary_obs_fish$obs_less_fish, na.rm = TRUE)

cat("<br>", check_icon(check_ok_match), "Observed DaS ≤ fishing DaS by métier (L4).")

# Check if any metierL4 is reported in D2 but no in D1
# Comprobar si algún métier de D2 no aparece en D1
missing_metier <- any(is.na(summary_obs_fish$obs_less_fish))

metiers_missing_D1 <- summary_obs_fish %>% 
  filter(is.na(obs_less_fish)) %>%
  distinct(metierL4)

cat("<br>", check_icon(!missing_metier), 
    if (!missing_metier) {"All métiers with monitoring effort in D2 have reported fishing effort in D1."} else {
    paste0("The following métier(s) have monitoring effort in D2 , but are not reported in D1: ",
      paste(metiers_missing_D1$metierL4, collapse = ", "),".")},
  "<br>"
)

```

### Information on VesselsObs is provided
```{r , echo = FALSE}
missing_codes <- c("", "NK", "Unknown")

VesselsObs_check <- trimws(as.character(D2_country$vesselsOb))
VesselsObs_provided <- !is.na(VesselsObs_check) &
    !VesselsObs_check %in% missing_codes

n_total <- length(VesselsObs_check)
n_VesselsObs_provided <- sum(VesselsObs_provided)
n_VesselsObs_missing <- sum(!VesselsObs_provided)

pct_VesselsObs_provided <- if (n_total > 0L) {
  100 * n_VesselsObs_provided / n_total
} else {NA_real_}

any_VesselsObs_provided <- n_VesselsObs_provided > 0L

cat(
  check_icon(any_VesselsObs_provided),
  sprintf(
    paste0(
      "Information on VesselsObs is provided for %s of %s records ",
      "(%.1f%%)."
    ),
    format(n_VesselsObs_provided, big.mark = ","),
    format(n_total, big.mark = ","),
    pct_VesselsObs_provided
  ),
  "<br>"
)


```

### Check effort (DaS) units and range
```{r , echo = FALSE}
range(D2_country$daysAtSeaOb)
```

### Check unique ICES area codes & no "9999"|"-9" are reported
```{r , echo = FALSE, results = "asis"}
no_invalid_area_D2 <- !any(D2_country$areaCode %in% c(-9, 9999, "-9", "9999"))

cat(check_icon(no_invalid_area_D2), "Valid areaCode values (-9 / 9999) have been reported.<br>")

cat("Monitoring data for", length(unique(D2_country$areaCode)), "ICES areas has been reported.")

# Proportion of the data missing because -9/9999
cat("Missing ICES rectangles information represents ", round((length(which(D2_country$areaCode %in% c(-9, 9999, "-9", "9999")))/length(D2_country$areaCode))*100, 3), "% of the monitoring effort data.")


```


### Monitoring & fishing effort data reported per year
```{r , echo = FALSE, results="asis", warning = FALSE, fig.width = 16, fig.height = 20, out.width = "100%"}
# First merge both DaS & DaSObs
das_fish <- D1_country %>%
  group_by(metierL4, year) %>%
  summarise(das_fish = sum(daysAtSeaF, na.rm = TRUE), .groups = "drop" )
#das_fish

das_obs <- D2_country %>%
  group_by(metierL4, year) %>%
  summarise(das_obs = sum(daysAtSeaOb, na.rm = TRUE), .groups = "drop" )
#das_obs

summary_das <- das_fish %>% 
  full_join(das_obs, by = c("metierL4", "year")) %>%
  mutate(obs_less_fish = das_obs <= das_fish)

# Check if das_obs <= das_fish per metier & per year
check_ok_monitor <- all(summary_das$das_obs <= summary_das$das_fish, na.rm = TRUE)


cat(check_icon(check_ok_monitor), "Observed DaS ≤ fishing DaS (by metier and year) ")

# Plot both efforts per year

das_long <- summary_das %>%
  pivot_longer(cols = c(das_obs, das_fish),
    names_to = "das_type",
    values_to = "das")

# Add the number of incidents reported per metier and year

incidents_df <- D3_country %>% 
  group_by(metierL4, year) %>% 
  summarise(
    inc_with = sum(incidentsWithPingers[incidentsWithPingers > 0], na.rm = T),
    inc_without = sum(incidentsWithoutPingers[incidentsWithoutPingers > 0], na.rm = T),
    n_incidents = inc_with+inc_without,
    .groups = "drop")
  
  # filter(incidentsWithPingers >= 1) %>% 
  # count(metierL4, year, name = "n_incidents")

ggplot(das_long, aes(x = factor(year), y = das, fill = das_type)) +
  geom_col() +
  facet_wrap(~ metierL4+das_type, scales = "free_y",
             labeller = label_wrap_gen(multi_line=T)) +
  geom_text(data = incidents_df, aes(x = factor(year), y = Inf, 
                                     label = n_incidents, color = "n_incidents"), 
            inherit.aes = FALSE, vjust = 1.2, size = 3.5, fontface = "bold",
            family = "sans") +
  labs(x = "Year", y = "Days at sea", fill = "") +
  scale_fill_manual(values = c("das_obs" = "steelblue", "das_fish" = "grey70"),
    labels = c("Fishing DaS", "Observed DaS")) +
  scale_color_manual(name = "", values = c("n_incidents" = "firebrick"),
  labels = c("Number of incidents")) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```


### Check reported metier levels are consistent (3, 4 and 5) - Commission Decision 2008/949/EC
```{r , echo = FALSE, results = "asis"}
datos_validated <- D2_country %>%
  left_join(valid_l3, by = "metierL3") %>%
  left_join(valid_l3_l4, by = c("metierL3", "metierL4")) %>%
  left_join(valid_l3_l4_l5, by = c("metierL3", "metierL4", "metierL5")) %>%
  mutate(
    across(c(valid_L3, valid_L3_L4, valid_L3_L4_L5), ~ coalesce(.x, FALSE)),
    validation_status = case_when(
      is.na(metierL3) | is.na(metierL4) | is.na(metierL5) ~ "Missing métier information",
      !valid_L3 ~ "Invalid L3 code",
      !valid_L3_L4 ~ "L4 code is not valid for the stated L3 code",
      !valid_L3_L4_L5 ~ "L5 code is not valid for the stated L3-L4 combination",
      TRUE ~ "Valid combination"))

# Print
#datos_validated %>% count(validation_status)

validation_summary <- datos_validated %>%
  dplyr::count(validation_status, name = "n") %>%
  dplyr::mutate(percentage = 100 * n / sum(n))

knitr::kable(validation_summary,
             col.names = c("Validation status", "Number of records", "Percentage (%)"),
             digits = 2, align = c("l", "r", "r"),
             caption = "Validation of Level 3, Level 4 and Level 5 métier combinations")

# TRUE only when every record has a valid métier combination
no_invalid_metiers_D2 <- all(datos_validated$validation_status == "Valid combination", na.rm = FALSE)

cat(check_icon(no_invalid_metiers_D2), "All records contain a valid metier L3-L4-L5 combination.<br>")
```


## D3 Table - Bycatch events
### Note which species are included among the bycatch events & the spelling of the scientific names
```{r , echo = FALSE}
has_bycatch <- nrow(D3_country) > 0 &&
               any(!is.na(D3_country$species))

if(!has_bycatch){cat("No bycatch events have been reported.")}else{
  cat(paste(sort(unique(na.omit(D3_country$species))),collapse = ", "))}

```

### Check monitoring protocols reported
```{r , echo = FALSE}

if(!has_bycatch){
  cat("No monitoring protocols reported because no bycatch events occurred.")}else{
  protocols_sps <- D3_country %>%  
    group_by(samplingProtocol) %>% 
    summarise(n_species = n_distinct(species, na.rm = TRUE),.groups = "drop")
  knitr::kable(protocols_sps,
               caption = "Monitoring protocols and number of bycatch species reported")
}

```

### Fields IndividualsWithPingers, IndividualsWithoutPingers, IncidentsWithoutPingers & IncidentsWithPingers do not contain values representing unknown information or not valid values (e.g. 99999, -9)
```{r , echo = FALSE}

fields_bycatch <- c("individualsWithPingers", "individualsWithoutPingers",
                    "incidentsWithPingers", "incidentsWithoutPingers")

invalid_values <- c(-9, 99999, "-9", "99999", NA, "NA", "NK")


check_bycatch_fields <- function(D, fields) {
  res <- lapply(fields, function(f) {
    x <- D[[f]]
    invalid <- x %in% invalid_values | x < 0
    list(field   = f,
         passed  = !any(invalid, na.rm = TRUE),
         n_bad   = sum(invalid, na.rm = TRUE))
    })
  do.call(rbind, lapply(res, as.data.frame))
}

bycatch_check <- check_bycatch_fields(D3_country, fields_bycatch)

check_ok_bycatch <- all(bycatch_check$passed)


cat(check_icon(check_ok_bycatch), "Bycatch-related fields contain no unknown or not valid values.")

knitr::kable(bycatch_check)

```

### Check that when incidents are reported, individuals are reported
```{r , echo = FALSE}
incidents_check <- D3_country %>%
  mutate(check_with_pingers = incidentsWithPingers > 0 & individualsWithPingers == 0,
         check_without_pingers = incidentsWithoutPingers > 0 & individualsWithoutPingers == 0)
# Problematic records

incidents_problems <- incidents_check %>%
  filter(check_with_pingers | check_without_pingers)

check_inc_ok <- nrow(incidents_problems) == 0

cat(check_icon(check_inc_ok), "When incidents are reported, individuals are reported.",
    if(!check_inc_ok)
    paste0(" (", nrow(incidents_problems), " row(s) present issues, out of ", nrow(D3_country), ").") else "")

```

### Check that all rows reported have data
```{r, echo = FALSE}
cols_to_check <- c("individualsWithPingers", "individualsWithoutPingers",
                   "incidentsWithPingers", "incidentsWithoutPingers")

all_rows_data <- D3_country %>% 
  mutate(
    any_zero = any(c_across(all_of(cols_to_check)) > 0, na.rm = TRUE))

zero_rows <- D3_country %>%
  filter(rowSums(across(all_of(cols_to_check), ~ . > 0), na.rm = TRUE) == 0)

check_all_rows_ok <- nrow(zero_rows) == 0

cat(check_icon(check_all_rows_ok), "All rows reported have data.", 
    if(!check_all_rows_ok)
      paste0("", nrow(zero_rows), " row(s) are reported with no bycatch data.")
    else "")
```

### Check unique ICES area codes & no “9999”|“-9” are reported
```{r , echo = FALSE}
no_invalid_area_D3 <- !any(D3_country$areaCode %in% c(-9, 9999, "-9", "9999"))

cat(check_icon(no_invalid_area_D3), "Valid areaCode values (-9 / 9999) have been reported.<br>")

cat("Bycatch data for", length(unique(D3_country$areaCode)), "ICES areaCode has been reported.")

```

### Check reported metier levels are consistent (3, 4 and 5) - Commission Decision 2008/949/EC
```{r , echo = FALSE, results = "asis"}
datos_validated <- D3_country %>%
  left_join(valid_l3, by = "metierL3") %>%
  left_join(valid_l3_l4, by = c("metierL3", "metierL4")) %>%
  left_join(valid_l3_l4_l5, by = c("metierL3", "metierL4", "metierL5")) %>%
  mutate(
    across(c(valid_L3, valid_L3_L4, valid_L3_L4_L5), ~ coalesce(.x, FALSE)),
    validation_status = case_when(
      is.na(metierL3) | is.na(metierL4) | is.na(metierL5) ~ "Missing métier information",
      !valid_L3 ~ "Invalid L3 code",
      !valid_L3_L4 ~ "L4 code is not valid for the stated L3 code",
      !valid_L3_L4_L5 ~ "L5 code is not valid for the stated L3-L4 combination",
      TRUE ~ "Valid combination"))

# Print
#datos_validated %>% count(validation_status)

validation_summary <- datos_validated %>%
  dplyr::count(validation_status, name = "n") %>%
  dplyr::mutate(percentage = 100 * n / sum(n))

knitr::kable(validation_summary,
             col.names = c("Validation status", "Number of records", "Percentage (%)"),
             digits = 2, align = c("l", "r", "r"),
             caption = "Validation of Level 3, Level 4 and Level 5 métier combinations")

# TRUE only when every record has a valid métier combination
no_invalid_metiers_D3 <- all(datos_validated$validation_status == "Valid combination", na.rm = FALSE)

cat(check_icon(no_invalid_metiers_D3), "All records contain a valid metier L3-L4-L5 combination.<br>")
```